### 2021-05-06 prospectus/prospectus.tex
Start prospectus document
Here we go!

I'm not entirely sure what's needed for the prospectus document, but this starts filling out information based off of what was in the given example. I've sent an email to set up a meeting with my director to get more information

\documentclass[11pt]{article}

%%%%% Preamble

\usepackage[margin=1in]{geometry} % Margin size
\usepackage{supertabular}
\usepackage[export]{adjustbox}

%% Title Information.

\title{Thesis Prospectus}
\author{Dylan Lathrum}
%\date{Jan 1 2000} % By default, LaTeX uses the current date
\date{} % Don't display date

%%%%% The Document

\begin{document}

\maketitle

\section*{Topic Overview}

Personal computers have greatly evolved and specialized over the past few decades to the point where a computer can be tailored to specific needs and usages ranging from the portability of a laptop to the raw power of a desktop.
Though it hasn't been until very recently that these two classes of computers have begun to share traits with each other.
While a desktop computer has been the standard for intensive computing and work since the beginning of computers, technology has progressed to the point where a laptop can provide comparable power in a much smaller footprint.
With the recent shift in attention towards remote work and mobile computing, the prospect of taking a workstation wherever needed is incredibly promising.
Though even the emerging laptops today struggle to compete with desktops in terms of cost, maintenance, and future upgrades.
The cost barrier to entry for laptops is considerably higher compared to desktop computers, and most laptops are manufactured in a way that makes upgrading parts of the machine difficult or impossible, forcing a complete purchase in the event of failure or a component needs to be upgraded.

However, in the case where someone already owns a desktop computer and needs to be mobile, instead of needing to purchase a second device at full price, it may be possible to develop a low-cost computer that has just enough power to connect to the existing desktop and run all computing there, using the mobile device only as a user interface.
While the idea of a thin-client is not new, it's current usage in the real world has mostly been limited to computing clusters and server terminals due to their dependence on stable internet connections.
With the advent of small consumer electronics such as Raspberry Pis and their improvements in networking capabilities, the idea of a remote desktop has become a real possibility.

\section*{Research Plan}

I will begin my research by investigating the minimum cost barrier to create a computer using quality parts that can be used like a regular consumer product that has enough power to connect with a desktop computer at speeds sufficient for intense computing.
% Do I want to mention video games as a benchmark, as that would prove both processing speed as well as responsiveness?
Once parts have been selected and order, I will begin developing the hardware enclosure, wiring plans, and the software required to make the computer functional.
Throughout development, I will be benchmarking my product against existing technologies that support remote usage of computers, such as remote desktops.
I plan to investigate the specific technical specifications of the remote connection between my device and the host computer such as the latency, responsiveness, and image quality.
Due to the lack of existing all-in-one device for this purpose, I will research the advantages such a device has over comparable alternatives.
Once development has concluded, I will use the data collected to synthesize a comprehensive overview of the produced device and analyze its benefits over purchasing a high-powered laptop or using current alternatives.

\section*{Meetings}

I will meet with my thesis director on a regular basis every other week to discuss progress on the project and to answer any questions that arise from either party.
I will also meet with my other committee members to discuss my progress and ask for their advice and expertise.
Other meetings may be scheduled should new information or urgent questions arise.

\section*{Timeline}

\begin{tabular}{r|p{0.8\linewidth}}
	Date       & Task                                                                                 \\
	\hline
	           & Begin research into hardware and software to be used in the project                  \\
	           & Order parts and begin researching and designing enclosure for parts                  \\
	August 11  & Begin regular meetings with director, review schedule                                \\
	August     & Begin creating hardware and software for project                                     \\
	January 10 & Start working on thesis paper, continue research and development on physical project \\
	March 6    & Complete rough draft of thesis paper                                                 \\
	April      & Thesis Defense                                                                       \\
	           & Submit Thesis                                                                        \\
\end{tabular}

\end{document}

### 2021-05-24 prospectus/prospectus.tex
Fix typo and experiment with latex editing formatting
It's a little awkward to write regular text into VSCode

@@
 % Do I want to mention video games as a benchmark, as that would prove both processing speed as well as responsiveness?
-Once parts have been selected and order, I will begin developing the hardware enclosure, wiring plans, and the software required to make the computer functional.
+Once parts have been selected and ordered, I will begin developing the hardware enclosure, wiring plans, and the software required to make the computer functional.
 Throughout development, I will be benchmarking my product against existing technologies that support remote usage of computers, such as remote desktops.

### 2021-06-07 prospectus/prospectus.tex
Update prospectus to include PCB plan

@@
 
-Personal computers have greatly evolved and specialized over the past few decades to the point where a computer can be tailored to specific needs and usages ranging from the portability of a laptop to the raw power of a desktop.
-Though it hasn't been until very recently that these two classes of computers have begun to share traits with each other.
+Personal computers have greatly evolved and specialized over the past few decades to the point where the word "computer" now refers to a whole family of devices that can be tailored to any particular use case.
+A computer can be built to specifically fill a need or role ranging from the portability of a laptop to the raw power of a desktop.
+For the longest time these roles have been completely seperate from one another, though in the past couple of years these two classes of computers have begun to share traits with each other.
 While a desktop computer has been the standard for intensive computing and work since the beginning of computers, technology has progressed to the point where a laptop can provide comparable power in a much smaller footprint.
@@
 Though even the emerging laptops today struggle to compete with desktops in terms of cost, maintenance, and future upgrades.
-The cost barrier to entry for laptops is considerably higher compared to desktop computers, and most laptops are manufactured in a way that makes upgrading parts of the machine difficult or impossible, forcing a complete purchase in the event of failure or a component needs to be upgraded.
+The cost barrier to entry for powerful laptops is considerably higher compared to desktop computers, and most laptops are manufactured in a way that makes upgrading parts of the machine difficult or impossible, forcing a complete purchase in the event of failure or a component needs to be upgraded.
 
-However, in the case where someone already owns a desktop computer and needs to be mobile, instead of needing to purchase a second device at full price, it may be possible to develop a low-cost computer that has just enough power to connect to the existing desktop and run all computing there, using the mobile device only as a user interface.
+However, in the case where someone already owns a desktop computer and needs to be mobile, instead of needing to purchase a second device at full price, it may be possible to develop a low-cost computer that has just enough power to connect to the existing desktop and run all processing there, using the mobile device only as a user interface.
 While the idea of a thin-client is not new, it's current usage in the real world has mostly been limited to computing clusters and server terminals due to their dependence on stable internet connections.
@@
 I will begin my research by investigating the minimum cost barrier to create a computer using quality parts that can be used like a regular consumer product that has enough power to connect with a desktop computer at speeds sufficient for intense computing.
-% Do I want to mention video games as a benchmark, as that would prove both processing speed as well as responsiveness?
+Once I can build a proof-of-concept using consumer parts, I will begin researching and developing a custom Printed Circuit Board (PCB) that will produce the same hardware output as a single board computer at a small footprint.
+I believe creating a custom PCB will have benefits over a "DIY" solution as it will help prove the final product as something that could be used as a professional day-to-day replacement for a conventional powerful laptop.
 Once parts have been selected and ordered, I will begin developing the hardware enclosure, wiring plans, and the software required to make the computer functional.
@@
 I plan to investigate the specific technical specifications of the remote connection between my device and the host computer such as the latency, responsiveness, and image quality.
+These statistics can be benchmarked by attempting to run intensive programs remotely, such as training an AI model, compiling code, running CAD software, rendering images/video, or developing video game software.
 Due to the lack of existing all-in-one device for this purpose, I will research the advantages such a device has over comparable alternatives.
@@
 	\hline
-	           & Begin research into hardware and software to be used in the project                  \\
-	           & Order parts and begin researching and designing enclosure for parts                  \\
+	June       & Begin research into hardware and software to be used in the project                  \\
+	July       & Order parts and begin researching and designing enclosure for parts                  \\
 	August 11  & Begin regular meetings with director, review schedule                                \\

### 2021-06-07 prospectus/prospectus.tex
Tweak prospectus schedule

@@
 
-Personal computers have greatly evolved and specialized over the past few decades to the point where the word "computer" now refers to a whole family of devices that can be tailored to any particular use case.
+Personal computers have greatly evolved and specialized over the past few decades to the point where the word ``computer" now refers to a whole family of devices that can be tailored to any particular use case.
 A computer can be built to specifically fill a need or role ranging from the portability of a laptop to the raw power of a desktop.
@@
 	August     & Begin creating hardware and software for project                                     \\
+	December   & Create final prototype for the project                                               \\
 	January 10 & Start working on thesis paper, continue research and development on physical project \\
-	March 6    & Complete rough draft of thesis paper                                                 \\
-	April      & Thesis Defense                                                                       \\
-	           & Submit Thesis                                                                        \\
+	Feburary   & Complete rough draft of thesis paper                                                 \\
+	March      & Thesis Defense                                                                       \\
+	April 15   & Submit Thesis                                                                        \\
 \end{tabular}

### 2022-01-13 code/moonlight/src/video/x11.c
Add Stats logging to moonlight

@@ -18,6 +18,7 @@
  */
 
 #include "video.h"
+#include "stats.h"
 #include "egl.h"
 #include "ffmpeg.h"
 #ifdef HAVE_VAAPI
@@ -52,6 +53,10 @@ static int pipefd[2];
 static int display_width;
 static int display_height;
 
+int m_LastFrameNumber = 0;
+VIDEO_STATS m_ActiveWndVideoStats;
+VIDEO_STATS m_GlobalVideoStats;
+
 static int frame_handle(int pipefd) {
   AVFrame* frame = NULL;
   while (read(pipefd, &frame, sizeof(void*)) > 0);
@@ -68,6 +73,9 @@ static int frame_handle(int pipefd) {
 }
 
 int x11_init(bool vdpau, bool vaapi) {
+  SDL_zero(m_ActiveWndVideoStats);
+  SDL_zero(m_GlobalVideoStats);
+  
   XInitThreads();
   display = XOpenDisplay(NULL);
   if (!display)
@@ -161,15 +169,122 @@ int x11_setup_vaapi(int videoFormat, int width, int height, int redrawRate, void
   return x11_setup(videoFormat, width, height, redrawRate, context, drFlags | X11_VAAPI_ACCELERATION);
 }
 
+void stringifyVideoStats(VIDEO_STATS* stats, char* output)
+{
+    int offset = 0;
+
+    // Start with an empty string
+    output[offset] = 0;
+
+    if (stats->receivedFps > 0) {
+        offset += sprintf(&output[offset],
+                          "Incoming frame rate from network: %.2f FPS\n"
+                          "Decoding frame rate: %.2f FPS\n"
+                          "Rendering frame rate: %.2f FPS\n",
+                          stats->receivedFps,
+                          stats->decodedFps,
+                          stats->renderedFps);
+    }
+
+    if (stats->renderedFrames != 0) {
+        char rttString[32];
+
+        if (stats->lastRtt != 0) {
+            sprintf(rttString, "%u ms (variance: %u ms)", stats->lastRtt, stats->lastRttVariance);
+        }
+        else {
+            sprintf(rttString, "N/A");
+        }
+
+        offset += sprintf(&output[offset],
+                          "Frames dropped by your network connection: %.2f%%\n"
+                          "Average network latency: %s\n"
+                          "Average decoding time: %.2f ms\n"
[…]

### 2022-01-15 paper/main.tex
Load LaTeX template
This template is based on a template by:
Steve Gunn (http://users.ecs.soton.ac.uk/srg/softwaretools/document/templates/)
Sunil Patel (http://www.sunilpatel.co.uk/thesis-template/)
Template license:
CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Undergraduate Thesis for Barrett, The Honors College at Arizona State University
% For the degree of Software Engineering
% By Dylan Lathrum
%
% This document is based on a template by:
% Vel (vel@latextemplates.com)
% Steve Gunn (http://users.ecs.soton.ac.uk/srg/softwaretools/document/templates/)
% Sunil Patel (http://www.sunilpatel.co.uk/thesis-template/)
%
% Under the license:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[
11pt, % The default document font size, options: 10pt, 11pt, 12pt
%oneside, % Two side (alternating margins) for binding by default, uncomment to switch to one side
english, % ngerman for German
singlespacing, % Single line spacing, alternatives: onehalfspacing or doublespacing
%draft, % Uncomment to enable draft mode (no pictures, no links, overfull hboxes indicated)
%nolistspacing, % If the document is onehalfspacing or doublespacing, uncomment this to set spacing in lists to single
%liststotoc, % Uncomment to add the list of figures/tables/etc to the table of contents
%toctotoc, % Uncomment to add the main table of contents to the table of contents
%parskip, % Uncomment to add space between paragraphs
%nohyperref, % Uncomment to not load the hyperref package
headsepline, % Uncomment to get a line under the header
%chapterinoneline, % Uncomment to place the chapter title next to the number on one line
%consistentlayout, % Uncomment to change the layout of the declaration, abstract and acknowledgements pages to match the default layout
]{MastersDoctoralThesis} % The class file specifying the document structure

\usepackage[utf8]{inputenc} % Required for inputting international characters
\usepackage[T1]{fontenc} % Output font encoding for international characters

\usepackage{mathpazo} % Use the Palatino font by default

\usepackage[backend=bibtex,style=authoryear,natbib=true]{biblatex} % Use the bibtex backend with the authoryear citation style (which resembles APA)

\addbibresource{example.bib} % The filename of the bibliography

\usepackage[autostyle=true]{csquotes} % Required to generate language-dependent quotes in the bibliography

%----------------------------------------------------------------------------------------
%	MARGIN SETTINGS
%----------------------------------------------------------------------------------------

\geometry{
	paper=a4paper, % Change to letterpaper for US letter
	inner=2.5cm, % Inner margin
	outer=3.8cm, % Outer margin
	bindingoffset=.5cm, % Binding offset
	top=1.5cm, % Top margin
	bottom=1.5cm, % Bottom margin
	%showframe, % Uncomment to show how the type block is set on the page
}

%----------------------------------------------------------------------------------------
%	THESIS INFORMATION
%----------------------------------------------------------------------------------------

\thesistitle{Thesis Title} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
\supervisor{Dr. James \textsc{Smith}} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
\examiner{} % Your examiner's name, this is not currently used anywhere in the template, print it elsewhere with \examname
\degree{Doctor of Philosophy} % Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
\author{John \textsc{Smith}} % Your name, this is used in the title page and abstract, print it elsewhere with \authorname
\addresses{} % Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname

\subject{Biological Sciences} % Your subject area, this is not currently used anywhere in the template, print it elsewhere with \subjectname
\keywords{} % Keywords for your thesis, this is not currently used anywhere in the template, print it elsewhere with \keywordnames
\university{\href{http://www.university.com}{University Name}} % Your university's name and URL, this is used in the title page and abstract, print it elsewhere with \univname
\department{\href{http://department.university.com}{Department or School Name}} % Your department's name and URL, this is used in the title page and abstract, print it elsewhere with \deptname
\group{\href{http://researchgroup.university.com}{Research Group Name}} % Your research group's name and URL, this is used in the title page, print it elsewhere with \groupname
\faculty{\href{http://faculty.university.com}{Faculty Name}} % Your faculty's name and URL, this is used in the title page and abstract, print it elsewhere with \facname

\AtBeginDocument{
\hypersetup{pdftitle=\ttitle} % Set the PDF's title to your title
\hypersetup{pdfauthor=\authorname} % Set the PDF's author to your name
\hypersetup{pdfkeywords=\keywordnames} % Set the PDF's keywords to your keywords
}

\begin{document}

\frontmatter % Use roman page numbering style (i, ii, iii, iv...) for the pre-content pages

\pagestyle{plain} % Default to the plain heading style until the thesis style is called for the body content

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\begin{titlepage}
	\begin{center}

		\vspace*{.06\textheight}
		{\scshape\LARGE \univname\par}\vspace{1.5cm} % University name
		\textsc{\Large Doctoral Thesis}\\[0.5cm] % Thesis type

		\HRule \\[0.4cm] % Horizontal line
		{\huge \bfseries \ttitle\par}\vspace{0.4cm} % Thesis title
		\HRule \\[1.5cm] % Horizontal line

		\begin{minipage}[t]{0.4\textwidth}
			\begin{flushleft} \large
				\emph{Author:}\\
				\href{http://www.johnsmith.com}{\authorname} % Author name - remove the \href bracket to remove the link
			\end{flushleft}
		\end{minipage}
		\begin{minipage}[t]{0.4\textwidth}
			\begin{flushright} \large
				\emph{Supervisor:} \\
				\href{http://www.jamessmith.com}{\supname} % Supervisor name - remove the \href bracket to remove the link  
			\end{flushright}
		\end{minipage}\\[3cm]

		\vfill

		\large \textit{A thesis submitted in fulfillment of the requirements\\ for the degree of \degreename}\\[0.3cm] % University requirement text
		\textit{in the}\\[0.4cm]
		\groupname\\\deptname\\[2cm] % Research group name and department name

		\vfill

		{\large \today}\\[4cm] % Date
		%\includegraphics{Logo} % University/department logo - uncomment to place it

		\vfill
	\end{center}
\end{titlepage}

%----------------------------------------------------------------------------------------
%	DECLARATION PAGE
%----------------------------------------------------------------------------------------

\begin{declaration}
	\addchaptertocentry{\authorshipname} % Add the declaration to the table of contents
	\noindent I, \authorname, declare that this thesis titled, \enquote{\ttitle} and the work presented in it are my own. I confirm that:

	\begin{itemize}
		\item This work was done wholly or mainly while in candidature for a research degree at this University.
		\item Where any part of this thesis has previously been submitted for a degree or any other qualification at this University or any other institution, this has been clearly stated.
		\item Where I have consulted the published work of others, this is always clearly attributed.
		\item Where I have quoted from the work of others, the source is always given. With the exception of such quotations, this thesis is entirely my own work.
		\item I have acknowledged all main sources of help.
		\item Where the thesis is based on work done by myself jointly with others, I have made clear exactly what was done by others and what I have contributed myself.\\
	\end{itemize}

	\noindent Signed:\\
	\rule[0.5em]{25em}{0.5pt} % This prints a line for the signature

	\noindent Date:\\
	\rule[0.5em]{25em}{0.5pt} % This prints a line to write the date
\end{declaration}

\cleardoublepage

%----------------------------------------------------------------------------------------
%	QUOTATION PAGE
%----------------------------------------------------------------------------------------

\vspace*{0.2\textheight}

\noindent\enquote{\itshape Thanks to my solid academic training, today I can write hundreds of words on virtually any topic without possessing a shred of information, which is how I got a good job in journalism.}\bigbreak

\hfill Dave Barry

%----------------------------------------------------------------------------------------
%	ABSTRACT PAGE
%----------------------------------------------------------------------------------------

\begin{abstract}
	\addchaptertocentry{\abstractname} % Add the abstract to the table of contents
	The Thesis Abstract is written here (and usually kept to just this page). The page is kept centered vertically so can expand into the blank space above the title too\ldots
\end{abstract}

%----------------------------------------------------------------------------------------
%	ACKNOWLEDGEMENTS
%----------------------------------------------------------------------------------------

\begin{acknowledgements}
	\addchaptertocentry{\acknowledgementname} % Add the acknowledgements to the table of contents
	The acknowledgments and the people to thank go here, don't forget to include your project advisor\ldots
\end{acknowledgements}

%----------------------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES PAGES
%----------------------------------------------------------------------------------------

\tableofcontents % Prints the main table of contents

\listoffigures % Prints the list of figures

\listoftables % Prints the list of tables

%----------------------------------------------------------------------------------------
%	ABBREVIATIONS
%----------------------------------------------------------------------------------------

\begin{abbreviations}{ll} % Include a list of abbreviations (a table of two columns)

	\textbf{LAH} & \textbf{L}ist \textbf{A}bbreviations \textbf{H}ere\\
	\textbf{WSF} & \textbf{W}hat (it) \textbf{S}tands \textbf{F}or\\

\end{abbreviations}

%----------------------------------------------------------------------------------------
%	PHYSICAL CONSTANTS/OTHER DEFINITIONS
%----------------------------------------------------------------------------------------

\begin{constants}{lr@{${}={}$}l} % The list of physical constants is a three column table

	% The \SI{}{} command is provided by the siunitx package, see its documentation for instructions on how to use it

	Speed of Light & $c_{0}$ & \SI{2.99792458e8}{\meter\per\second} (exact)\\
	%Constant Name & $Symbol$ & $Constant Value$ with units\\

\end{constants}

%----------------------------------------------------------------------------------------
%	SYMBOLS
%----------------------------------------------------------------------------------------

\begin{symbols}{lll} % Include a list of Symbols (a three column table)

	$a$ & distance & \si{\meter} \\
	$P$ & power & \si{\watt} (\si{\joule\per\second}) \\
	%Symbol & Name & Unit \\

	\addlinespace % Gap to separate the Roman symbols from the Greek

	$\omega$ & angular frequency & \si{\radian} \\

\end{symbols}

%----------------------------------------------------------------------------------------
%	DEDICATION
%----------------------------------------------------------------------------------------

\dedicatory{For/Dedicated to/To my\ldots}

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - CHAPTERS
%----------------------------------------------------------------------------------------

\mainmatter % Begin numeric (1,2,3...) page numbering

\pagestyle{thesis} % Return the page headers back to the "thesis" style

% Include the chapters of the thesis as separate files from the Chapters folder
% Uncomment the lines as you write the chapters

\include{Chapters/Chapter1}
%\include{Chapters/Chapter2} 
%\include{Chapters/Chapter3}
%\include{Chapters/Chapter4} 
%\include{Chapters/Chapter5}

%----------------------------------------------------------------------------------------
%	THESIS CONTENT - APPENDICES
%----------------------------------------------------------------------------------------

\appendix % Cue to tell LaTeX that the following "chapters" are Appendices

% Include the appendices of the thesis as separate files from the Appendices folder
% Uncomment the lines as you write the Appendices

\include{Appendices/AppendixA}
%\include{Appendices/AppendixB}
%\include{Appendices/AppendixC}

%----------------------------------------------------------------------------------------
%	BIBLIOGRAPHY
%----------------------------------------------------------------------------------------

\printbibliography[heading=bibintoc]

%----------------------------------------------------------------------------------------

\end{document}

### 2022-01-15 paper/main.tex
Fill out basic information in template
Just to get the basic variables filled in

@@
 
-\thesistitle{Thesis Title} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
-\supervisor{Dr. James \textsc{Smith}} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
+\thesistitle{Building a Mobile Device that Uses the Power of a Desktop Computer} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
+\supervisor{Dr. Robert \textsc{Heinrichs}} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
 \examiner{} % Your examiner's name, this is not currently used anywhere in the template, print it elsewhere with \examname
-\degree{Doctor of Philosophy} % Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
-\author{John \textsc{Smith}} % Your name, this is used in the title page and abstract, print it elsewhere with \authorname
+\degree{Software Engineering} % Your degree name, this is used in the title page and abstract, print it elsewhere with \degreename
+\author{Dylan \textsc{Lathrum}} % Your name, this is used in the title page and abstract, print it elsewhere with \authorname
 \addresses{} % Your address, this is not currently used anywhere in the template, print it elsewhere with \addressname
 
-\subject{Biological Sciences} % Your subject area, this is not currently used anywhere in the template, print it elsewhere with \subjectname
+\subject{Software Engineering} % Your subject area, this is not currently used anywhere in the template, print it elsewhere with \subjectname
 \keywords{} % Keywords for your thesis, this is not currently used anywhere in the template, print it elsewhere with \keywordnames
-\university{\href{http://www.university.com}{University Name}} % Your university's name and URL, this is used in the title page and abstract, print it elsewhere with \univname
-\department{\href{http://department.university.com}{Department or School Name}} % Your department's name and URL, this is used in the title page and abstract, print it elsewhere with \deptname
-\group{\href{http://researchgroup.university.com}{Research Group Name}} % Your research group's name and URL, this is used in the title page, print it elsewhere with \groupname
-\faculty{\href{http://faculty.university.com}{Faculty Name}} % Your faculty's name and URL, this is used in the title page and abstract, print it elsewhere with \facname
+\university{\href{https://www.asu.edu/}{Arizona State University}} % Your university's name and URL, this is used in the title page and abstract, print it elsewhere with \univname
+\department{\href{https://barretthonors.asu.edu/}{Barrett, The Honors College}} % Your department's name and URL, this is used in the title page and abstract, print it elsewhere with \deptname
+% \group{} % Your research group's name and URL, this is used in the title page, print it elsewhere with \groupname
+% \faculty{\href{http://faculty.university.com}{Faculty Name}} % Your faculty's name and URL, this is used in the title page and abstract, print it elsewhere with \facname
 
@@
 		{\scshape\LARGE \univname\par}\vspace{1.5cm} % University name
-		\textsc{\Large Doctoral Thesis}\\[0.5cm] % Thesis type
+		\textsc{\Large Undergraduate Thesis}\\[0.5cm] % Thesis type
 
@@
 				\emph{Author:}\\
-				\href{http://www.johnsmith.com}{\authorname} % Author name - remove the \href bracket to remove the link
+				\href{https://www.linkedin.com/in/dylanlathrum/}{\authorname} % Author name - remove the \href bracket to remove the link
 			\end{flushleft}
@@
 				\emph{Supervisor:} \\
-				\href{http://www.jamessmith.com}{\supname} % Supervisor name - remove the \href bracket to remove the link  
+				\href{https://isearch.asu.edu/profile/3180015}{\supname} % Supervisor name - remove the \href bracket to remove the link  
 			\end{flushright}
@@
 		\large \textit{A thesis submitted in fulfillment of the requirements\\ for the degree of \degreename}\\[0.3cm] % University requirement text
-		\textit{in the}\\[0.4cm]
-		\groupname\\\deptname\\[2cm] % Research group name and department name
+		\textit{for}\\[0.4cm]
+		\deptname\\[2cm] % Research group name and department name
 
@@
 
-\noindent\enquote{\itshape Thanks to my solid academic training, today I can write hundreds of words on virtually any topic without possessing a shred of information, which is how I got a good job in journalism.}\bigbreak
+\noindent\enquote{\itshape Premature optimization is the root of all evil.}\bigbreak
 
-\hfill Dave Barry
+\hfill Sir Tony Hoare
 

### 2022-01-26 paper/Chapters/Evaluation.tex
First draft of outline
This also creates the file structure for each of the chapters, which is nice

\chapter{Evaluation} % Main chapter title

\label{Chapter6} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

\section{Testing Methodology}

\todosection

\section{Responsivity and Latency}

\todosection

\section{Performance}

\todosection

\section{Quality}

\todosection

\section{Summary}

\todosection

### 2022-01-26 paper/Chapters/Example.tex
First draft of outline
This also creates the file structure for each of the chapters, which is nice

% Chapter 1

\chapter{Chapter Title Here} % Main chapter title

\label{Chapter1} % For referencing the chapter elsewhere, use \ref{Chapter1}

%----------------------------------------------------------------------------------------

% Define some commands to keep the formatting separated from the content 
\newcommand{\keyword}[1]{\textbf{#1}}
\newcommand{\tabhead}[1]{\textbf{#1}}
\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\file}[1]{\texttt{\bfseries#1}}
\newcommand{\option}[1]{\texttt{\itshape#1}}

%----------------------------------------------------------------------------------------

\section{Welcome and Thank You}
Welcome to this \LaTeX{} Thesis Template, a beautiful and easy to use template for writing a thesis using the \LaTeX{} typesetting system.

If you are writing a thesis (or will be in the future) and its subject is technical or mathematical (though it doesn't have to be), then creating it in \LaTeX{} is highly recommended as a way to make sure you can just get down to the essential writing without having to worry over formatting or wasting time arguing with your word processor.

\LaTeX{} is easily able to professionally typeset documents that run to hundreds or thousands of pages long. With simple mark-up commands, it automatically sets out the table of contents, margins, page headers and footers and keeps the formatting consistent and beautiful. One of its main strengths is the way it can easily typeset mathematics, even \emph{heavy} mathematics. Even if those equations are the most horribly twisted and most difficult mathematical problems that can only be solved on a super-computer, you can at least count on \LaTeX{} to make them look stunning.

%----------------------------------------------------------------------------------------

\section{Learning \LaTeX{}}

\LaTeX{} is not a \textsc{wysiwyg} (What You See is What You Get) program, unlike word processors such as Microsoft Word or Apple's Pages. Instead, a document written for \LaTeX{} is actually a simple, plain text file that contains \emph{no formatting}. You tell \LaTeX{} how you want the formatting in the finished document by writing in simple commands amongst the text, for example, if I want to use \emph{italic text for emphasis}, I write the \verb|\emph{text}| command and put the text I want in italics in between the curly braces. This means that \LaTeX{} is a \enquote{mark-up} language, very much like HTML.

\subsection{A (not so short) Introduction to \LaTeX{}}

If you are new to \LaTeX{}, there is a very good eBook -- freely available online as a PDF file -- called, \enquote{The Not So Short Introduction to \LaTeX{}}. The book's title is typically shortened to just \emph{lshort}. You can download the latest version (as it is occasionally updated) from here:
\url{http://www.ctan.org/tex-archive/info/lshort/english/lshort.pdf}

It is also available in several other languages. Find yours from the list on this page: \url{http://www.ctan.org/tex-archive/info/lshort/}

It is recommended to take a little time out to learn how to use \LaTeX{} by creating several, small `test' documents, or having a close look at several templates on:\\
\url{http://www.LaTeXTemplates.com}\\
Making the effort now means you're not stuck learning the system when what you \emph{really} need to be doing is writing your thesis.

\subsection{A Short Math Guide for \LaTeX{}}

If you are writing a technical or mathematical thesis, then you may want to read the document by the AMS (American Mathematical Society) called, \enquote{A Short Math Guide for \LaTeX{}}. It can be found online here:
\url{http://www.ams.org/tex/amslatex.html}
under the \enquote{Additional Documentation} section towards the bottom of the page.

\subsection{Common \LaTeX{} Math Symbols}
There are a multitude of mathematical symbols available for \LaTeX{} and it would take a great effort to learn the commands for them all. The most common ones you are likely to use are shown on this page:
\url{http://www.sunilpatel.co.uk/latex-type/latex-math-symbols/}

You can use this page as a reference or crib sheet, the symbols are rendered as large, high quality images so you can quickly find the \LaTeX{} command for the symbol you need.

\subsection{\LaTeX{} on a Mac}

The \LaTeX{} distribution is available for many systems including Windows, Linux and Mac OS X. The package for OS X is called MacTeX and it contains all the applications you need -- bundled together and pre-customized -- for a fully working \LaTeX{} environment and work flow.

MacTeX includes a custom dedicated \LaTeX{} editor called TeXShop for writing your `\file{.tex}' files and BibDesk: a program to manage your references and create your bibliography section just as easily as managing songs and creating playlists in iTunes.

%----------------------------------------------------------------------------------------

\section{Getting Started with this Template}

If you are familiar with \LaTeX{}, then you should explore the directory structure of the template and then proceed to place your own information into the \emph{THESIS INFORMATION} block of the \file{main.tex} file. You can then modify the rest of this file to your unique specifications based on your degree/university. Section \ref{FillingFile} on page \pageref{FillingFile} will help you do this. Make sure you also read section \ref{ThesisConventions} about thesis conventions to get the most out of this template.

If you are new to \LaTeX{} it is recommended that you carry on reading through the rest of the information in this document.

Before you begin using this template you should ensure that its style complies with the thesis style guidelines imposed by your institution. In most cases this template style and layout will be suitable. If it is not, it may only require a small change to bring the template in line with your institution's recommendations. These modifications will need to be done on the \file{MastersDoctoralThesis.cls} file.

\subsection{About this Template}

This \LaTeX{} Thesis Template is originally based and created around a \LaTeX{} style file created by Steve R.\ Gunn from the University of Southampton (UK), department of Electronics and Computer Science. You can find his original thesis style file at his site, here:
\url{http://www.ecs.soton.ac.uk/~srg/softwaretools/document/templates/}

Steve's \file{ecsthesis.cls} was then taken by Sunil Patel who modified it by creating a skeleton framework and folder structure to place the thesis files in. The resulting template can be found on Sunil's site here:
\url{http://www.sunilpatel.co.uk/thesis-template}

Sunil's template was made available through \url{http://www.LaTeXTemplates.com} where it was modified many times based on user requests and questions. Version 2.0 and onwards of this template represents a major modification to Sunil's template and is, in fact, hardly recognisable. The work to make version 2.0 possible was carried out by \href{mailto:vel@latextemplates.com}{Vel} and Johannes Böttcher.

%----------------------------------------------------------------------------------------

\section{What this Template Includes}

\subsection{Folders}

This template comes as a single zip file that expands out to several files and folders. The folder names are mostly self-explanatory:

\keyword{Appendices} -- this is the folder where you put the appendices. Each appendix should go into its own separate \file{.tex} file. An example and template are included in the directory.

\keyword{Chapters} -- this is the folder where you put the thesis chapters. A thesis usually has about six chapters, though there is no hard rule on this. Each chapter should go in its own separate \file{.tex} file and they can be split as:
\begin{itemize}
  \item Chapter 1: Introduction to the thesis topic
  \item Chapter 2: Background information and theory
  \item Chapter 3: (Laboratory) experimental setup
  \item Chapter 4: Details of experiment 1
  \item Chapter 5: Details of experiment 2
  \item Chapter 6: Discussion of the experimental results
  \item Chapter 7: Conclusion and future directions
\end{itemize}
This chapter layout is specialised for the experimental sciences, your discipline may be different.

\keyword{Figures} -- this folder contains all figures for the thesis. These are the final images that will go into the thesis document.

\subsection{Files}

Included are also several files, most of them are plain text and you can see their contents in a text editor. After initial compilation, you will see that more auxiliary files are created by \LaTeX{} or BibTeX and which you don't need to delete or worry about:

\keyword{example.bib} -- this is an important file that contains all the bibliographic information and references that you will be citing in the thesis for use with BibTeX. You can write it manually, but there are reference manager programs available that will create and manage it for you. Bibliographies in \LaTeX{} are a large subject and you may need to read about BibTeX before starting with this. Many modern reference managers will allow you to export your references in BibTeX format which greatly eases the amount of work you have to do.

\keyword{MastersDoctoralThesis.cls} -- this is an important file. It is the class file that tells \LaTeX{} how to format the thesis.

\keyword{main.pdf} -- this is your beautifully typeset thesis (in the PDF file format) created by \LaTeX{}. It is supplied in the PDF with the template and after you compile the template you should get an identical version.

\keyword{main.tex} -- this is an important file. This is the file that you tell \LaTeX{} to compile to produce your thesis as a PDF file. It contains the framework and constructs that tell \LaTeX{} how to layout the thesis. It is heavily commented so you can read exactly what each line of code does and why it is there. After you put your own information into the \emph{THESIS INFORMATION} block -- you have now started your thesis!

Files that are \emph{not} included, but are created by \LaTeX{} as auxiliary files include:

\keyword{main.aux} -- this is an auxiliary file generated by \LaTeX{}, if it is deleted \LaTeX{} simply regenerates it when you run the main \file{.tex} file.

\keyword{main.bbl} -- this is an auxiliary file generated by BibTeX, if it is deleted, BibTeX simply regenerates it when you run the \file{main.aux} file. Whereas the \file{.bib} file contains all the references you have, this \file{.bbl} file contains the references you have actually cited in the thesis and is used to build the bibliography section of the thesis.

\keyword{main.blg} -- this is an auxiliary file generated by BibTeX, if it is deleted BibTeX simply regenerates it when you run the main \file{.aux} file.

\keyword{main.lof} -- this is an auxiliary file generated by \LaTeX{}, if it is deleted \LaTeX{} simply regenerates it when you run the main \file{.tex} file. It tells \LaTeX{} how to build the \emph{List of Figures} section.

\keyword{main.log} -- this is an auxiliary file generated by \LaTeX{}, if it is deleted \LaTeX{} simply regenerates it when you run the main \file{.tex} file. It contains messages from \LaTeX{}, if you receive errors and warnings from \LaTeX{}, they will be in this \file{.log} file.

\keyword{main.lot} -- this is an auxiliary file generated by \LaTeX{}, if it is deleted \LaTeX{} simply regenerates it when you run the main \file{.tex} file. It tells \LaTeX{} how to build the \emph{List of Tables} section.

\keyword{main.out} -- this is an auxiliary file generated by \LaTeX{}, if it is deleted \LaTeX{} simply regenerates it when you run the main \file{.tex} file.

So from this long list, only the files with the \file{.bib}, \file{.cls} and \file{.tex} extensions are the most important ones. The other auxiliary files can be ignored or deleted as \LaTeX{} and BibTeX will regenerate them.

%----------------------------------------------------------------------------------------

\section{Filling in Your Information in the \file{main.tex} File}\label{FillingFile}

You will need to personalise the thesis template and make it your own by filling in your own information. This is done by editing the \file{main.tex} file in a text editor or your favourite LaTeX environment.

Open the file and scroll down to the third large block titled \emph{THESIS INFORMATION} where you can see the entries for \emph{University Name}, \emph{Department Name}, etc \ldots

Fill out the information about yourself, your group and institution. You can also insert web links, if you do, make sure you use the full URL, including the \code{http://} for this. If you don't want these to be linked, simply remove the \verb|\href{url}{name}| and only leave the name.

When you have done this, save the file and recompile \code{main.tex}. All the information you filled in should now be in the PDF, complete with web links. You can now begin your thesis proper!

%----------------------------------------------------------------------------------------

\section{The \code{main.tex} File Explained}

The \file{main.tex} file contains the structure of the thesis. There are plenty of written comments that explain what pages, sections and formatting the \LaTeX{} code is creating. Each major document element is divided into commented blocks with titles in all capitals to make it obvious what the following bit of code is doing. Initially there seems to be a lot of \LaTeX{} code, but this is all formatting, and it has all been taken care of so you don't have to do it.

Begin by checking that your information on the title page is correct. For the thesis declaration, your institution may insist on something different than the text given. If this is the case, just replace what you see with what is required in the \emph{DECLARATION PAGE} block.

Then comes a page which contains a funny quote. You can put your own, or quote your favourite scientist, author, person, and so on. Make sure to put the name of the person who you took the quote from.

Following this is the abstract page which summarises your work in a condensed way and can almost be used as a standalone document to describe what you have done. The text you write will cause the heading to move up so don't worry about running out of space.

Next come the acknowledgements. On this page, write about all the people who you wish to thank (not forgetting parents, partners and your advisor/supervisor).

The contents pages, list of figures and tables are all taken care of for you and do not need to be manually created or edited. The next set of pages are more likely to be optional and can be deleted since they are for a more technical thesis: insert a list of abbreviations you have used in the thesis, then a list of the physical constants and numbers you refer to and finally, a list of mathematical symbols used in any formulae. Making the effort to fill these tables means the reader has a one-stop place to refer to instead of searching the internet and references to try and find out what you meant by certain abbreviations or symbols.

The list of symbols is split into the Roman and Greek alphabets. Whereas the abbreviations and symbols ought to be listed in alphabetical order (and this is \emph{not} done automatically for you) the list of physical constants should be grouped into similar themes.

The next page contains a one line dedication. Who will you dedicate your thesis to?

Finally, there is the block where the chapters are included. Uncomment the lines (delete the \code{\%} character) as you write the chapters. Each chapter should be written in its own file and put into the \emph{Chapters} folder and named \file{Chapter1}, \file{Chapter2}, etc\ldots Similarly for the appendices, uncomment the lines as you need them. Each appendix should go into its own file and placed in the \emph{Appendices} folder.

After the preamble, chapters and appendices finally comes the bibliography. The bibliography style (called \option{authoryear}) is used for the bibliography and is a fully featured style that will even include links to where the referenced paper can be found online. Do not underestimate how grateful your reader will be to find that a reference to a paper is just a click away. Of course, this relies on you putting the URL information into the BibTeX file in the first place.

%----------------------------------------------------------------------------------------

\section{Thesis Features and Conventions}\label{ThesisConventions}

To get the best out of this template, there are a few conventions that you may want to follow.

One of the most important (and most difficult) things to keep track of in such a long document as a thesis is consistency. Using certain conventions and ways of doing things (such as using a Todo list) makes the job easier. Of course, all of these are optional and you can adopt your own method.

\subsection{Printing Format}

This thesis template is designed for double sided printing (i.e. content on the front and back of pages) as most theses are printed and bound this way. Switching to one sided printing is as simple as uncommenting the \option{oneside} option of the \code{documentclass} command at the top of the \file{main.tex} file. You may then wish to adjust the margins to suit specifications from your institution.

The headers for the pages contain the page number on the outer side (so it is easy to flick through to the page you want) and the chapter name on the inner side.

The text is set to 11 point by default with single line spacing, again, you can tune the text size and spacing should you want or need to using the options at the very start of \file{main.tex}. The spacing can be changed similarly by replacing the \option{singlespacing} with \option{onehalfspacing} or \option{doublespacing}.

\subsection{Using US Letter Paper}

The paper size used in the template is A4, which is the standard size in Europe. If you are using this thesis template elsewhere and particularly in the United States, then you may have to change the A4 paper size to the US Letter size. This can be done in the margins settings section in \file{main.tex}.

Due to the differences in the paper size, the resulting margins may be different to what you like or require (as it is common for institutions to dictate certain margin sizes). If this is the case, then the margin sizes can be tweaked by modifying the values in the same block as where you set the paper size. Now your document should be set up for US Letter paper size with suitable margins.

\subsection{References}

The \code{biblatex} package is used to format the bibliography and inserts references such as this one \parencite{Reference1}. The options used in the \file{main.tex} file mean that the in-text citations of references are formatted with the author(s) listed with the date of the publication. Multiple references are separated by semicolons (e.g. \parencite{Reference2, Reference1}) and references with more than three authors only show the first author with \emph{et al.} indicating there are more authors (e.g. \parencite{Reference3}). This is done automatically for you. To see how you use references, have a look at the \file{Chapter1.tex} source file. Many reference managers allow you to simply drag the reference into the document as you type.

Scientific references should come \emph{before} the punctuation mark if there is one (such as a comma or period). The same goes for footnotes\footnote{Such as this footnote, here down at the bottom of the page.}. You can change this but the most important thing is to keep the convention consistent throughout the thesis. Footnotes themselves should be full, descriptive sentences (beginning with a capital letter and ending with a full stop). The APA6 states: \enquote{Footnote numbers should be superscripted, [...], following any punctuation mark except a dash.} The Chicago manual of style states: \enquote{A note number should be placed at the end of a sentence or clause. The number follows any punctuation mark except the dash, which it precedes. It follows a closing parenthesis.}

The bibliography is typeset with references listed in alphabetical order by the first author's last name. This is similar to the APA referencing style. To see how \LaTeX{} typesets the bibliography, have a look at the very end of this document (or just click on the reference number links in in-text citations).

\subsubsection{A Note on bibtex}

The bibtex backend used in the template by default does not correctly handle unicode character encoding (i.e. "international" characters). You may see a warning about this in the compilation log and, if your references contain unicode characters, they may not show up correctly or at all. The solution to this is to use the biber backend instead of the outdated bibtex backend. This is done by finding this in \file{main.tex}: \option{backend=bibtex} and changing it to \option{backend=biber}. You will then need to delete all auxiliary BibTeX files and navigate to the template directory in your terminal (command prompt). Once there, simply type \code{biber main} and biber will compile your bibliography. You can then compile \file{main.tex} as normal and your bibliography will be updated. An alternative is to set up your LaTeX editor to compile with biber instead of bibtex, see \href{http://tex.stackexchange.com/questions/154751/biblatex-with-biber-configuring-my-editor-to-avoid-undefined-citations/}{here} for how to do this for various editors.

\subsection{Tables}

Tables are an important way of displaying your results, below is an example table which was generated with this code:

{\small
\begin{verbatim}
\begin{table}
\caption{The effects of treatments X and Y on the four groups studied.}
\label{tab:treatments}
\centering
\begin{tabular}{l l l}
\toprule
\tabhead{Groups} & \tabhead{Treatment X} & \tabhead{Treatment Y} \\
\midrule
1 & 0.2 & 0.8\\
2 & 0.17 & 0.7\\
3 & 0.24 & 0.75\\
4 & 0.68 & 0.3\\
\bottomrule\\
\end{tabular}
\end{table}
\end{verbatim}
}

\begin{table}
  \caption{The effects of treatments X and Y on the four groups studied.}
  \label{tab:treatments}
  \centering
  \begin{tabular}{l l l}
    \toprule
    \tabhead{Groups} & \tabhead{Treatment X} & \tabhead{Treatment Y} \\
    \midrule
    1                & 0.2                   & 0.8                   \\
    2                & 0.17                  & 0.7                   \\
    3                & 0.24                  & 0.75                  \\
    4                & 0.68                  & 0.3                   \\
    \bottomrule                                                      \\
  \end{tabular}
\end{table}

You can reference tables with \verb|\ref{<label>}| where the label is defined within the table environment. See \file{Chapter1.tex} for an example of the label and citation (e.g. Table~\ref{tab:treatments}).

\subsection{Figures}

There will hopefully be many figures in your thesis (that should be placed in the \emph{Figures} folder). The way to insert figures into your thesis is to use a code template like this:
\begin{verbatim}
\begin{figure}
\centering
\includegraphics{Figures/Electron}
\decoRule
\caption[An Electron]{An electron (artist's impression).}
\label{fig:Electron}
\end{figure}
\end{verbatim}
Also look in the source file. Putting this code into the source file produces the picture of the electron that you can see in the figure below.

\begin{figure}[th]
  \centering
  \includegraphics{Figures/Electron}
  \decoRule
  \caption[An Electron]{An electron (artist's impression).}
  \label{fig:Electron}
\end{figure}

Sometimes figures don't always appear where you write them in the source. The placement depends on how much space there is on the page for the figure. Sometimes there is not enough room to fit a figure directly where it should go (in relation to the text) and so \LaTeX{} puts it at the top of the next page. Positioning figures is the job of \LaTeX{} and so you should only worry about making them look good!

Figures usually should have captions just in case you need to refer to them (such as in Figure~\ref{fig:Electron}). The \verb|\caption| command contains two parts, the first part, inside the square brackets is the title that will appear in the \emph{List of Figures}, and so should be short. The second part in the curly brackets should contain the longer and more descriptive caption text.

The \verb|\decoRule| command is optional and simply puts an aesthetic horizontal line below the image. If you do this for one image, do it for all of them.

\LaTeX{} is capable of using images in pdf, jpg and png format.

\subsection{Typesetting mathematics}

If your thesis is going to contain heavy mathematical content, be sure that \LaTeX{} will make it look beautiful, even though it won't be able to solve the equations for you.

The \enquote{Not So Short Introduction to \LaTeX} (available on \href{http://www.ctan.org/tex-archive/info/lshort/english/lshort.pdf}{CTAN}) should tell you everything you need to know for most cases of typesetting mathematics. If you need more information, a much more thorough mathematical guide is available from the AMS called, \enquote{A Short Math Guide to \LaTeX} and can be downloaded from:
\url{ftp://ftp.ams.org/pub/tex/doc/amsmath/short-math-guide.pdf}

There are many different \LaTeX{} symbols to remember, luckily you can find the most common symbols in \href{http://ctan.org/pkg/comprehensive}{The Comprehensive \LaTeX~Symbol List}.

You can write an equation, which is automatically given an equation number by \LaTeX{} like this:
\begin{verbatim}
\begin{equation}
E = mc^{2}
\label{eqn:Einstein}
\end{equation}
\end{verbatim}

This will produce Einstein's famous energy-matter equivalence equation:
\begin{equation}
  E = mc^{2}
  \label{eqn:Einstein}
\end{equation}

All equations you write (which are not in the middle of paragraph text) are automatically given equation numbers by \LaTeX{}. If you don't want a particular equation numbered, use the unnumbered form:
\begin{verbatim}
\[ a^{2}=4 \]
\end{verbatim}

%----------------------------------------------------------------------------------------

\section{Sectioning and Subsectioning}

You should break your thesis up into nice, bite-sized sections and subsections. \LaTeX{} automatically builds a table of Contents by looking at all the \verb|\chapter{}|, \verb|\section{}|  and \verb|\subsection{}| commands you write in the source.

The Table of Contents should only list the sections to three (3) levels. A \verb|chapter{}| is level zero (0). A \verb|\section{}| is level one (1) and so a \verb|\subsection{}| is level two (2). In your thesis it is likely that you will even use a \verb|subsubsection{}|, which is level three (3). The depth to which the Table of Contents is formatted is set within \file{MastersDoctoralThesis.cls}. If you need this changed, you can do it in \file{main.tex}.

%----------------------------------------------------------------------------------------

\section{In Closing}

You have reached the end of this mini-guide. You can now rename or overwrite this pdf file and begin writing your own \file{Chapter1.tex} and the rest of your thesis. The easy work of setting up the structure and framework has been taken care of for you. It's now your job to fill it out!

Good luck and have lots of fun!

\begin{flushright}
  Guide written by ---\\
  Sunil Patel: \href{http://www.sunilpatel.co.uk}{www.sunilpatel.co.uk}\\
  Vel: \href{http://www.LaTeXTemplates.com}{LaTeXTemplates.com}
\end{flushright}

### 2022-01-26 paper/main.tex
First draft of outline
This also creates the file structure for each of the chapters, which is nice

@@
 singlespacing, % Single line spacing, alternatives: onehalfspacing or doublespacing
-%draft, % Uncomment to enable draft mode (no pictures, no links, overfull hboxes indicated)
+% draft, % Uncomment to enable draft mode (no pictures, no links, overfull hboxes indicated)
+% final, % Uncomment to enable final mode (hide todos)
 %nolistspacing, % If the document is onehalfspacing or doublespacing, uncomment this to set spacing in lists to single
@@
 \usepackage[autostyle=true]{csquotes} % Required to generate language-dependent quotes in the bibliography
+
+\usepackage[colorinlistoftodos,prependcaption,textsize=tiny,obeyFinal]{todonotes} % Todo notes in margin
+
+\newcommand{\todosection}{\todo[linecolor=olive,backgroundcolor=olive!25,bordercolor=olive,inline]{TODO}}
 
@@
 
-	\textbf{LAH} & \textbf{L}ist \textbf{A}bbreviations \textbf{H}ere\\
-	\textbf{WSF} & \textbf{W}hat (it) \textbf{S}tands \textbf{F}or\\
+	\textbf{CRD} & \textbf{C}hrome \textbf{R}emote \textbf{D}esktop\\
+	\textbf{PCB} & \textbf{P}rinted \textbf{C}ircuit \textbf{B}oard\\
+	\textbf{RDP} & \textbf{R}emote \textbf{D}esktop \textbf{P}rotocol\\
+	\textbf{SSH} & \textbf{S}ecure \textbf{Sh}ell (Protocol)\\
+	\textbf{VNC} & \textbf{V}irtual \textbf{N}etwork \textbf{C}omputing\\
 
@@
 
-\begin{constants}{lr@{${}={}$}l} % The list of physical constants is a three column table
+% \begin{constants}{lr@{${}={}$}l} % The list of physical constants is a three column table
 
-	% The \SI{}{} command is provided by the siunitx package, see its documentation for instructions on how to use it
+% 	% The \SI{}{} command is provided by the siunitx package, see its documentation for instructions on how to use it
 
-	Speed of Light & $c_{0}$ & \SI{2.99792458e8}{\meter\per\second} (exact)\\
-	%Constant Name & $Symbol$ & $Constant Value$ with units\\
+% 	Speed of Light & $c_{0}$ & \SI{2.99792458e8}{\meter\per\second} (exact)\\
+% 	%Constant Name & $Symbol$ & $Constant Value$ with units\\
 
-\end{constants}
+% \end{constants}
 
@@
 
-\begin{symbols}{lll} % Include a list of Symbols (a three column table)
+% \begin{symbols}{lll} % Include a list of Symbols (a three column table)
 
-	$a$ & distance & \si{\meter} \\
-	$P$ & power & \si{\watt} (\si{\joule\per\second}) \\
-	%Symbol & Name & Unit \\
+% 	$a$ & distance & \si{\meter} \\
+% 	$P$ & power & \si{\watt} (\si{\joule\per\second}) \\
+% 	%Symbol & Name & Unit \\
 
-	\addlinespace % Gap to separate the Roman symbols from the Greek
+% 	\addlinespace % Gap to separate the Roman symbols from the Greek
 
-	$\omega$ & angular frequency & \si{\radian} \\
+% 	$\omega$ & angular frequency & \si{\radian} \\
 
-\end{symbols}
+% \end{symbols}
 
@@
 
-\include{Chapters/Chapter1}
-%\include{Chapters/Chapter2} 
-%\include{Chapters/Chapter3}
-%\include{Chapters/Chapter4} 
-%\include{Chapters/Chapter5}
+\include{Chapters/Introduction}
+\include{Chapters/Background}
+\include{Chapters/StateOfTheArt}
+\include{Chapters/Hardware}
+\include{Chapters/Software}
+\include{Chapters/Evaluation}
+\include{Chapters/Conclusion}
 

### 2022-01-31 paper/Chapters/Background.tex

\chapter{Background} % Main chapter title

\label{Chapter2} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

\section{Specialization of Computers}

\todosection

\subsection{Power of the Desktop}

\todosection

\subsection{Convenience of the Laptop}

\todosection

\subsection{Rise of the Gaming Laptop}

\todosection

\section{Thin Clients}

\todosection

\section{Application to Modern Day}

\todosection

### 2022-01-31 paper/Chapters/Conclusion.tex

\chapter{Conclusion} % Main chapter title

\label{Chapter7} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

\section{Summary}

\todosection

\section{Limitations}

\todosection

\section{Future Research}

\todosection

### 2022-01-31 paper/Chapters/Hardware.tex

\chapter{Developing the Hardware} % Main chapter title

\label{Chapter4} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

\section{Requirements}

\todosection

\subsection{Performance}

\todosection

\subsection{Cost}

\todosection

\section{Choosing Parts}

\todosection

\section{Prototyping}

\todosection

\section{Designing the Printed Circuit Board}

\todosection

\section{Manufacturing the Circuit Board}

\todosection

### 2022-01-31 paper/Chapters/Introduction.tex

\chapter{Introduction} % Main chapter title

\label{Chapter1} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

\section{Power and Portability}

\todosection

\section{Purpose of this Thesis}

\todosection

\section{Thesis Overview}

\todosection

### 2022-01-31 paper/Chapters/Software.tex

\chapter{Developing the Software} % Main chapter title

\label{Chapter5} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

\section{Requirements}

\todosection

\section{Potential Avenues}

\todosection

\subsection{NVIDIA GameStream}

\todosection

\subsection{Moonlight}

\todosection

\section{Developing for ARM}

\todosection

### 2022-01-31 paper/Chapters/StateOfTheArt.tex

\chapter{State of the Art} % Main chapter title

\label{Chapter3} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

\section{Introduction}

\todosection

\section{Hardware Solutions}

\todosection

\subsection{Thin Clients}

\todosection

\subsection{Web Clients}

\todosection

\section{Software Solutions}

\todosection

\subsection{Remote Desktop Protocol}

\todosection

\subsection{Virtual Network Computing}

\todosection

\subsection{Chrome Remote Desktop}

\todosection

\subsection{Secure Shell Protocol}

\todosection

### 2022-02-01 paper/Chapters/Background.tex
Begin working on the Background section
The first words are always the hardest

@@
 \section{Specialization of Computers}
+\todo{Should this paragraph be moved above?}
+In the age of monolithic computers the size of entire rooms, it was expected of a computer to only be capable of serving a single function.
+Before the technology could be miniaturized, every single computer had to be specialized for the job at hand.
+As time passed and technology progressed, a single computer could be produced that served multiple functions as well as be adaptable programmable to handle new tasks that were not considered when the machine was originally built.
+This gave rise to what was known as \enquote{Autonomic Computing}, a system of characteristics that are built into a computer to help it self-manage it's resources and adapt to it's administrator's requirements without the need to be redesigned for it's new purpose \todo{Correct citation format?}\citep{AutonomicComputing}.
+Autonomic Computing was designed to combat the exponential complexity crisis that came from the widespread availability of computers in different disciplines, since now anyone who could afford a computer could program it for any purpose.
+As new use cases were found for computers in day-to-day life and new technologies were being developed to interface with the world around us, the need for componentization became ever more apparent.
+Researchers at IBM knew that the best way to address the looming problem of runaway complexity in computers was to develop a way for the computers to automatically interface with any new components that are installed, and to configure itself to only use what is necessary for the job at hand \citep[p.~43]{AutonomicComputing}.
+This allowed hardware and software developers to focus on building their products to work with a common standard rather than having to manually integrate their products with every computer.
 
-\todosection
+Once the idea of modularity began to take hold in the computer industry, attention was turned once again towards specializing individual computers.
+Now that a computer's physical footprint can be minified, and peripheral components can be added and removed without a complete reengineering of the device, a single computer can be optimized to handle a single task without the overhead cost of building a monolithic computer \citep{Burbeck2007ComplexityAT}.
+Now a computer can be specialized to serve a particular purpose or set of purposes while keeping development time comparatively short.
+The world has seen this realized through numerous applications such as cell phones, designed to be user-friendly portable devices, computing clusters, built for high-performance computing, or even internet routers, which are built to be a plug-and-play solution for a problem posed to users of all levels of familiarity with computers.
+Without this idea of Autonomic Computing, each of these devices would have to be reengineered from the ground up every time a new use case was developed.
+
+This leads into the specialization of home computers in from the perspective of a consumer.
+It used to be that a \enquote{home computer} was a device that was bought off the shelf as-is and served it's purpose.
+Now a home computer can be a desktop PC that can be upgraded with time and seldom moves from it's place, or it could be a laptop that is portable and easy to cary around.
+These specializations bring more choices for the consumer to pick a computer that best suits their needs, but they often come with compromises.
 
@@
 
-\todosection
+The classical manifestation of a personal computer is the desktop computer.
+Known for it's componentization, user repairability, and direct connections to power and the network, desktop computers are the best option for a user looking for a workhorse system.
+Even though they aren't very portable, desktop computers allow consumers to access greater processor power for a lower cost compared to portable devices \citep{Meyer20145RS}.
+Due to this, a desktop PC is often seen at one's home or place of work hardwired to the wall where it remains unmoving unless it needs upgrading, cleaning, or repairs.
+Adding on the fact the desktop is modular, it can be easily customized to fit any user's needs, as well as enable particular parts to be replaced or upgraded without needing to rebuild the entire system.
+
+While desktops were the staple of personal computers for decades, advancing technology and the rising need of portability and flexibility has led to the growth of laptops.
 
 \subsection{Convenience of the Laptop}
+
+As the world moved towards a more mobile lifestyle, the laptop shifted from being a luxury to being a necessity.
 

### 2022-02-01 paper/main.tex
Begin working on the Background section
The first words are always the hardest

@@
 singlespacing, % Single line spacing, alternatives: onehalfspacing or doublespacing
-% draft, % Uncomment to enable draft mode (no pictures, no links, overfull hboxes indicated)
+draft, % Uncomment to enable draft mode (no pictures, no links, overfull hboxes indicated)
 % final, % Uncomment to enable final mode (hide todos)
@@
 
-\addbibresource{example.bib} % The filename of the bibliography
+\addbibresource{bibliography.bib} % The filename of the bibliography
 
@@
 
+	\textbf{CPU} & \textbf{C}entral \textbf{P}rocessing \textbf{U}nit\\
 	\textbf{CRD} & \textbf{C}hrome \textbf{R}emote \textbf{D}esktop\\
+	\textbf{GPU} & \textbf{G}raphics \textbf{P}rocessing \textbf{U}nit\\
+	\textbf{PC} & \textbf{P}ersonal \textbf{C}computer\\
 	\textbf{PCB} & \textbf{P}rinted \textbf{C}ircuit \textbf{B}oard\\
@@
 
+% \include{Chapters/Example}
 \include{Chapters/Introduction}

### 2022-02-08 paper/Chapters/Background.tex
Draft of background section

@@
 
-\section{Specialization of Computers}
-\todo{Should this paragraph be moved above?}
+Computers have a long history of being built to serve a particular purpose, often compromising some aspects of the system to bolster others.
+Chapter \ref{SpecializationOfComputers} discusses the history of how computers were built for a particular function, and how the advancement of technology has brought the \enquote{Personal Computer} seen know today.
+\todo{Too much detail?}
+Chapter \ref{PowerOfTheDesktop} discusses the benefits of a desktop computer, namely the raw power and serviceability.
+Chapter \ref{ConvenienceOfTheLaptop} discusses the benefits and drawbacks of a laptop computer, namely the portability and convenience at the cost of power.
+Chapter \ref{RiseOfTheGamingLaptop} discusses the rise of the gaming laptop, and the closing gap between power and portability at the cost of increased price.
+Chapter \ref{ThinClients} briefly discusses thin clients, a corporate solution for remote access to a computer.
+Chapter \ref{ApplicationToModernDay} introduces current world wide context that has an impact on the computing world and the need for power and portability.
+
+\section{Specialization of Computers}\label{SpecializationOfComputers}
+
 In the age of monolithic computers the size of entire rooms, it was expected of a computer to only be capable of serving a single function.
@@
 As time passed and technology progressed, a single computer could be produced that served multiple functions as well as be adaptable programmable to handle new tasks that were not considered when the machine was originally built.
-This gave rise to what was known as \enquote{Autonomic Computing}, a system of characteristics that are built into a computer to help it self-manage it's resources and adapt to it's administrator's requirements without the need to be redesigned for it's new purpose \todo{Correct citation format?}\citep{AutonomicComputing}.
+This gave rise to what was known as \enquote{Autonomic Computing}, a system of characteristics that are built into a computer to help it self-manage it's resources and adapt to it's administrator's requirements without the need to be redesigned for it's new purpose \citep{AutonomicComputing}.
 Autonomic Computing was designed to combat the exponential complexity crisis that came from the widespread availability of computers in different disciplines, since now anyone who could afford a computer could program it for any purpose.
@@
 
-\subsection{Power of the Desktop}
+\subsection{Power of the Desktop}\label{PowerOfTheDesktop}
 
@@
 
-\subsection{Convenience of the Laptop}
+\subsection{Convenience of the Laptop}\label{ConvenienceOfTheLaptop}
 
 As the world moved towards a more mobile lifestyle, the laptop shifted from being a luxury to being a necessity.
+People began to need computing power on the go, whether it be for working remotely, attending classes and following along, or simply needing more power than their phone can offer.
+Companies began to take note too and started building applications to work fully within a web browser, such as Google's implementation of Google Docs, allowing users to edit documents on the go without having to install software like Microsoft Word directly to their computers.
+This shift towards mobility provided the general public the ability to be more flexible with how they used computers, no longer requiring them to dedicating a spot in their house to be taken up by a desktop computer.
 
-\todosection
+Reducing the barrier to entry for computers, such as the knowledge to buy the computer the best fits their needs and the restrictions that come with setting up and keeping a desktop, led to the widespread adoption of laptops.
+In fact, by 2021, laptops accounted for 80\% of total computer sales \cite{idc_2021}.
+For most consumers, the laptop is all they need; it may not have the raw power of a desktop, but it handles everything they need it to do.
+The biggest limiting factor then becomes technology, with manufacturers constantly trying to minimize existing desktop technology into the footprint of a laptop while keeping energy efficiency high enough to be feasible in a mobile device.
 
-\subsection{Rise of the Gaming Laptop}
+\subsection{Rise of the Gaming Laptop}\label{RiseOfTheGamingLaptop}
 
-\todosection
+All computer products are a balance, it simply isn't possible to have the perfect combination of portable, powerful, serviceable, and cheap.
+While desktops focus on power and serviceability, laptops often focus portability at the expense of other traits\todo{Theres a word at the tip of my tounge thats a lot better here dangit}.
+As mobility becomes ever more important, gaming laptops have become more popular to bring power to the portable form factor.
+This understandably comes with a cost, with the literal cost of a gaming laptop being the most obvious compromise.
+It is more expensive, and often less user-serviceable, to buy a gaming laptop that has the same performance of a desktop computer -- but given that it is an all-in-one device, it's portability gives it a niche that a desktop cannot hope to rival.
+\todo{Things could be said about the current chip shortage leading to laptops being easier to buy compared to desktops, but that might fit better in the section below.}
+The compromise presented by a gaming laptop helps close the gap between power and portability, but it still falls short of being an all around good solution.
 
-\section{Thin Clients}
+\section{Thin Clients}\label{ThinClients}
 
-\todosection
+On the corporate side of computing, the solution for portability has been the thin client.
+Rather than integrating the power of a desktop to a portable machine, a thin client will connect to a powerful server and provide a thin interface to the user.
+This allows the user to utilize the power of the host machine from a cheap computer that usually doesn't have enough power to be used as a standalone computer.
+While these clients work well for simple interfacing tasks such as data entry or system management, they are not designed to be used for complex tasks that require quick response times or graphical fidelity.
+\todo{Is more information needed here?}
 
-\section{Application to Modern Day}
+\section{Application to Modern Day}\label{ApplicationToModernDay}
 
-\todosection
+In 2022, the world is currently going through some chaotic times that demand flexibility and adaptability like it hasn't faced before.
+From the COVID-19 pandemic to massive chip shortages and supply line disruptions, people are now more than ever looking for a way to be flexible in their work and personal lives now and in the future.
+The COVID-19 pandemic has forced people to be able to operate remotely, whether that be for work, school, or communicating with friends and family.
+\todo{It's a little hard to cite this since they are current events. Is that ok?}
+Many office workers used to working on a desktop computer onsite are now needed to work from home, or classes usually taught in person are now held remotely.
+This shift to remote operation has fueled the need for mobile computers, many of which need to be powerful enough to handle the workload of a desktop computer from anywhere in the world.
+The global chip shortage is also causing issues for the purchase of new devices.
+Parts which are usually readily available for purchase are out of stock with months of delay on back order.
+Resellers are taking advantage of the situation by hiking prices to two or three times the original listing price, and it's become harder than ever to build a new PC.
+
+It doesn't make sense to build or purchase a powerful PC for use at home and another power laptop to handle working on the go, especially with rising prices, so many people are moving to exclusively powerful laptops or looking for other solutions.

### 2022-02-08 paper/main.tex
Draft of background section

--- 2022-01-26 paper/main.tex
@@
 
-\usepackage[backend=bibtex,style=authoryear,natbib=true]{biblatex} % Use the bibtex backend with the authoryear citation style (which resembles APA)
+\usepackage[backend=bibtex,natbib=true]{biblatex} % Use the bibtex backend
 
-\addbibresource{example.bib} % The filename of the bibliography
+\addbibresource{bibliography.bib} % The filename of the bibliography
 

### 2022-02-17 paper/Chapters/Background.tex
Feedback from Advisor
Mostly just notes, they will be fixed next

@@
 \todo{Too much detail?}
+\todoadvisor{Yes, don't show content}
 Chapter \ref{PowerOfTheDesktop} discusses the benefits of a desktop computer, namely the raw power and serviceability.
@@
 
-In the age of monolithic computers the size of entire rooms, it was expected of a computer to only be capable of serving a single function.
+In the age of \todoadvisor{Better describe monolithic computers, maybe add the year}monolithic computers the size of entire rooms, it was expected of a computer to only be capable of serving a single function.
 Before the technology could be miniaturized, every single computer had to be specialized for the job at hand.
@@
 The classical manifestation of a personal computer is the desktop computer.
-Known for it's componentization, user repairability, and direct connections to power and the network, desktop computers are the best option for a user looking for a workhorse system.
+Known for it's \todoadvisor{Give an example}componentization, user repairability, and direct connections to power and the network, desktop computers are the best option for a user looking for a workhorse system.
 Even though they aren't very portable, desktop computers allow consumers to access greater processor power for a lower cost compared to portable devices \citep{Meyer20145RS}.
@@
 This understandably comes with a cost, with the literal cost of a gaming laptop being the most obvious compromise.
-It is more expensive, and often less user-serviceable, to buy a gaming laptop that has the same performance of a desktop computer -- but given that it is an all-in-one device, it's portability gives it a niche that a desktop cannot hope to rival.
+It is more expensive, and often less \todoadvisor{Give an example}user-serviceable, to buy a gaming laptop that has the same performance of a desktop computer -- but given that it is an all-in-one device, it's portability gives it a niche that a desktop cannot hope to rival.
 \todo{Things could be said about the current chip shortage leading to laptops being easier to buy compared to desktops, but that might fit better in the section below.}
@@
 On the corporate side of computing, the solution for portability has been the thin client.
-Rather than integrating the power of a desktop to a portable machine, a thin client will connect to a powerful server and provide a thin interface to the user.
+Rather than integrating the power of a desktop to a portable machine, a thin client will connect to a powerful server and provide a \todoadvisor{Define thin interface}thin interface to the user.
 This allows the user to utilize the power of the host machine from a cheap computer that usually doesn't have enough power to be used as a standalone computer.
@@
 \todo{Is more information needed here?}
+\todoadvisor{Pictures?}
 
@@
 From the COVID-19 pandemic to massive chip shortages and supply line disruptions, people are now more than ever looking for a way to be flexible in their work and personal lives now and in the future.
-The COVID-19 pandemic has forced people to be able to operate remotely, whether that be for work, school, or communicating with friends and family.
+The COVID-19 pandemic has forced people to be able to \todoadvisor{cite (even ASU's announcements)}operate remotely, whether that be for work, school, or communicating with friends and family.
 \todo{It's a little hard to cite this since they are current events. Is that ok?}
-Many office workers used to working on a desktop computer onsite are now needed to work from home, or classes usually taught in person are now held remotely.
+\todoadvisor{News sources (and less academic) are fine}
+Many office workers used to working on a desktop computer onsite are now needed to \todoadvisor{cite}work from home, or classes usually taught in person are now held remotely.
 This shift to remote operation has fueled the need for mobile computers, many of which need to be powerful enough to handle the workload of a desktop computer from anywhere in the world.
-The global chip shortage is also causing issues for the purchase of new devices.
+The \todoadvisor{cite}global chip shortage is also causing issues for the purchase of new devices.
 Parts which are usually readily available for purchase are out of stock with months of delay on back order.
-Resellers are taking advantage of the situation by hiking prices to two or three times the original listing price, and it's become harder than ever to build a new PC.
+Resellers are taking advantage of the situation by \todoadvisor{cite}hiking prices to two or three times the original listing price, and it's become harder than ever to build a new PC.
 

### 2022-02-22 paper/Chapters/Background.tex
Fix comments for Background chapter

--- 2022-02-08 paper/Chapters/Background.tex
@@
 Chapter \ref{SpecializationOfComputers} discusses the history of how computers were built for a particular function, and how the advancement of technology has brought the \enquote{Personal Computer} seen know today.
-\todo{Too much detail?}
-Chapter \ref{PowerOfTheDesktop} discusses the benefits of a desktop computer, namely the raw power and serviceability.
-Chapter \ref{ConvenienceOfTheLaptop} discusses the benefits and drawbacks of a laptop computer, namely the portability and convenience at the cost of power.
-Chapter \ref{RiseOfTheGamingLaptop} discusses the rise of the gaming laptop, and the closing gap between power and portability at the cost of increased price.
+Chapter \ref{PowerOfTheDesktop} discusses the benefits of a desktop computer, Chapter \ref{ConvenienceOfTheLaptop} discusses the benefits and drawbacks of a laptop computer, and Chapter \ref{RiseOfTheGamingLaptop} discusses the rise of the gaming laptop and the closing gap between power and portability at the cost of increased price.
 Chapter \ref{ThinClients} briefly discusses thin clients, a corporate solution for remote access to a computer.
@@
 
-In the age of monolithic computers the size of entire rooms, it was expected of a computer to only be capable of serving a single function.
+In the age of punch cards and monolithic computers the size of entire rooms, it was expected of a computer to only be capable of serving a single function.
 Before the technology could be miniaturized, every single computer had to be specialized for the job at hand.
@@
 Known for it's componentization, user repairability, and direct connections to power and the network, desktop computers are the best option for a user looking for a workhorse system.
+Pieces of the computer could be bought separately, and single parts could be upgraded or replaced by anyone will a little hardware knowledge.
 Even though they aren't very portable, desktop computers allow consumers to access greater processor power for a lower cost compared to portable devices \citep{Meyer20145RS}.
@@
 This understandably comes with a cost, with the literal cost of a gaming laptop being the most obvious compromise.
+The componentization of such a laptop also often takes a hit with parts being directly soldered to the motherboard, reducing a user's options for repairs or upgrades.
 It is more expensive, and often less user-serviceable, to buy a gaming laptop that has the same performance of a desktop computer -- but given that it is an all-in-one device, it's portability gives it a niche that a desktop cannot hope to rival.
-\todo{Things could be said about the current chip shortage leading to laptops being easier to buy compared to desktops, but that might fit better in the section below.}
 The compromise presented by a gaming laptop helps close the gap between power and portability, but it still falls short of being an all around good solution.
@@
 
+\begin{wrapfigure}{r}{0.375\textwidth}
+  \centering
+  \includegraphics[width=0.3\textwidth]{Figures/ClientronU700}
+  \caption[Thin Client]{A thin client next to a traditional desktop computer \cite{ImageThinClient}.}
+  \label{fig:ThinClient}
+\end{wrapfigure}
+
 On the corporate side of computing, the solution for portability has been the thin client.
@@
 This allows the user to utilize the power of the host machine from a cheap computer that usually doesn't have enough power to be used as a standalone computer.
+Because of it's low power, the thin client cannot be used as a standalone computer, but it's low cost and portability make it a great solution for a corporate user.
 While these clients work well for simple interfacing tasks such as data entry or system management, they are not designed to be used for complex tasks that require quick response times or graphical fidelity.
-\todo{Is more information needed here?}
 
@@
 From the COVID-19 pandemic to massive chip shortages and supply line disruptions, people are now more than ever looking for a way to be flexible in their work and personal lives now and in the future.
-The COVID-19 pandemic has forced people to be able to operate remotely, whether that be for work, school, or communicating with friends and family.
-\todo{It's a little hard to cite this since they are current events. Is that ok?}
-Many office workers used to working on a desktop computer onsite are now needed to work from home, or classes usually taught in person are now held remotely.
+The COVID-19 pandemic has forced people to be able to operate remotely, whether that be for work, school, or communicating with friends and family \cite{levanon_2020}.
+Many office workers used to working on a desktop computer onsite are now needed to work from home, or classes usually taught in person are now held remotely \cite{ASU_covid_19}.
 This shift to remote operation has fueled the need for mobile computers, many of which need to be powerful enough to handle the workload of a desktop computer from anywhere in the world.
-The global chip shortage is also causing issues for the purchase of new devices.
+The global chip shortage is also causing issues for the purchase of new devices \cite{chipshortage_jpmorgan}.
 Parts which are usually readily available for purchase are out of stock with months of delay on back order.
-Resellers are taking advantage of the situation by hiking prices to two or three times the original listing price, and it's become harder than ever to build a new PC.
+Resellers are taking advantage of the situation by hiking prices to two or three times the original listing price, and it's become harder than ever to build a new PC \cite{tamarov_2021}.
+All of these factors combined lead to a very difficult situation for anyone looking to purchase a new computer or parts for their existing computer.
 

### 2022-02-22 paper/Chapters/Background.tex
Start State of the Art chapter

@@
 
-\section{Specialization of Computers}\label{SpecializationOfComputers}
+\section{Specialization of Computers}\label{sec:SpecializationOfComputers}
 
@@
 
-\subsection{Power of the Desktop}\label{PowerOfTheDesktop}
+\subsection{Power of the Desktop}\label{subsec:PowerOfTheDesktop}
 
@@
 
-\subsection{Convenience of the Laptop}\label{ConvenienceOfTheLaptop}
+\subsection{Convenience of the Laptop}\label{subsec:ConvenienceOfTheLaptop}
 
@@
 
-\subsection{Rise of the Gaming Laptop}\label{RiseOfTheGamingLaptop}
+\subsection{Rise of the Gaming Laptop}\label{subsec:RiseOfTheGamingLaptop}
 
@@
 
-\section{Thin Clients}\label{ThinClients}
+\section{Thin Clients}\label{sec:ThinClients}
 
@@
 
-\section{Application to Modern Day}\label{ApplicationToModernDay}
+\section{Application to Modern Day}\label{sec:ApplicationToModernDay}
 

### 2022-02-22 paper/Chapters/StateOfTheArt.tex
Begin working on the Background section
The first words are always the hardest
Feedback from Advisor
Mostly just notes, they will be fixed next
Start State of the Art chapter

@@
 
-\section{Introduction}
+\section{Introduction}\label{sec:Introduction}
 
@@
 
-\section{Hardware Solutions}
+\section{Software Solutions}\label{sec:SoftwareSolutions}
+
+While there are many applications in existence that allow a user to remotely control a computer from another location, few of them offer the power required to stream high-performing applications to a remote device.
+The following subsections will introduce existing attempts to solve this problem as well as drawbacks that come with each implementation.
+
+\subsection{Remote Desktop Protocol}\label{subsec:RemoteDesktopProtocol}
+
+Microsoft's Remote Desktop Protocol (RDP), is a protocol that defines communication between a terminal server and terminal server client for multimedia purposes \cite{rdpDocs}.
+Coming pre-installed on every Windows machine since Windows XP and with clients available for Windows, Mac, and Linux, RDP is an easy to use solution for remotely accessing Windows machines graphically.
+
+Although it is the built-in solution for Windows machines, it isn't without it's drawbacks.
+Firstly, only one graphical session is active at one time while using windows.
+This means if a user is logged into the host computer and someone initiates a connection with RDP, the user using the host computer will be logged out.
+While this isn't always an issue, sensitive programs that don't take well to being logged out may run into issues.
+Secondly, RDP does not support relative mouse movement \cite{burgin_2013}.
+This means every mouse input is sent to the host computer as an absolute position on the screen, rather than as a relative distance from it's previous location.
+Again, while this isn't always an issue, any application that moves the mouse for the user, such as controlling 3D applications or anything dealing with virtual cameras such as animation or game development, cannot be controlled properly.
+
+\subsection{Virtual Network Computing}\label{subsec:VirtualNetworkComputing}
+
+Virtual Network Computing (VNC), is a platform-independent system originally developed by Olivetti \& Oracle Research Labs and later bought and shelved by AT\&T \cite{vncFlavors}.
+While the original implementation is no longer used in a wide capacity, the protocol it was developed on has been expanded and improved to become one of the most flexible yet simple ways to control a computer remotely.
+VNC can run on any modern operating system, and even a web browser can serve as a VNC client.
+
+\subsection{Chrome Remote Desktop}\label{subsec:ChromeRemoteDesktop}
 
@@
 
-\subsection{Thin Clients}
+\subsection{Secure Shell Protocol}\label{subsec:SecureShellProtocol}
 
@@
 
-\subsection{Web Clients}
+\subsection{Web Clients}\label{subsec:WebClients}
 
+\todo{https://ets.engineering.asu.edu/fse-cloud-classroom/}
 \todosection
 
-\section{Software Solutions}
+\section{Hardware Solutions}\label{sec:HardwareSolutions}
 
+While Hardware solutions are not as popular or widespread as software solutions, there has been some innovation in the field that is worth considering.
+Hardware solutions usually focus on building a device that has just enough power to connect and stream data to a remote computer in order to leverage its power for the user to use.
+
+\subsection{Thin Clients}\label{subsec:ThinClients}
+
+As discussed in Section \ref{sec:ThinClients}, Thin Clients are a device commonly seen in the corporate world to allow employees to access data centers and computing clusters from their desks.
+Due to their low cost and ease of deployment, they are often the device of choice to enable employees to access the full power of the business' computing infrastructure without needed to purchase a powerful device for each employee.
+However, since a typical thin client is a headless device, meaning it doesn't come with a monitor, keyboard, mouse, or other peripherals, they are often less portable as one might thing.
+A thin client is usually set up once, and then left in that place for as long as a desktop would stay.
+Especially with laptop sales on the rise, thin clients do not have the benefit of being more portable than the alternatives.
+
+\subsection{Nvidia Shield and GameStream}\label{subsec:NvidiaShieldAndGameStream}
+
+\begin{wrapfigure}{r}{0.375\textwidth}
+  \centering
+  \includegraphics[width=0.3\textwidth]{Figures/nvidia-shield-open-ifixit}
+  \caption[Nvidia Shield]{An open Nvidia Shield displaying the home page \cite{ImageNvidiaShield}.}
+  \label{fig:nvidiashield}
+\end{wrapfigure}
+
+The Nvidia Shield game console was Nvidia's first foray into the realm of remote streaming.
+Though it was focused on gaming and media streaming, it was one of the first attempts at using a portable device to stream intensive applications such as games from a host computer \cite{brown_2013}.
+At the steep price point of \$349, it definitely was enthusiast hardware for a device that focused on mobile and desktop gaming over running console games or demanding titles locally.
+But while the the physical device was praised for performance, battery life, and the experience of streaming games to the console, it didn't perform too well in terms of sales.
+It was still widely regarded as a technological breakthrough though, and Nvidia continued to develop the technology into a new device called the Nvidia Shield TV \cite{daniel_2017}.
+Focused on bringing the power of a gaming desktop to a TV, the Nvidia Shield TV dropped the idea of portability in favor of filling a different market focused on comfort.
+While this is no longer a valid device for the purposes of this paper, the developments of the proprietary GameStream technology that powered the Shield and the Shield TV proved to be even more exciting than the physical devices themselves.
+\todo{Does the picture look ok?}
+
+Though the GameStream technology that powers the Nvidia Shield devices is closed source and only officially compatible with the products Nvidia releases, its enticing power and potential drew a community to reverse engineer the protocol.
+In 2013, a group of students at Case Western Reserve University developed Moonlight, an open source implementation of the GameStream protocol \cite{moonlight}.
+This open source implementation allows the development of GameStream compatible clients that can run on other devices, from other computers to embedded ARM devices.
+This stands as a good starting place to develop the solution described in this paper.
+
+\section{Research Questions}\label{sec:ResearchQuestions}
+\todoadvisor{What are the specific questions we're trying to answer}
 \todosection
-
-\subsection{Remote Desktop Protocol}
-
-\todosection
-
-\subsection{Virtual Network Computing}
-
-\todosection
-
-\subsection{Chrome Remote Desktop}
-
-\todosection
-
-\subsection{Secure Shell Protocol}
-
-\todosection

### 2022-02-23 paper/Chapters/Background.tex
Feedback from advisor

@@
 Computers have a long history of being built to serve a particular purpose, often compromising some aspects of the system to bolster others.
-Chapter \ref{SpecializationOfComputers} discusses the history of how computers were built for a particular function, and how the advancement of technology has brought the \enquote{Personal Computer} seen know today.
-Chapter \ref{PowerOfTheDesktop} discusses the benefits of a desktop computer, Chapter \ref{ConvenienceOfTheLaptop} discusses the benefits and drawbacks of a laptop computer, and Chapter \ref{RiseOfTheGamingLaptop} discusses the rise of the gaming laptop and the closing gap between power and portability at the cost of increased price.
-Chapter \ref{ThinClients} briefly discusses thin clients, a corporate solution for remote access to a computer.
-Chapter \ref{ApplicationToModernDay} introduces current world wide context that has an impact on the computing world and the need for power and portability.
+Chapter \ref{sec:SpecializationOfComputers} discusses the history of how computers were built for a particular function, and how the advancement of technology has brought the \enquote{Personal Computer} seen know today.
+Chapter \ref{subsec:PowerOfTheDesktop} discusses the benefits of a desktop computer, Chapter \ref{subsec:ConvenienceOfTheLaptop} discusses the benefits and drawbacks of a laptop computer, and Chapter \ref{subsec:RiseOfTheGamingLaptop} discusses the rise of the gaming laptop and the closing gap between power and portability at the cost of increased price.
+Chapter \ref{sec:ThinClients} briefly discusses thin clients, a corporate solution for remote access to a computer.
+Chapter \ref{sec:ApplicationToModernDay} introduces current world wide context that has an impact on the computing world and the need for power and portability.
 

### 2022-02-23 paper/Chapters/StateOfTheArt.tex
Feedback from advisor

@@
 
-While there are many applications in existence that allow a user to remotely control a computer from another location, few of them offer the power required to stream high-performing applications to a remote device.
-The following subsections will introduce existing attempts to solve this problem as well as drawbacks that come with each implementation.
+While there are many applications in existence that allow a user to remotely control a computer from another location, few of them offer the \todo{Clarify power, since the issue isn't processing power}power required to stream high-performing applications to a remote device.
+The following subsections will introduce existing solutions to this problem as well as drawbacks that come with each implementation.
 
@@
 Secondly, RDP does not support relative mouse movement \cite{burgin_2013}.
+\todo{explain further}
 This means every mouse input is sent to the host computer as an absolute position on the screen, rather than as a relative distance from it's previous location.
@@
 
-As discussed in Section \ref{sec:ThinClients}, Thin Clients are a device commonly seen in the corporate world to allow employees to access data centers and computing clusters from their desks.
+As discussed in Section \ref{sec:ThinClients}, Thin Clients are a family of devices commonly seen in the corporate world to allow employees to access data centers and computing clusters from their desks.
 Due to their low cost and ease of deployment, they are often the device of choice to enable employees to access the full power of the business' computing infrastructure without needed to purchase a powerful device for each employee.
-However, since a typical thin client is a headless device, meaning it doesn't come with a monitor, keyboard, mouse, or other peripherals, they are often less portable as one might thing.
+However, since a typical thin client is a headless device, meaning it doesn't come with a monitor, keyboard, mouse, or other peripherals, they are often less portable as one might think.
 A thin client is usually set up once, and then left in that place for as long as a desktop would stay.
@@
 While this is no longer a valid device for the purposes of this paper, the developments of the proprietary GameStream technology that powered the Shield and the Shield TV proved to be even more exciting than the physical devices themselves.
-\todo{Does the picture look ok?}
 

### 2022-03-12 paper/Chapters/StateOfTheArt.tex
Draft State of the Art Chapter
I need to pick up the pace

@@
 
-\todosection
+\todo[inline]{Should this be a section on it's own or collapse it like the previous chapter?}
+This chapter seeks to provide a summary of existing solutions for remote computer access.
+It will not cover specific applications, but rather the protocols that are used to communicate with a remote computer.
+Generally speaking, there are two types of solutions that have been developed, software solutions, which are covered in Chapter \ref{sec:SoftwareSolutions} and hardware solutions, which are covered in Chapter \ref{sec:HardwareSolutions}.
+Various protocols and implementations are covered in the subsections of those two categories.
 
@@
 
-While there are many applications in existence that allow a user to remotely control a computer from another location, few of them offer the \todo{Clarify power, since the issue isn't processing power}power required to stream high-performing applications to a remote device.
+While there are many applications in existence that allow a user to remotely control a computer from another location, few of them offer the capabilities required to stream high-performing applications to a remote device.
 The following subsections will introduce existing solutions to this problem as well as drawbacks that come with each implementation.
@@
 Secondly, RDP does not support relative mouse movement \cite{burgin_2013}.
-\todo{explain further}
 This means every mouse input is sent to the host computer as an absolute position on the screen, rather than as a relative distance from it's previous location.
-Again, while this isn't always an issue, any application that moves the mouse for the user, such as controlling 3D applications or anything dealing with virtual cameras such as animation or game development, cannot be controlled properly.
+This means any program that moves the user's mouse for them, such as 3D applications with a virtual camera, will be sent the absolute position of the client's mouse.
+This will cause the application to detect a large movement of the mouse instantly, which can result in erratic camera movement as the program interprets the mouse moving a large distance every time the cursor is reset to the center of the screen.
+Again, while this isn't always an issue, tasks such as 3D modeling and animation or game development cannot be controlled properly.
 
@@
 VNC can run on any modern operating system, and even a web browser can serve as a VNC client.
+It was originally built as the simplest form of remote computer control over LAN, allowing system administrators to control almost any kind of computer with a simple, robust, and extremely compatible system.
+However, since its use has gradually outgrown it's original scope, VNC isn't usually the right tool for any job greater than remote system management.
+Because of it's simplicity and focus on being as straightforward as possible, VNC is often found to be lacking in terms of security and speed \cite{vncFlavors}.
+It is not recommended to use VNC over the internet without some secondary form of security, such as SSL or a secure VPN.
 
@@
 
-\todosection
+Google Chromoting\todo{Should this chapter be called Chromoting?
+"Chrome Remote Desktop" is the much more recognizable name though}, implemented publicly as Chrome Remote Desktop, is an open source protocol that enables remote desktop control through any web browser running on the Chromium engine \cite{chromotingBuildInstructions}.
+It prioritizes low bandwidth usage and ease of access, allowing easy server install on every major desktop operating system, and enables any web browser or mobile device to act as a client.
+By employing the VP8 compression format (Most commonly seen in internet video streaming), Chromoting is able to keep network traffic low by only sending a few full images of the server's screen as keyframes and transmitting motion with compressed partial frames \cite{miniorange_chromoting}.
+These partial frames are calculated by the server using the difference between the previous and current frame, allowing the server to only send the difference between the two frames.
+
+This save in bandwidth and processing done by the client makes it the perfect solution for Google's ChromeOS laptops coming out around the same time.
+These low powered machines were built to be laptops running on the power of a web browser with a link back to more powerful computers when the tasks were too great \cite{upson_sengupta_2012}.
+While the protocol works well for remote access, it is limited in certain applications by it's conservative approach to bandwidth usage and reliance on the web browser.
 
@@
 
-\todosection
+The Secure Shell Protocol (SSH), is a protocol for securely transmitting data over an insecure network \cite{rfc4251}.
+Usually used for remote server management, SSH has become the de facto standard for connecting to a remote computer when nothing more than terminal access is needed.
+Paired with the associated SSH file transfer (SFTP) or secure copy (SCP) protocols, SSH provides an incredible amount of flexibility for working with remote computers.
+It's largest and most obvious drawback is limited graphical support.
+While not a problem for it's traditional use-case, it does limit what high-intensity applications can be used with it.
 
-\subsection{Web Clients}\label{subsec:WebClients}
-
-\todo{https://ets.engineering.asu.edu/fse-cloud-classroom/}
-\todosection
+One recent innovation in remote computing has been the introduction of Remote Development using SSH through the Visual Studio Code IDE.
+By opening multiple SSH connections to the host machine at once, Visual Studio Code is able to run the IDE's User interface on the client's machine, and simply send all the written code and commands through to the host machine, utilizing it's power to compile code, run applications, and debug software \cite{vscode_ssh}.
+This enables a development experience comparable to working on a local machine by only transmitting the data that is needed over the internet and constructing the visual user interface on the client.
+Because of this, high-intensity computational programs such as Artificial Intelligence training and simulation can be run on a powerful host machine and only need to transmit the output to the lower-powered client, but graphical applications are still limited.
 
@@
 \section{Research Questions}\label{sec:ResearchQuestions}
-\todoadvisor{What are the specific questions we're trying to answer}
-\todosection
+
+While there are many existing solutions for accessing a computer remotely, many struggle to be performant enough or efficient enough to stream demanding applications to the client without compromising on the user's experience.
+This thesis will seek to build a solution that enables the use the power of a desktop computer from a mobile device in a way that is performant and responsive in response to the following questions:
+
+\begin{enumerate}
+  \item \textbf{Hardware Feasibility:} \emph{What hardware is needed in order to power a mobile device capable of acting as a client?}
+  \item \textbf{Hardware Cost:} \emph{Is it feasible to construct such as device at a lower or comparable cost to existing solutions?}
+  \item \textbf{Communication Protocol:} \emph{Does a protocol exist that is efficient enough to stream demanding applications to a client?}
+  \item \textbf{Software Application:} \emph{Can a software solution be built to utilize this protocol in a manner that performant enough?}
+  \item \textbf{Software Performance:} \emph{Can this solution be powerful enough to stream demanding applications to a client while remaining performant enough to be usable?}
+\end{enumerate}

### 2022-03-15 paper/Chapters/Hardware.tex
Start working on the Hardware chapter

@@
 
-\section{Requirements}
+\section{Requirements}\label{sec:HardwareRequirements}
+
+This section will explain the different constraints taken into account while developing the hardware portion of this project.
+First, the hardware must be performant enough to provide a positive user experience (\ref{sec:HardwarePerformance}), and secondly the product must be cheap enough to justify it over similar alternatives (\ref{sec:HardwareCost}).
+
+\subsection{Performance}\label{sec:HardwarePerformance}
+
+When considering how much power needs to be incorporated in to the miniature computer, it is important to consider a number of factors, such as: having enough power to stream audio and video over the internet, not consuming too much power so as to require an expensive battery, have a fast enough connection to the internet to deliver data with the least possible latency, and having a reasonable footprint to be used as a mobile device.
+
+Immediately, the device must be capable of decoding and rendering video that is streamed over the internet; a relatively expensive process that cannot easily be done with something such as a micro controller or even a single core CPU seen on Single Board Computers (SBC) such as the \todo{pic?}Raspberry Pi Zero.
+Though at the same time, it is important to not simply build a powerful system akin to a laptop that will consume more power.
+It makes no sense to built a powerful mini-computer when all it will ever be used to is to connect to another computer, and the added power draw would require a larger battery in order to stay operational for decent periods of time.
+Instead it should be cut down and distilled into something that can do it's job as efficiently as possible.
+
+The second major performance requirement of the hardware is the ability to stream this data over the internet as quickly as possible.
+Because all of the data being sent to the device is live information from the host machine, the video and audio streams cannot simply be buffered in advance and played when it's ready like a recorded video.
+Instead, the data must be displayed as it is received so that the user can have a smooth experience without having to wait for their keystrokes or mouse movements to be registered.
+While Wi-Fi is growing ever quicker and more reliable, and should absolutely be included as a way to connected to the internet, the best way to ensure a stable and speedy connection is to use an Ethernet Cable.
+This will help the device work on a consistent connection when used at a desk, while also providing a way to connect while on-the-go.
+
+This leads to the desire for a portable design that can be used both in a conducive environnement such as a desk or while mobile.
+The hardware should have the ability to be powered by a chord plugging into the device, or by an internal battery.
+This alone presents a number of difficult concerns, but those will be addressed later in this chapter\todo{cite}.
+Such a requirement leads to a clear benefit of integrating crucial computers such as a battery and a screen into the hardware design, while leaving as many ports open such as ethernet and USB for the user to take advantage of.
+
+\subsection{Cost}\label{sec:HardwareCost}
+
+Because this project consists of both a Hardware and Software component to make up a complete product, there must be reason for both to exist.
+While the software has the benefit of being the foundation for the connection between host and client machines, there are already hardware solutions that would theoretically be able to run such software.
+Since the goal of this project is to save money by not needing to purchase an expensive mobile device to connect to a powerful desktop computer, there must be benefit in the hardware to warrant it's existence over repurposing something like an old laptop.
+The most obvious metric of this is the actual cost of the hardware.
+If the hardware can be produced at a high enough quality at a low enough cost to bring value over just a software solution, then it should be considered moving forward.
+Otherwise it may prove that the development of the software should focus on compatibility so the hardware can be provided by the end user.
+
+\section{Choosing Parts}\label{sec:ChoosingParts}
+
+When starting to choose parts for the hardware, \todo{First person, now this is me explaining what I did?}I had many difficult questions to face.
+First and foremost, as a Software Engineer with no experience with Electrical Engineering, I had to make sure I could build the hardware in a way that I could handle.
+Specifically, I knew I shouldn't even attempt to build a Single Board Computer on my own, but I knew I also did not want to purchase an off the shelf SBC that was not tailored to the task at hand such as a traditional Raspberry Pi.
+Instead I opted for a \emph{Raspberry Pi Compute Module 4 (CM4)} which has \enquote{The power of Raspberry Pi 4 in a compact form factor for deeply embedded applications} \cite{rpi_cm4}.
+Boasting an incredibly small size and streamlined I/O, there only way to communicate with the Compute Module was through two tiny mezzanine connectors as shown in Figure \ref{fig:rpi_cm4_mezzanine}.
+With no other pin outs for anything such as power, HDMI, or USB, everything had to be built on a custom Printed Circuit Board (PCB).
+
+\begin{figure}[h]
+  \centering
+  \begin{minipage}{0.45\textwidth}
+    \centering
+    \includegraphics[width=0.9\textwidth]{Figures/rpi_cm4}
+    \captionsetup{width=.75\linewidth}
+    \caption[Raspberry Pi Compute Module 4]{Raspberry Pi Compute Module 4, sizing\newline55 x 40mm}
+    \label{fig:rpi_cm4}
+  \end{minipage}\hfill
+  \begin{minipage}{0.45\textwidth}
+    \centering
+    \includegraphics[width=0.9\textwidth]{Figures/rpi_cm4_mezzanine}
+    \captionsetup{width=.75\linewidth}
+    \caption[Raspberry Pi Compute Module 4 Pin out]{Close-up of the pin out, with a 0.04mm pitch between pins}
+    \label{fig:rpi_cm4_mezzanine}
+  \end{minipage}
+\end{figure}
+
+This PCB would have to be designed so that the CM4 could be plugged into the board and use it for all it's inputs and outputs.
+This, however, is only the beginning of the story.
+Since I would not be working with an off the shelf SBC, I would have to not only design the PCB but also pick out all the components that would go onto the PCB, such as the USB ports themselves, the USB driver chip, and related components needed for operation.
+Because the CM4 is designed to be as low-powered and compact as possible, some features that would usually be expected of such a device, like USB ports, are disabled by default.
+Instead, the board designer must build out that functionality manually.
+While this helps greatly by removing extraneous components and features that aren't needed for this project, it also provides me with a \todo{does this sound sarcastic?}great opportunity to learn exactly what parts are being used in the hardware.
+
+\section{Prototyping}\label{sec:Prototyping}
 
@@
 
-\subsection{Performance}
+\section{Designing the Printed Circuit Board}\label{sec:DesigningThePCB}
 
@@
 
-\subsection{Cost}
+\section{Manufacturing the Circuit Board}\label{sec:ManufacturingThePCB}
 
 \todosection
-
-\section{Choosing Parts}
-
-\todosection
-
-\section{Prototyping}
-
-\todosection
-
-\section{Designing the Printed Circuit Board}
-
-\todosection
-
-\section{Manufacturing the Circuit Board}
-
-\todosection

### 2022-03-16 paper/Chapters/Hardware.tex
Draft prototyping sections
The next section is going to be meaty, it might need to be broken up somehow

@@
 
-This PCB would have to be designed so that the CM4 could be plugged into the board and use it for all it's inputs and outputs.
+This PCB would have to be designed so that the CM4 could be plugged into the board using the mezzanine connectors and rely on it for all inputs and outputs.
 This, however, is only the beginning of the story.
-Since I would not be working with an off the shelf SBC, I would have to not only design the PCB but also pick out all the components that would go onto the PCB, such as the USB ports themselves, the USB driver chip, and related components needed for operation.
-Because the CM4 is designed to be as low-powered and compact as possible, some features that would usually be expected of such a device, like USB ports, are disabled by default.
-Instead, the board designer must build out that functionality manually.
+Since I would not be working with an off the shelf SBC like a regular Raspberry PI, I would have to not only design the PCB but also pick out all the components that would go onto the PCB, such as the USB ports themselves, the USB driver chip, and related components needed for operation.
+Because the CM4 is designed to be as low-powered and compact as possible, some features that would usually be expected of such a device, like the USB ports, are disabled by default and must instead be built by the board designer.
 While this helps greatly by removing extraneous components and features that aren't needed for this project, it also provides me with a \todo{does this sound sarcastic?}great opportunity to learn exactly what parts are being used in the hardware.
+
+An invaluable resource while designing the PCB was the Official I/O board for the CM4 \cite{rpi_cm4_io}.
+Though I'd still have to build my PCB board on my own and replace components due to the ongoing chip shortage, I wouldn't need to start from scratch.
+Another massive benefit from working within the Raspberry Pi Ecosystem is their \todo{(re)move?}incredible documentation and community surrounding the boards.
+Around the time I was starting this project, the CM4 was just starting to get into the hands of makers and developers, and only a couple custom boards had actually been made by anyone other than Raspberry Pi themselves.
+Even though this leaves the community with less of a bulk of knowledge to work with, it has provided a sort of comradery with everyone helping each other learn the new Compute Module together.
+Members of the CM4 engineering team would even assist those on the forums to help them pick parts and design their own boards.
 
@@
 
-\todosection
+Whenever working with electrical projects such as this, it is always a good idea to build up prototypes before committing to a final design that will be produced at the highest quality.
+I have had experience with building breadboards and even some breakout boards for some components on other projects in the past, but due to the nature of the complex Surface Mounted Devices (SMD) required for this project as well as the dependence on the CM4's tiny mezzanine connectors, building a prototype using a breadboard is infeasible for the amount of time and resources at my disposal.
+For example, the USB controller's footprint packs 64 pins into as small of a package as possible, requiring a breakout board of considerable size in order to access each pin separately by hand in a way that could be easily prototyped with.
+Not to mention the difficulty in wiring up that many pins along with the other related components which would need their own breakout boards either purchased or custom-designed.
+
+\begin{figure}[h]
+  \centering
+  %Height instead of width because the image is rotated
+  \includegraphics[height=0.9\textwidth,angle=90]{Figures/smt_breakout}
+  \captionsetup{width=.8\linewidth}
+  \caption[SMD Breakout Board]{A breakout board similar to what would be needer for the USB controller used on the custom PCB}
+  \label{fig:smt_breakout}
+\end{figure}
+
+Instead, I opted to build out my first board in a piecewise fashion, allowing each section of the board a way to fail without interrupting the rest of the board.
+This allows me to test each function of the board individually and in isolation so that changes could be made incrementally without needing to worry about their integration with the rest of the board.
+For example, in order to build a system that would allow the board to be powered via USB or an internal battery, as well as charge the battery simultaneously, I would need to build a complex circuit consisting of parts neither I nor anyone I knew had used before.
+Since power is obviously a critical component of the board, I would also need to build a secondary way of feeding power directly the board to bypass all that circuitry in case it fails.
+In doing so, I can still power the board and test it's other features while developing a fix for that portion of the circuitry for the next iteration.
 

### 2022-03-19 paper/Chapters/Hardware.tex
Change first person to third person in Hardware chapter

@@
 
-When starting to choose parts for the hardware, \todo{First person, now this is me explaining what I did?}I had many difficult questions to face.
-First and foremost, as a Software Engineer with no experience with Electrical Engineering, I had to make sure I could build the hardware in a way that I could handle.
-Specifically, I knew I shouldn't even attempt to build a Single Board Computer on my own, but I knew I also did not want to purchase an off the shelf SBC that was not tailored to the task at hand such as a traditional Raspberry Pi.
-Instead I opted for a \emph{Raspberry Pi Compute Module 4 (CM4)} which has \enquote{The power of Raspberry Pi 4 in a compact form factor for deeply embedded applications} \cite{rpi_cm4}.
+When starting to choose parts for the hardware, there are many difficult questions to face.
+First and foremost, the hardware had to be built with the time and resources available for this project.
+Specifically, building a Single Board Computer from scratch would be completely infeasible, but the goal was also not to simply purchase an off the shelf SBC that was not tailored to the task at hand such as a traditional Raspberry Pi.
+Instead, a better suited solution would be to utilize a \emph{Raspberry Pi Compute Module 4 (CM4)} which has \enquote{The power of Raspberry Pi 4 in a compact form factor for deeply embedded applications} \cite{rpi_cm4}.
 Boasting an incredibly small size and streamlined I/O, there only way to communicate with the Compute Module was through two tiny mezzanine connectors as shown in Figure \ref{fig:rpi_cm4_mezzanine}.
@@
 This, however, is only the beginning of the story.
-Since I would not be working with an off the shelf SBC like a regular Raspberry PI, I would have to not only design the PCB but also pick out all the components that would go onto the PCB, such as the USB ports themselves, the USB driver chip, and related components needed for operation.
+Since the project would not use an off the shelf SBC like a regular Raspberry PI, the project would require only designing the PCB but also picking out all the components that would go onto the PCB, such as the USB ports themselves, the USB driver chip, and related components needed for operation.
 Because the CM4 is designed to be as low-powered and compact as possible, some features that would usually be expected of such a device, like the USB ports, are disabled by default and must instead be built by the board designer.
@@
 An invaluable resource while designing the PCB was the Official I/O board for the CM4 \cite{rpi_cm4_io}.
-Though I'd still have to build my PCB board on my own and replace components due to the ongoing chip shortage, I wouldn't need to start from scratch.
+Though the PCB board would still have to be custom built and components would need to be replaced due to the current ongoing chip shortage, it provides a great foundation instead of starting from scratch.
 Another massive benefit from working within the Raspberry Pi Ecosystem is their \todo{(re)move?}incredible documentation and community surrounding the boards.
-Around the time I was starting this project, the CM4 was just starting to get into the hands of makers and developers, and only a couple custom boards had actually been made by anyone other than Raspberry Pi themselves.
+Around the time this project was being started, the CM4 was just starting to get into the hands of makers and developers, and only a couple custom boards had actually been made by anyone other than Raspberry Pi themselves.
 Even though this leaves the community with less of a bulk of knowledge to work with, it has provided a sort of comradery with everyone helping each other learn the new Compute Module together.
@@
 Whenever working with electrical projects such as this, it is always a good idea to build up prototypes before committing to a final design that will be produced at the highest quality.
-I have had experience with building breadboards and even some breakout boards for some components on other projects in the past, but due to the nature of the complex Surface Mounted Devices (SMD) required for this project as well as the dependence on the CM4's tiny mezzanine connectors, building a prototype using a breadboard is infeasible for the amount of time and resources at my disposal.
+Usually breadboards and some breakout boards for Surface Mounted Devices (SMD) are a great way to develop a circuit without creating a full PCB; but due to the nature of the complex microchips required for this project as well as the dependence on the CM4's tiny mezzanine connectors, building a prototype using a breadboard is impractical given the time and resources available.
 For example, the USB controller's footprint packs 64 pins into as small of a package as possible, requiring a breakout board of considerable size in order to access each pin separately by hand in a way that could be easily prototyped with.
@@
 
-Instead, I opted to build out my first board in a piecewise fashion, allowing each section of the board a way to fail without interrupting the rest of the board.
-This allows me to test each function of the board individually and in isolation so that changes could be made incrementally without needing to worry about their integration with the rest of the board.
-For example, in order to build a system that would allow the board to be powered via USB or an internal battery, as well as charge the battery simultaneously, I would need to build a complex circuit consisting of parts neither I nor anyone I knew had used before.
-Since power is obviously a critical component of the board, I would also need to build a secondary way of feeding power directly the board to bypass all that circuitry in case it fails.
-In doing so, I can still power the board and test it's other features while developing a fix for that portion of the circuitry for the next iteration.
+Instead, the first board was built out in a piecewise fashion, allowing each section of the board a way to fail without interrupting the rest of the board.
+This allows each function of the board to be tested individually and in isolation so that changes could be made incrementally without needing to worry about their integration with the rest of the board.
+\todo{convert first person to third person}For example, in order to build a system that would allow the board to be powered via USB or an internal battery, as well as charge the battery simultaneously, I would need to build a complex circuit consisting of parts neither I nor anyone I knew had used before.
+Since power is obviously a critical component of the board, A secondary way of feeding power directly the board was built to bypass all that circuitry in case it fails.
+In doing so, the board can still be powered and it's other features can be tested while a fix for that portion of the circuitry is developed for the next iteration.
 

### 2022-03-21 paper/Chapters/Hardware.tex
Feedback from advisor

@@
 \label{Chapter4} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
+
+\todosection
 
@@
 
-Because this project consists of both a Hardware and Software component to make up a complete product, there must be reason for both to exist.
+Because this project consists of both a Hardware and Software component to make up a complete product, there must be reason for both to be developed.
 While the software has the benefit of being the foundation for the connection between host and client machines, there are already hardware solutions that would theoretically be able to run such software.
@@
 First and foremost, the hardware had to be built with the time and resources available for this project.
-Specifically, building a Single Board Computer from scratch would be completely infeasible, but the goal was also not to simply purchase an off the shelf SBC that was not tailored to the task at hand such as a traditional Raspberry Pi.
+Specifically, building a Single Board Computer\todoadvisor{introduce acronyms the first time they're used}\todo{This was introduced in 4.1.1, should it also be done here?} from scratch would be completely infeasible, but the goal was also not to simply purchase an off the shelf SBC that was not tailored to the task at hand such as a traditional Raspberry Pi.
 Instead, a better suited solution would be to utilize a \emph{Raspberry Pi Compute Module 4 (CM4)} which has \enquote{The power of Raspberry Pi 4 in a compact form factor for deeply embedded applications} \cite{rpi_cm4}.
@@
 This PCB would have to be designed so that the CM4 could be plugged into the board using the mezzanine connectors and rely on it for all inputs and outputs.
-This, however, is only the beginning of the story.
+This, however, is only the beginning of what is required to develop the product.
 Since the project would not use an off the shelf SBC like a regular Raspberry PI, the project would require only designing the PCB but also picking out all the components that would go onto the PCB, such as the USB ports themselves, the USB driver chip, and related components needed for operation.
@@
 Usually breadboards and some breakout boards for Surface Mounted Devices (SMD) are a great way to develop a circuit without creating a full PCB; but due to the nature of the complex microchips required for this project as well as the dependence on the CM4's tiny mezzanine connectors, building a prototype using a breadboard is impractical given the time and resources available.
-For example, the USB controller's footprint packs 64 pins into as small of a package as possible, requiring a breakout board of considerable size in order to access each pin separately by hand in a way that could be easily prototyped with.
+\begin{wrapfigure}{o}{0.375\textwidth}
+  \centering
+  \includegraphics[width=0.3\textwidth]{Figures/64-LQFP}
+  \caption[USB Controller]{The USB controller, with 0.5mm between pins}
+  \label{fig:USBController}
+\end{wrapfigure}
+For example, the USB controller chip only has 0.5mm of space between each of it's 64 pins (Figure \ref{fig:USBController}), which would require a breakout board at least 7 times larger than the chip so that each each pin can be easily accessed by hand.
 Not to mention the difficulty in wiring up that many pins along with the other related components which would need their own breakout boards either purchased or custom-designed.

### 2022-03-21 paper/Chapters/StateOfTheArt.tex
Feedback from advisor

@@
 
-\section{Introduction}\label{sec:Introduction}
-
-\todo[inline]{Should this be a section on it's own or collapse it like the previous chapter?}
 This chapter seeks to provide a summary of existing solutions for remote computer access.
@@
 
+\section{Summary}\label{sec:StateOfTheArtSummary}
+
+\todosection
+
 \section{Research Questions}\label{sec:ResearchQuestions}
@@
   \item \textbf{Software Application:} \emph{Can a software solution be built to utilize this protocol in a manner that performant enough?}
-  \item \textbf{Software Performance:} \emph{Can this solution be powerful enough to stream demanding applications to a client while remaining performant enough to be usable?}
 \end{enumerate}

### 2022-03-22 paper/Appendices/AppendixA.tex
Appendix A and B

\chapter{PCB Specifications} % Main appendix title

\label{AppendixA} % For referencing this appendix elsewhere, use \ref{AppendixA}

\section{PCB Schematic}

\begin{figure}[h]
  \centering
  \includegraphics[width=1\textwidth,page=1]{Figures/kicad/lathrum_thesis_schematic.pdf}
  \captionsetup{width=.8\linewidth}
  \caption[Top Level Schematic]{The top level of the PCB board schematic}
  \label{fig:pcb_schematic_top}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=1\textwidth,page=2]{Figures/kicad/lathrum_thesis_schematic.pdf}
  \captionsetup{width=.8\linewidth}
  \caption[Power Schematic]{Schematic layout for the power circuit, including battery management}
  \label{fig:pcb_schematic_battery}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=1\textwidth,page=3]{Figures/kicad/lathrum_thesis_schematic.pdf}
  \captionsetup{width=.8\linewidth}
  \caption[CM4 Schematic]{Schematic layout for interfacing with the Compute Module 4}
  \label{fig:pcb_schematic_cm4}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[width=1\textwidth,page=4]{Figures/kicad/lathrum_thesis_schematic.pdf}
  \captionsetup{width=.8\linewidth}
  \caption[USB Schematic]{Schematic layout for the USB hub and ports}
  \label{fig:pcb_schematic_usb}
\end{figure}

% Make sure next section is on a new page
\clearpage
\section{PCB Layout}

\begin{figure}[h]
  \centering
  %Height instead of width because the image is rotated
  \includegraphics[height=0.9\textwidth,angle=90,page=1]{Figures/kicad/lathrum_thesis_layout_front.pdf}
  \captionsetup{width=.8\linewidth}
  \caption[PCB Front Layout]{Physical board layout for the front side of the PCB}
  \label{fig:pcb_layout_front}
\end{figure}

\begin{figure}[h]
  \centering
  %Height instead of width because the image is rotated
  \includegraphics[height=0.9\textwidth,angle=90,page=1]{Figures/kicad/lathrum_thesis_layout_back.pdf}
  \captionsetup{width=.8\linewidth}
  \caption[PCB Back Layout]{Physical board layout for the back side of the PCB}
  \label{fig:pcb_layout_back}
\end{figure}

### 2022-03-22 paper/Appendices/AppendixB.tex
Appendix A and B

\chapter{Bill of Materials} % Main appendix title

\label{AppendixB} % For referencing this appendix elsewhere, use \ref{AppendixA}

\begin{longtable}{|>{\raggedright\arraybackslash}p{0.2\linewidth}|c|l|l|}%
  \hline
  \label{tab:BillOfMaterials}
  \bfseries Reference \newline Designator& \bfseries Count & \bfseries Manufacturer's Number &\bfseries Value % Columns
  \csvreader[respect percent=true, respect underscore=true, respect sharp=true, table head=\toprule, table foot=\bottomrule]{Figures/kicad/lathrum_thesis_bom.csv}{}% Data
  {\\\hline\csvcoli&\csvcolii&\csvcolv&\csvcoliii}% Rows
  \\\hline
\end{longtable}

### 2022-03-24 paper/Chapters/Hardware.tex
Fix feedback issues and start Design PCB section

@@
 
-Immediately, the device must be capable of decoding and rendering video that is streamed over the internet; a relatively expensive process that cannot easily be done with something such as a micro controller or even a single core CPU seen on Single Board Computers (SBC) such as the \todo{pic?}Raspberry Pi Zero.
+Immediately, the device must be capable of decoding and rendering video that is streamed over the internet.
+Such video processing is a relatively expensive process that cannot easily be done with something such as a micro controller or even a single core CPU seen on Single Board Computers (SBC) such as the Raspberry Pi Zero due to the lack of working memory and processing power \cite{picockpit_2021}.
+\todoadvisor{Add more detail?}
 Though at the same time, it is important to not simply build a powerful system akin to a laptop that will consume more power.
@@
 Otherwise it may prove that the development of the software should focus on compatibility so the hardware can be provided by the end user.
+
+Generally speaking, the goal for the hardware of this project is to cost approximately \$100-\$200 to produce a single unit.
+This comes with the understanding that hardware production in bulk often costs significantly less than assembling a single product.
+The reasoning for this cost comes down to reasonableness: the price of the final product should be less than what it would cost to purchase or refurbish an existing computer to do the job of this project.
+Otherwise, as stated before, the software of this project may prove more valuable than the hardware as long as compatibility is kept in the forefront of development.
 
@@
 
+Following the example put forward by the existing boards, PCB parts were chosen to be integrated into the circuit as safely as possible.
+Thankfully due to most parts following international standards, many parts were chosen simply because they followed the same standard as the CM4 (such as USB 2.0, or \emph{1PORT 1000 BASE-T} Ethernet) and were in stock.
+What follows is a table of some components that had to be purchased apart from those that had been tested by the community, and the reason for their selection.
+A full Bill of Materials can be found in Appendix \ref{AppendixB}.
+
+\renewcommand*{\arraystretch}{1.5} % inner table padding
+\begin{longtable}{|>{\raggedright\arraybackslash}p{0.2\linewidth}|>{\raggedright\arraybackslash}p{0.33\linewidth}|>{\raggedright\arraybackslash}p{0.46\linewidth}|}
+  \hline
+  \bfseries Purpose & \bfseries Manufacturer's Number &\bfseries Reason \\% Columns
+  \hline
+  USB-C Power in & \emph{USB4125-GF-A} by \enquote{GCT} & Most projects of this scale are powered by micro usb, but with USB-C gaining popularity, this component was picked for its low price and exclusive ability to transfer power without transfering data.
+\\
+  \hline
+  Ethernet Port & \emph{ARJM11D7-502-AB-EW2} by \enquote{Abracon LLC} & One of the few in-stock ethernet ports that supports \emph{1000 BASE-T} usage.
+The CM4 recommends the \emph{MagJack-A70-112-331N126} produced by \enquote{LINK-PP}, but that part is out of stock for at least another year as of writing.
+\\
+  \hline
+  USB Port & \emph{SS-52100-001} by \enquote{Stewart Connector} & Single flat port that supports USB 2.0. \\
+  \hline
+  HDMI Port & \emph{SS-53000-001} by \enquote{Stewart Connector} & Single flat port that supports HDMI 1.4. \\
+  \hline
+  Flat Flexible Cable Connector & \emph{FFC3B07-20-T} by \enquote{GCT} & A generic Flat Flexible Cable port that can be connected to anything.
+In this case, one is connected to the build-in screen's HDMI port, and another is connected to the screen's touch/power USB port.
+\\
+  \hline
+  2-Pin Through Hole Header & \emph{PREC002SAAN-RC} by \enquote{Sullins Connector Solutions} & A generic 2-pin header than can be used to bypass the power circuit as discussed in Section \ref{sec:Prototyping}.
+\\
+  \hline
+  HDMI Power Regulator & \emph{RT9742ENGJ5} by \enquote{Richtek USA Inc.} & The CM4 recommends the \emph{RT9742SNGV} from the same manufacturer, but it is out of stock.
+This part has slight differences that call for tweaks in the final layout.
+\\
+  \hline
+  SD-Card Power Regulator & \emph{RT9742VGJ5} by \enquote{Richtek USA Inc.} & The CM4 recommends the \emph{RT9742GGJ5F} from the same manufacturer, but it is out of stock.
+This part has slight differences that call for tweaks in the final layout.
+\\
+  \hline
+  Battery Voltage Regulator and Boost Converter & \emph{TPS61090RSAR} by \enquote{TPS61090RSAR} & Used to charge balance power usage from the USB-C port or the attached battery, favoring the external power when available.
+\\
+  \hline
+  Li-Po Battery Charge Management Controller & \emph{MCP73871-2CCI\_ML} by \enquote{Microchip Technology} & Used to charge the attached battery when attached to external power through the USB-C port.
+\\
+  \hline
+  4 Port USB Hub Controller & \emph{USB2504A-JT} by \enquote{Microchip Technology} & Used to control the four USB 2.0 ports and send the data to the CM4.
+The CM4 recommends the \emph{USB2514B-I/M2} by the same manufacturer, however it and all the parts in it's family are out of stock with at least a 52-week back order.
+Out of all the components, this has the highest potential to be problematic.\\
+  \hline
+  \caption[Selected PCB Components]{A non-exhaustive list of components that were chosen for this project.}
+  \label{tab:SelectedComponents}
+\end{longtable}
+\todo[inline]{Should the above table have pictures? If so, where? Since the table is already pretty dense}
+
 \section{Prototyping}\label{sec:Prototyping}
 
-Whenever working with electrical projects such as this, it is always a good idea to build up prototypes before committing to a final design that will be produced at the highest quality.
-Usually breadboards and some breakout boards for Surface Mounted Devices (SMD) are a great way to develop a circuit without creating a full PCB; but due to the nature of the complex microchips required for this project as well as the dependence on the CM4's tiny mezzanine connectors, building a prototype using a breadboard is impractical given the time and resources available.
-\begin{wrapfigure}{o}{0.375\textwidth}
+\begin{wrapfigure}{o}{0.37\textwidth}
   \centering
@@
 \end{wrapfigure}
+Whenever working with electrical projects such as this, it is always a good idea to build up prototypes before committing to a final design that will be produced at the highest quality.
+Usually breadboards and some breakout boards for Surface Mounted Devices (SMD) are a great way to develop a circuit without creating a full PCB; but due to the nature of the complex microchips required for this project as well as the dependence on the CM4's tiny mezzanine connectors, building a prototype using a breadboard is impractical given the time and resources available.
 For example, the USB controller chip only has 0.5mm of space between each of it's 64 pins (Figure \ref{fig:USBController}), which would require a breakout board at least 7 times larger than the chip so that each each pin can be easily accessed by hand.
@@
 
-\todosection
+Because of the lack of prototyping options and the complexity of the circuit and it's components, the Printed Circuit Board was designed in a modular and piecewise fashion.
+This allowed any single part of the circuit to fail without interfering with the other components.
+Since the time and budget for this project was limited, the amount of revisioning was minimized by running the circuit by as many people as possible to catch any potential flaws before the board was manufactured.
+
+\begin{wrapfigure}{r}{0.375\textwidth}
+  \centering
+  \includegraphics[width=0.3\textwidth]{Figures/kicad/close-ups/power-front}
+  \caption[PCB Power Circuit]{The front side of the PCB's power circuit}
+  \label{fig:PowerCircuit}
+\end{wrapfigure}
+Development started with arguably the most important part of the board: the power supply.
+If the CM4 could not receive power, none of the project would work.
+A feature of modern devices that has become ubiquitous as much as it has been taken for granted is the ability to charge a battery at the same time it is in use by the device.
+This enables the device to be used portably until the battery runs out and then charge the device without needing to stop using it before moving again.
+Though it proves to be slightly more complicated than simply connecting the two power supplies.
+As seen in Figure \ref{fig:PowerCircuit},
 

### 2022-03-25 paper/Chapters/Hardware.tex
Designing the power, usb, hdmi and start cm4 circuits

@@
   \hline
-  Battery Voltage Regulator and Boost Converter & \emph{TPS61090RSAR} by \enquote{TPS61090RSAR} & Used to charge balance power usage from the USB-C port or the attached battery, favoring the external power when available.
+  Battery Voltage Regulator and Boost Converter & \emph{TPS61090RSAR} by \enquote{TPS61090RSAR} & Used to balance power usage from the USB-C port or the attached battery, favoring the external power when available.
 \\
@@
 
-\begin{wrapfigure}{r}{0.375\textwidth}
+\begin{figure}[b]
   \centering
-  \includegraphics[width=0.3\textwidth]{Figures/kicad/close-ups/power-front}
+  \includegraphics[width=0.33\textwidth]{Figures/kicad/close-ups/power-front}
   \caption[PCB Power Circuit]{The front side of the PCB's power circuit}
   \label{fig:PowerCircuit}
-\end{wrapfigure}
-Development started with arguably the most important part of the board: the power supply.
+\end{figure}
+
+Development started with arguably the most important part of the board: the power supply (schematic in Figure \ref{fig:pcb_schematic_battery}).
 If the CM4 could not receive power, none of the project would work.
@@
 Though it proves to be slightly more complicated than simply connecting the two power supplies.
-As seen in Figure \ref{fig:PowerCircuit},
+As seen in Figure \ref{fig:PowerCircuit}, the source of external power, the USB-C port labeled \emph{J1}, has to travel through two integrated circuits before reaching the CM4 through the trace on the right side.
+The first chip, \emph{U2}, detailed in Table \ref{tab:SelectedComponents} as \emph{MCP73871-2CCI\_ML}, is used to charge the internal battery attached to the left side of the board through the connector labeled \emph{B1}.
+The second chip, \emph{U1}, also detailed in Table \ref{tab:SelectedComponents} as \emph{TPS61090RSAR}, is used to balance the power draw from the USB-C port or the attached battery, favoring the external power when the device is plugged in.
+Thankfully, due to the sensitivity of the two chips, their provided datasheets offer specific configurations that they should be designed in to ensure proper functionality.
+Although very dense, these blueprints promised individual functionality for each chip, all that was left was connecting them together.
+The two of these chips must work in tandem for power to be successfully delivered to the CM4, if either failed, none of the other parts of the board could be tested.
+In the event this happened, a secondary method of powering the board directly was added in the form of the generic two-pin header \emph{J2}.
+These two pins are simply 5v input and ground through-hole pins that bypass all the power circuitry just described.
+Once the CM4 could be powered, the rest of the I/O could be designed.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.33\textwidth]{Figures/kicad/close-ups/usb-front}
+  \caption[PCB USB Circuit]{The front side of the PCB's USB circuit}
+  \label{fig:USBCircuit}
+\end{figure}
+
+The next prominent circuit on the board is the section that controls the USB ports.
+As seen in Figure \ref{fig:USBCircuit} (schematic in Figure \ref{fig:pcb_schematic_usb}), the dense circuit is mostly comprised of the USB controller chip, \emph{U8}, it's associated components, and the four USB ports, two of which are pictured as \emph{J7} and \emph{J8}.
+Similarly to the power circuit, the USB controller chip is a very sensitive component that must be arranged in relation to other components in a certain way in order to ensure it will function properly.
+This includes surrounding the chip with numerous capacitors to normalize any small power fluctuations, using a specific load balancing power distribution switch (\emph{U9}), and an oscillating crystal to provide a stable clock for the chip.
+The chip connects to each of the four downstream USB ports and handles all of the digital signal processing before sending the data to the upstream CM4.
+
+\begin{figure}[b]
+  \centering
+  \begin{subfigure}{.5\textwidth}
+    \centering
+    \includegraphics[width=.6\linewidth]{Figures/kicad/close-ups/hdmi-front}
+    \caption{The front side of the PCB}
+    \label{fig:HDMICircuitFront}
+  \end{subfigure}%
+  \begin{subfigure}{.5\textwidth}
+    \centering
+    \includegraphics[width=.6\linewidth]{Figures/kicad/close-ups/hdmi-back}
+    \caption{The back side of the PCB}
+    \label{fig:HDMICircuitBack}
+  \end{subfigure}
+  \caption{The PCB's HDMI circuit}
+  \label{fig:HDMICircuit}
+\end{figure}
+
+Another vital part of the board is the two HDMI ports provided by the device.
+Thankfully, unlike the USB ports, HDMI ports send information in only one direction, so an extra chip is not necessary to handle the data as an intermediary.
+All that is necessary is to connect the two HDMI ports to the CM4 in the right order and the device will be able to output the rendered video to a display.
+Because this device is intended to be a standalone machine, it makes sense to have essential components like a screen integrated as part of the design.
+The only issue with simply attaching an HDMI cable inside of the device is that the ports are relatively bulky and inflexible, meaning that another solution must be created to fit in the tight space.
+This is accomplished with the Flat Flexible Cable Connector, seen on the right side of Figure \ref{fig:HDMICircuitFront} as \emph{J4}.
+Flat Flexible Cables (FFC) are generic bendable cables that can be used in tight spaces.
+Even though their connectors are completely generic, meaning that they have no set pinout, it is possible to use them given that the receiving screen will take in the FFC in the same order as a HDMI cable.
+Using this knowledge, an internal FFC can be wired for the device's primary screen, while a second tradition HDMI port can be exposed to the user in case they want to use an external monitor.
+This, along with the FFC used for one of the four USB ports, allows a screen to be hooked up to the CM4 internally without needing to expose anything to the user.
+
+Finally, everything must connect to the CM4.
+Figure \ref{fig:CM4Circuit} shows the the two mezzanine connectors the CM4 uses to communicate with the rest of the PCB (schematic in Figure \ref{fig:pcb_schematic_cm4}).
+Remember that there is only 0.04mm of space in between each of the pins, leaving very little space to route the traces where they need to go.
+The first thing to notice is the large stretch of traces on the back side of the board coming from the HDMI ports.
+Because most of the traces come in positive and negative pairs, called differential pairs, the two tracks must be roughly the same length as each other.
+This is because as data is transmitted across the pair of tracks, the two signals need to arrive at their destination at the same time, a time that is determined by the physical distance the signals need to cover.
+Since the speed of electricity is constant, the traces may need to wind back and forth so that the lengths positive and negative tracks are equal.
+Also note the cutout area at the top of the CM4, whose perimeter is stitched by vias; this is where the CM4 houses it's bluetooth and wifi antennas.
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}{.5\textwidth}
+    \centering
+    \includegraphics[width=.5\linewidth]{Figures/kicad/close-ups/cm4-front}
+    \caption{The front side of the PCB}
+    \label{fig:CM4CircuitFront}
+  \end{subfigure}%
+  \begin{subfigure}{.5\textwidth}
+    \centering
+    \includegraphics[width=.5\linewidth]{Figures/kicad/close-ups/cm4-back}
+    \caption{The back side of the PCB}
+    \label{fig:CM4CircuitBack}
+  \end{subfigure}
+  \caption{The circuitry surrounding the CM4's two mezzanine connectors}
+  \label{fig:CM4Circuit}
+\end{figure}
 

### 2022-03-29 paper/Chapters/Hardware.tex
Tweaks and updates to the hardware sections
Now a nice big sprint can begin. Just need the time

@@
 
-\todosection
+There are a lot of factors that go into designing and building an electrical system that fits the needs for this project.
+This process begins by outlining the requirements in Section \ref{sec:HardwareRequirements}, followed by describing the process and methodology of choosing parts in Section \ref{sec:ChoosingParts}.
+Then, the feasibility of prototyping such as system is discussed in Section \ref{sec:Prototyping}, and finally the system is designed and manufactured in Sections \ref{sec:DesigningThePCB} and \ref{sec:ManufacturingThePCB} respectively.
 
@@
 First and foremost, the hardware had to be built with the time and resources available for this project.
-Specifically, building a Single Board Computer\todoadvisor{introduce acronyms the first time they're used}\todo{This was introduced in 4.1.1, should it also be done here?} from scratch would be completely infeasible, but the goal was also not to simply purchase an off the shelf SBC that was not tailored to the task at hand such as a traditional Raspberry Pi.
+Specifically, building a Single Board Computer\todoadvisor{introduce acronyms the first time they're used}\todoquestion{This was introduced in 4.1.1, should it also be done here?} from scratch would be completely infeasible, but the goal was also not to simply purchase an off the shelf SBC that was not tailored to the task at hand such as a traditional Raspberry Pi.
 Instead, a better suited solution would be to utilize a \emph{Raspberry Pi Compute Module 4 (CM4)} which has \enquote{The power of Raspberry Pi 4 in a compact form factor for deeply embedded applications} \cite{rpi_cm4}.
@@
 Because the CM4 is designed to be as low-powered and compact as possible, some features that would usually be expected of such a device, like the USB ports, are disabled by default and must instead be built by the board designer.
-While this helps greatly by removing extraneous components and features that aren't needed for this project, it also provides me with a \todo{does this sound sarcastic?}great opportunity to learn exactly what parts are being used in the hardware.
+While this helps greatly by removing extraneous components and features that aren't needed for this project, it also provides me with a \todoquestion{does this sound sarcastic?}great opportunity to learn exactly what parts are being used in the hardware.
 
@@
 Though the PCB board would still have to be custom built and components would need to be replaced due to the current ongoing chip shortage, it provides a great foundation instead of starting from scratch.
-Another massive benefit from working within the Raspberry Pi Ecosystem is their \todo{(re)move?}incredible documentation and community surrounding the boards.
+Another massive benefit from working within the Raspberry Pi Ecosystem is their \todoquestion{(re)move?}incredible documentation and community surrounding the boards.
 Around the time this project was being started, the CM4 was just starting to get into the hands of makers and developers, and only a couple custom boards had actually been made by anyone other than Raspberry Pi themselves.
@@
 Members of the CM4 engineering team would even assist those on the forums to help them pick parts and design their own boards.
+
+\todo{Add paragraph on DSI vs HDMI}
+
+\todo{Add paragraph RTC}
 
@@
 \end{longtable}
-\todo[inline]{Should the above table have pictures? If so, where? Since the table is already pretty dense}
+\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,inline]{Should the above table have pictures? If so, where? Since the table is already pretty dense}
 
@@
 This allows each function of the board to be tested individually and in isolation so that changes could be made incrementally without needing to worry about their integration with the rest of the board.
-\todo{convert first person to third person}For example, in order to build a system that would allow the board to be powered via USB or an internal battery, as well as charge the battery simultaneously, I would need to build a complex circuit consisting of parts neither I nor anyone I knew had used before.
+\todoquestion{How do I convert this from first person to third person}For example, in order to build a system that would allow the board to be powered via USB or an internal battery, as well as charge the battery simultaneously, I would need to build a complex circuit consisting of parts neither I nor anyone I knew had used before.
 Since power is obviously a critical component of the board, A secondary way of feeding power directly the board was built to bypass all that circuitry in case it fails.
@@
 Because most of the traces come in positive and negative pairs, called differential pairs, the two tracks must be roughly the same length as each other.
-This is because as data is transmitted across the pair of tracks, the two signals need to arrive at their destination at the same time, a time that is determined by the physical distance the signals need to cover.
-Since the speed of electricity is constant, the traces may need to wind back and forth so that the lengths positive and negative tracks are equal.
-Also note the cutout area at the top of the CM4, whose perimeter is stitched by vias; this is where the CM4 houses it's bluetooth and wifi antennas.
+This is because as data is transmitted across the pair of tracks, they need to arrive at their destination at the same time based on the physical distance the signals need to cover.
+Since the speed of electricity is constant throughout the PCB board, the traces may need to wind back and forth so that the lengths of the positive and negative tracks are equal.
+Also note the blank area surrounded by vias at the top of Figure \ref{fig:CM4Circuit}.
+This is where the WiFi and Bluetooth antennas are located on the CM4 board.
+By removing the copper pour from this area on the PCB board and stitching the perimeter with vias, it prevents the board from acting like another antenna and interfering with the signals being sent to and from the CM4.
+Finally, each of the other components are connected to the CM4 board in the same manner as described previously in the chapter, such as the Ethernet adapter, the SD card reader, and remaining ports.
+See Appendices \ref{AppendixA} and \ref{AppendixB} for further information.
 
@@
     \centering
-    \includegraphics[width=.5\linewidth]{Figures/kicad/close-ups/cm4-front}
+    \includegraphics[width=.75\linewidth]{Figures/kicad/close-ups/cm4-front}
     \caption{The front side of the PCB}
@@
     \centering
-    \includegraphics[width=.5\linewidth]{Figures/kicad/close-ups/cm4-back}
+    \includegraphics[width=.75\linewidth]{Figures/kicad/close-ups/cm4-back}
     \caption{The back side of the PCB}

### 2022-03-29 paper/main.tex
Feedback from Advisor
Mostly just notes, they will be fixed next
Fix comments for Background chapter
Draft State of the Art Chapter
I need to pick up the pace
Start working on the Hardware chapter
Draft prototyping sections
The next section is going to be meaty, it might need to be broken up somehow
Appendix A and B
Designing the power, usb, hdmi and start cm4 circuits
Tweaks and updates to the hardware sections
Now a nice big sprint can begin. Just need the time

@@
 
+\usepackage{wrapfig} % Required to wrap figures
+\usepackage{subcaption} % Required for subfigures
+
 \usepackage[colorinlistoftodos,prependcaption,textsize=tiny,obeyFinal]{todonotes} % Todo notes in margin
 
+\setlength{\marginparwidth}{3cm} % The width of the margin for todo notes
+\newcommand{\todoadvisor}[1]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue]{#1}}
+\newcommand{\todoquestion}[1]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green]{#1}}
 \newcommand{\todosection}{\todo[linecolor=olive,backgroundcolor=olive!25,bordercolor=olive,inline]{TODO}}
+
+\usepackage{csvsimple} % Import CSV files for use in tables
+\usepackage{longtable} % Tables that span multiple pages
 
@@
 
-\thesistitle{Building a Mobile Device that Uses the Power of a Desktop Computer} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
+\thesistitle{Building a Mobile Device that Leverages the Power of a Desktop Computer} % Your thesis title, this is used in the title and abstract, print it elsewhere with \ttitle
 \supervisor{Dr. Robert \textsc{Heinrichs}} % Your supervisor's name, this is used in the title page, print it elsewhere with \supname
@@
 		{\scshape\LARGE \univname\par}\vspace{1.5cm} % University name
-		\textsc{\Large Undergraduate Thesis}\\[0.5cm] % Thesis type
+		\textsc{\Large Honors Thesis}\\[0.5cm] % Thesis type
 
@@
 \addchaptertocentry{\abstractname} % Add the abstract to the table of contents
-	The Thesis Abstract is written here (and usually kept to just this page). The page is kept centered vertically so can expand into the blank space above the title too\ldots
+
+With the recent focus of attention towards remote work and mobile computing, the possibility of taking a powerful workstation wherever needed is enticing.
+However, even emerging laptops today struggle to compete with desktops in terms of cost, maintenance, and future upgrades.
+The price point of a powerful laptop is considerably higher compared to an equally powerful desktop computer, and most laptops are manufactured in a way that makes upgrading parts of the machine difficult or impossible, forcing a complete purchase in the event of failure or a component needing an upgrade.
+
+In the case where someone already owns a desktop computer and must be mobile, instead of needing to purchase a second device at full price, it may be possible to develop a low-cost computer that has just enough power to connect to the existing desktop and run all processing there, using the mobile device only as a user interface.
+
 \end{abstract}
@@
 \begin{abbreviations}{ll} % Include a list of abbreviations (a table of two columns)
-
+	\addchaptertocentry{\abbrevname} % Add the Abbreviations to the table of contents
+	\textbf{CM4} & Raspberry Pi \textbf{C}ompute \textbf{M}odule \textbf{4}\\
 	\textbf{CPU} & \textbf{C}entral \textbf{P}rocessing \textbf{U}nit\\
 	\textbf{CRD} & \textbf{C}hrome \textbf{R}emote \textbf{D}esktop\\
+	\textbf{FFC} & \textbf{F}lat \textbf{F}lexible \textbf{C}able\\
 	\textbf{GPU} & \textbf{G}raphics \textbf{P}rocessing \textbf{U}nit\\
-	\textbf{PC} & \textbf{P}ersonal \textbf{C}computer\\
+	\textbf{IDE} & \textbf{I}ntegrated \textbf{D}evelopment \textbf{E}nvironment\\
+	\textbf{PC} & \textbf{P}ersonal \textbf{C}omputer\\
 	\textbf{PCB} & \textbf{P}rinted \textbf{C}ircuit \textbf{B}oard\\
 	\textbf{RDP} & \textbf{R}emote \textbf{D}esktop \textbf{P}rotocol\\
+	\textbf{SBC} & \textbf{S}ingle \textbf{B}oard \textbf{C}omputer\\
+	\textbf{SMD} & \textbf{S}urface \textbf{M}ounted \textbf{D}evice\\
 	\textbf{SSH} & \textbf{S}ecure \textbf{Sh}ell (Protocol)\\
@@
 \include{Appendices/AppendixA}
-%\include{Appendices/AppendixB}
+\include{Appendices/AppendixB}
 %\include{Appendices/AppendixC}

### 2022-03-30 paper/Chapters/Background.tex
(almost) finish draft of hardware section
Still need to investigate the USB issues and maybe add more detail to the end of the chapter, but other than that I have a couple of questions to resolve then I'm happyish with the chapter

@@
 All computer products are a balance, it simply isn't possible to have the perfect combination of portable, powerful, serviceable, and cheap.
-While desktops focus on power and serviceability, laptops often focus portability at the expense of other traits\todo{Theres a word at the tip of my tounge thats a lot better here dangit}.
+While desktops focus on power and serviceability, laptops often focus portability at the expense of other attributes.
 As mobility becomes ever more important, gaming laptops have become more popular to bring power to the portable form factor.

### 2022-03-30 paper/Chapters/Hardware.tex
(almost) finish draft of hardware section
Still need to investigate the USB issues and maybe add more detail to the end of the chapter, but other than that I have a couple of questions to resolve then I'm happyish with the chapter

@@
 Such video processing is a relatively expensive process that cannot easily be done with something such as a micro controller or even a single core CPU seen on Single Board Computers (SBC) such as the Raspberry Pi Zero due to the lack of working memory and processing power \cite{picockpit_2021}.
-\todoadvisor{Add more detail?}
 Though at the same time, it is important to not simply build a powerful system akin to a laptop that will consume more power.
@@
 The hardware should have the ability to be powered by a chord plugging into the device, or by an internal battery.
-This alone presents a number of difficult concerns, but those will be addressed later in this chapter\todo{cite}.
+This alone presents a number of difficult concerns which will be addressed in Section \ref{sec:Prototyping}.
 Such a requirement leads to a clear benefit of integrating crucial computers such as a battery and a screen into the hardware design, while leaving as many ports open such as ethernet and USB for the user to take advantage of.
@@
 
+Thought must also be given to the peripherals that will be included in the design.
+Most importantly, if a screen is going to be included in the design, what kind of screen will be used?
+For projects of this scale, the usual candidates for a relatively performant screen are HDMI screens or DSI screens.
+While HDMI screens are much more common and recognizable, DSI screens are a new development that promises high performance in low cost and low power applications such as embedded systems.
+However, when looking into existing DSI screens, there are a couple of issues that prevent it from being the best choice for this project.
+Even though DSI screens are more power efficient and have built in touch support, it is a closed format which greatly restricts manufactures from producing their own screens \cite{dsi_vs_hdmi}.
+Due to this, there is only one screen available for purchase outside of the official implementation, and both have a lower resolution that what is possible with HDMI screens of the same size.
+So even though HDMI screens are more expensive at this size, they are a more reliable choice compared to the still-developing DSI standard.
+
+Another potential feature to consider is whether to integrate a Real Time Clock (RTC) into the design.
+An RTC is a component commonly found within laptop and desktop computers that is used to keep track of time even though the computer is powered off.
+This is done with a small button battery powering an RTC chip integrated into the motherboard to keep the system's clock in sync with the real world.
+While this is almost never necessary in embedded systems, it is worth considering for this project because of it's use as a mobile computer.
+However, when considering the primary use case of connecting to another computer over the internet, and the potential of needing a linux kernel patch in order to add support for the RTC \cite{rtc_kernelpatch}, it would be much simpler to tell the operating system to simply resync it's clock with an internet clock every time it connects to a network.
+
 An invaluable resource while designing the PCB was the Official I/O board for the CM4 \cite{rpi_cm4_io}.
@@
 Members of the CM4 engineering team would even assist those on the forums to help them pick parts and design their own boards.
-
-\todo{Add paragraph on DSI vs HDMI}
-
-\todo{Add paragraph RTC}
 
@@
 
-\todosection
+\begin{figure}[b]
+  \centering
+  \includegraphics[width=0.5\textwidth]{Figures/kicad/close-ups/stencil}
+  \caption[PCB Stencil]{The PCB's stencil. Areas colored black are where solder paste is applied to solder SMD parts to the board}
+  \label{fig:PCBStencil}
+\end{figure}
+
+\todo[inline]{Insert pictures of PCB. Thinking at least one showing the flaws of the first board, another of the completed board}
+
+Once the board's design has been completed and checked against all electrical, functional, and manufacturing constraints, the Printed Circuit Board can be produced.
+The first step in fabrication is to print the actual PCB.
+While ASU offers a basic printing service for use in classes and academic projects, the requirements of this board, specifically the tiny traces and pads required for the CM4's mezzanine connectors (pictured in Figure \ref{fig:rpi_cm4_mezzanine}), are outside of the printer's capabilities.
+Instead, by recommendation of a professor, the printing was done through JLCPC (\href{https://www.jlcpc.com/}{JLCPCB.com})\todoquestion{Would this be the right format for linking JLCPCB?
+It's not really a source or a reference that would be linked in the bibliography}, a company that provides quick turnaround and low costs for printing PCBs.
+A set of five boards was ordered, allowing for multiple boards to be assembled in case of error, as well as a stencil.
+A PCB stencil is a thin sheet of metal with holes cut in it that line up with the pads on the PCB board.
+This enables the easy application of solder paste onto all of the pads on the board as once, including the difficulty small pads of the CM4's mezzanine connectors.
+While the PCB was being printed, all of the parts found in the Bill of Materials (found in Appendix \ref{AppendixB}) were ordered from sites such as Digi-Key and Mouser.\todoquestion{Should I mention small issues such as the time when I ordered the PCB and before I could order the parts one of them went out of stock?
+I worry that that would break the flow of the paper, and I already mentioned stock issues in a previous section.
+I feel like it's a detail that makes an interesting story, but since this isn't a story paper it's not needed here.}
+In order to account for the production of multiple prototype boards and the possibility of broken parts, enough parts were ordered to produce five complete boards.
+
+Once the PCB and it's component parts arrived, it was time for assembly.
+The first board was assembled by hand using the PCB stencil and a handheld heat gun.
+The stencil, shown as an image in Figure \ref{fig:PCBStencil}, is used by securing it the PCB board so that the holes line up with the pads, and spreading solder paste across it allowing the paste to cover the exposed pads through the holes.
+The stencil was then removed, and each of the surface mounted components were then placed onto the board using a set of tweezers.
+For tricky parts such as the CM4's mezzanine connectors and other small parts with the pads below the component itself, a microscope was used to assist in lining up the pads with the components.
+Once the parts were placed, a heat gun was used to melt the solder paste and fuse the components to the board.
+Following the SMDs, the through-hole components, such as the ports lining the side of the board, were then placed and soldered using a hand-held \todoquestion{soldering iron or solder iron?}soldering iron.
+
+The first board ran into two issues that prevented it from working.
+Firstly, due to improper use of the heat gun, most of the LEDs had popped, resulting in all hardware-level diagnostic indicators being inoperable.
+Secondly, the CM4 connectors were not properly connected to the board, and in an attempt to repair it, the connector components were irreparably damaged and the board itself was damaged slightly.\todo{Definitely want a picture here}
+Instead of attempting to assemble another board by hand, a local third-party company was contacted to assemble the board professionally.
+After delivering the board, stencil, and parts, the second board was assembled.
+
+\begin{figure}[b]
+  \centering
+  \begin{subfigure}{.5\textwidth}
+    \centering
+    \includegraphics[width=.8\linewidth]{Figures/kicad/QFN-20-1EP_4x4mm}
+    \caption{\emph{QFN-20-1EP\_4x4mm}}
+    \label{fig:FootprintQFN204x4_Original}
+  \end{subfigure}%
+  \begin{subfigure}{.5\textwidth}
+    \centering
+    \includegraphics[width=.8\linewidth]{Figures/kicad/QFN-20-1EP_4x4mm_RevA}
+    \caption{\emph{QFN-20-1EP\_4x4mm\_RevA}}
+    \label{fig:FootprintQFN204x4_RevA}
+  \end{subfigure}
+  \caption[Two versions of the QFN-20-1EP\_4x4mm package]{The two footprints for components with the \emph{QFN-20-1EP\_4x4mm} package.
+The original footprint is shown in Figure \ref{fig:FootprintQFN204x4_Original}, while the revised footprint is shown in Figure \ref{fig:FootprintQFN204x4_RevA}}
+  \label{fig:FootprintQFN204x4}
+\end{figure}
+
+The second board was assembled mostly successfully, and allowed the first stage of testing to begin.
+The only issue apart from some ports being too close together with real-world cables was one component having the wrong footprint on the board.
+Unfortunately, the component \emph{U2 (MCP73871-2CCI\_ML)}, is described as using the \emph{QFN-20-1EP\_4x4mm} footprint in it's datasheet\todoquestion{Does this need to be cited?
+Can you cite a datasheet like this?}, but actually uses a revised version of the footprint, shown in Figure \ref{fig:FootprintQFN204x4}.
+This discrepancy meant that that component could not be soldered, and it's associated circuit would not function properly.
+In this case, the component was a part of the battery charging circuit, causing the entire primary power supply to be unusable.
+Thankfully, as described in Section \ref{sec:Prototyping}, the secondary method of powering the board could be used to test the remaining features of the board.
+After making the necessary fixes, a second round of five PCB boards and a stencil was ordered from JLCPCB.
+Because there was still enough components to assemble three more boards, and no components needed to be changed, no more components were ordered at this stage.
+Once the boards were delivered, the revised board was again sent to be professionally assembled.
+
+Once the third board was assembled, the board could be tested in full.
+While each part of the board was at least partially functional, there were still issues preventing it from working completely.
+In particular, USB devices plugged into any of the ports on the board were able to receive power from the board but were unable to transfer data in either direction.
+\todo[inline]{I am currently going through and diagnosing this issue}

### 2022-03-30 paper/Chapters/Hardware.tex
Start the software section

@@
 This section will explain the different constraints taken into account while developing the hardware portion of this project.
-First, the hardware must be performant enough to provide a positive user experience (\ref{sec:HardwarePerformance}), and secondly the product must be cheap enough to justify it over similar alternatives (\ref{sec:HardwareCost}).
+First, the hardware must be performant enough to provide a positive user experience (\ref{subsec:HardwarePerformance}), and secondly the product must be cheap enough to justify it over similar alternatives (\ref{subsec:HardwareCost}).
 
-\subsection{Performance}\label{sec:HardwarePerformance}
+\subsection{Performance}\label{subsec:HardwarePerformance}
 
@@
 
-\subsection{Cost}\label{sec:HardwareCost}
+\subsection{Cost}\label{subsec:HardwareCost}
 

### 2022-03-30 paper/Chapters/Software.tex
Start the software section

@@
 
-\section{Requirements}
+\todosection
+
+\section{Requirements}\label{sec:SoftwareRequirements}
+
+Much like the hardware requirements in section \ref{sec:HardwareRequirements}, the primary requirement of the software is to be performant enough to provide the user with a positive experience.
+This is broken down to include being able to stream and decode video over the internet with as little latency as possible, stream inputs from the user to the remote computer, and retain visual fidelity while the host machine is under heavy load.
+While much of the latency between the host and client is due to the network both devices are connected to, there is still some optimizations that can be made to reduce the time between data being send and images being displayed on the screen.
+Similarly, while the rate at which data is transferred between the host and client cannot exceed the speed of the slowest internet connection of the two devices, there can still be a balance between the bandwidth-saving but lower quality streaming seen in Chrome Remote Desktop (\ref{subsec:ChromeRemoteDesktop}), and the local network but higher quality Nvidia Shield (\ref{subsec:NvidiaShieldAndGameStream}).
+To accomplish this, a compression method similar to CRD will be used, but it will not be bound by the traditional web browser which limit CRD due to the software solution being developed as a native application.
+
+As a secondary objective, it would also be beneficial for the software to be able to run on a variety of different hardware platforms in the event the hardware is determined to be infeasible to develop, as discussed in Section \ref{subsec:HardwareCost}.
+
+\section{Potential Avenues}\label{sec:PotentialAvenues}
+
+As with any software project, there are many potential ways to solve the problem.
+In many cases, the easiest place to start is to begin by ruling out potential avenues that are not feasible.
+For example, as evident by the number of existing solutions that attempt to solve this problem (Chapter \ref{Chapter3}), taking into special account Nvidia's recent foray into remote gaming with the Shield (\ref{subsec:NvidiaShieldAndGameStream}), it is not reasonable to attempt to create a brand new software solution from scratch with the time and resources available.
+It has taken Nvidia, a company worth hundreds of billions of dollars, years to get to the point they are at now \cite{brown_2013}.
+Developing a rival protocol or new software platform is simply not possible for this project.
+Instead, given that Nvidia's GameStream technology is the current leader in the realm of high-performance remote computing, the best way to build a software product that meets the requirements laid out in Section \ref{sec:ResearchQuestions} is to build upon what already exists.
+
+\subsection{NVIDIA GameStream}\label{subsec:NVIDIAGameStream}
+
+\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,inline]{Should this subsection and the following subsection be top level sections?
+The organization of the headers is a little strange, but the content flows pretty well}
+
+Unfortunately, the Nvidia GameStream protocol is both proprietary and a closed source project.
+Very little is officially known about the protocol, other than it uses specific video encoders built into Nvidia Graphics cards from the 600 series from 2012 and newer \cite{gamestream_userguide}.
+Apart from that, everything has been researched and reverse engineered by the community.
+A GameStream session is started by first connecting to a host computer which has enabled the GameStream service through Nvidia's GeForce Experience application.
+Once the service is enabled, the computer will listen on port \emph{47989}\todoquestion{Should this be italicized? quoted? plaintext?} for the following web requests:
+\begin{itemize}
+  \item /serverinfo
+  \item /pair
+  \item /unpair
+  \item /applist
+  \item /launch
+  \item /resume
+  \item /cancel
+\end{itemize}
+Once a session has been started with a \emph{/launch} request, raw data covering video, audio, and input is passed back and forth over a number of ports until the session is terminated.
+An end-to-end example of the protocol's usage is as follows:
+
+\begin{enumerate}
+  \item A user knows that a capable host computer is ready to stream at the address \emph{192.168.1.123}.
+        \begin{enumerate}
+          \item This can be confirmed by sending a HTTP GET request \emph{/serverinfo} to the host computer on port \emph{47989}.
+For example: \newline\emph{http://192.168.1.123:47989/serverinfo?uniqueid=1234}
+        \end{enumerate}
+  \item The user can make a request to \emph{/pair} with a number of parameters to authorize itself with the host.
+        \begin{enumerate}
+          \item Pairing is a two step-process.
+                First, a request to pair is made, at which point the host machine will display a four digit code on it's screen.
+                Then the client must send that code back to the host as a second form of authentication.
+          \item The authentication created by this pairing process lasts until either the client makes an \emph{/unpair} request with it's unique ID or a user on the host machine unpairs the client.
+        \end{enumerate}
+  \item The user can make a request to \emph{/applist} to get a list of applications that the host is able to stream.
+  \item The user can make a request to \emph{/launch} with a number of parameters to instruct the host to create a session with the given settings and launch the specified application.
+        \begin{enumerate}
+          \item The configuration options given include the application to launch, the video resolution to stream, audio setup, input mapping, and multiple forms of authentication.
+        \end{enumerate}
+  \item If the user unexpectedly or accidentally leaves a session without closing it, they can make a request to \emph{/resume} to instruct the host to reopen and resume the session the previous session.
+  \item Once the user is finished, that can make a request to \emph{/cancel} to instruct the host to close the session.
+  \item If the user is completely finished with the host, they can return it to the state they found it in by making a request to \emph{/unpair} with their unique ID, unauthorizing themselves with the host.
+\end{enumerate}
+
+The most difficult part of this process is handling the transfer of data after a session has begun.
+Once a session has launched, data is sent directly between the host and client over a number of ports without going through an easy to understand web protocol.
+This is where the community project Moonlight steps in.
+
+\subsection{Moonlight}\label{subsec:Moonlight}
+
+Moonlight is an unofficial third-party open-source client for the Nvidia GameStream protocol \cite{moonlight}.
+The project is structured as a core library written in \emph{C} with a number of community built clients built on top of it for various platforms.
+In essence, it turns the process described in the previous section (\ref{subsec:NVIDIAGameStream}) into a usable API.
+By reverse engineering the protocol originally implemented by the Nvidia Shield, Moonlight enables the development of a GameStream client for any other platform.
+The Moonlight community has already developed fantastic clients for many platforms such as Windows, Mac, Linux, Android, iOS, Web, and even other gaming consoles.
+All that is needed for the purposes of this project is to make tweaks to best suit the CM4 that drives the hardware and develop a suite of tools to test it's functionality.
+
+\section{Developing for ARM}\label{sec:DevelopingForARM}
 
@@
 
-\section{Potential Avenues}
+\section{Developing Testing and Measurement Tools}\label{sec:DevelopingTestingAndMeasurementTools}
 
 \todosection
-
-\subsection{NVIDIA GameStream}
-
-\todosection
-
-\subsection{Moonlight}
-
-\todosection
-
-\section{Developing for ARM}
-
-\todosection

### 2022-03-31 paper/Appendices/AppendixA.tex
Fixes from feedback, add images of physical board

@@
   \centering
-  \includegraphics[width=1\textwidth,page=1]{Figures/kicad/lathrum_thesis_schematic.pdf}
-  \captionsetup{width=.8\linewidth}
+  %Height instead of width because the image is rotated
+  \includegraphics[height=.8\textwidth,angle=90,page=1]{Figures/kicad/lathrum_thesis_schematic.pdf}
+  \captionsetup{width=1\linewidth}
   \caption[Top Level Schematic]{The top level of the PCB board schematic}
-  \label{fig:pcb_schematic_top}
+  \label{app:pcb_schematic_top}
 \end{figure}
@@
   \centering
-  \includegraphics[width=1\textwidth,page=2]{Figures/kicad/lathrum_thesis_schematic.pdf}
-  \captionsetup{width=.8\linewidth}
+  %Height instead of width because the image is rotated
+  \includegraphics[height=1\textwidth,angle=90,page=2]{Figures/kicad/lathrum_thesis_schematic.pdf}
+  \captionsetup{width=1\linewidth}
   \caption[Power Schematic]{Schematic layout for the power circuit, including battery management}
-  \label{fig:pcb_schematic_battery}
+  \label{app:pcb_schematic_battery}
 \end{figure}
@@
   \centering
-  \includegraphics[width=1\textwidth,page=3]{Figures/kicad/lathrum_thesis_schematic.pdf}
-  \captionsetup{width=.8\linewidth}
+  %Height instead of width because the image is rotated
+  \includegraphics[height=1\textwidth,angle=90,page=3]{Figures/kicad/lathrum_thesis_schematic.pdf}
+  \captionsetup{width=1\linewidth}
   \caption[CM4 Schematic]{Schematic layout for interfacing with the Compute Module 4}
-  \label{fig:pcb_schematic_cm4}
+  \label{app:pcb_schematic_cm4}
 \end{figure}
@@
   \centering
-  \includegraphics[width=1\textwidth,page=4]{Figures/kicad/lathrum_thesis_schematic.pdf}
-  \captionsetup{width=.8\linewidth}
+  %Height instead of width because the image is rotated
+  \includegraphics[height=1\textwidth,angle=90,page=4]{Figures/kicad/lathrum_thesis_schematic.pdf}
+  \captionsetup{width=1\linewidth}
   \caption[USB Schematic]{Schematic layout for the USB hub and ports}
-  \label{fig:pcb_schematic_usb}
+  \label{app:pcb_schematic_usb}
 \end{figure}
@@
   %Height instead of width because the image is rotated
-  \includegraphics[height=0.9\textwidth,angle=90,page=1]{Figures/kicad/lathrum_thesis_layout_front.pdf}
-  \captionsetup{width=.8\linewidth}
+  \includegraphics[height=.91\textwidth,angle=90,page=1]{Figures/kicad/lathrum_thesis_layout_front.pdf}
+  \captionsetup{width=1\linewidth}
   \caption[PCB Front Layout]{Physical board layout for the front side of the PCB}
-  \label{fig:pcb_layout_front}
+  \label{app:pcb_layout_front}
 \end{figure}
@@
   %Height instead of width because the image is rotated
-  \includegraphics[height=0.9\textwidth,angle=90,page=1]{Figures/kicad/lathrum_thesis_layout_back.pdf}
-  \captionsetup{width=.8\linewidth}
+  \includegraphics[height=1\textwidth,angle=90,page=1]{Figures/kicad/lathrum_thesis_layout_back.pdf}
+  \captionsetup{width=1\linewidth}
   \caption[PCB Back Layout]{Physical board layout for the back side of the PCB}
-  \label{fig:pcb_layout_back}
+  \label{app:pcb_layout_back}
 \end{figure}

### 2022-03-31 paper/Chapters/Hardware.tex
Fixes from feedback, add images of physical board

@@
 This comes with the understanding that hardware production in bulk often costs significantly less than assembling a single product.
-The reasoning for this cost comes down to reasonableness: the price of the final product should be less than what it would cost to purchase or refurbish an existing computer to do the job of this project.
+The reasoning for this cost comes down to feasibility: the price of the final product should be less than what it would cost to purchase or refurbish an existing computer to do the job of this project.
 Otherwise, as stated before, the software of this project may prove more valuable than the hardware as long as compatibility is kept in the forefront of development.
@@
 First and foremost, the hardware had to be built with the time and resources available for this project.
-Specifically, building a Single Board Computer\todoadvisor{introduce acronyms the first time they're used}\todoquestion{This was introduced in 4.1.1, should it also be done here?} from scratch would be completely infeasible, but the goal was also not to simply purchase an off the shelf SBC that was not tailored to the task at hand such as a traditional Raspberry Pi.
+Specifically, building a Single Board Computer from scratch would be completely infeasible, but the goal was also not to simply purchase an off the shelf SBC that was not tailored to the task at hand such as a traditional Raspberry Pi.
 Instead, a better suited solution would be to utilize a \emph{Raspberry Pi Compute Module 4 (CM4)} which has \enquote{The power of Raspberry Pi 4 in a compact form factor for deeply embedded applications} \cite{rpi_cm4}.
@@
 Because the CM4 is designed to be as low-powered and compact as possible, some features that would usually be expected of such a device, like the USB ports, are disabled by default and must instead be built by the board designer.
-While this helps greatly by removing extraneous components and features that aren't needed for this project, it also provides me with a \todoquestion{does this sound sarcastic?}great opportunity to learn exactly what parts are being used in the hardware.
+While this helps greatly by removing extraneous components and features that aren't needed for this project, it also provides a great opportunity to learn exactly what parts are being used in the hardware.
 
@@
 For projects of this scale, the usual candidates for a relatively performant screen are HDMI screens or DSI screens.
-While HDMI screens are much more common and recognizable, DSI screens are a new development that promises high performance in low cost and low power applications such as embedded systems.
+While HDMI screens are much more common and recognizable, DSI screens are a new development that promises 60Hz performance in low cost and low power applications such as embedded systems.
 However, when looking into existing DSI screens, there are a couple of issues that prevent it from being the best choice for this project.
@@
 Though the PCB board would still have to be custom built and components would need to be replaced due to the current ongoing chip shortage, it provides a great foundation instead of starting from scratch.
-Another massive benefit from working within the Raspberry Pi Ecosystem is their \todoquestion{(re)move?}incredible documentation and community surrounding the boards.
-Around the time this project was being started, the CM4 was just starting to get into the hands of makers and developers, and only a couple custom boards had actually been made by anyone other than Raspberry Pi themselves.
+Another massive benefit from working within the Raspberry Pi Ecosystem is their incredible documentation and community surrounding the boards \cite{rpi_cm4_docs}.
+Around the time this project was being started, the CM4 was just beginning to get into the hands of makers and developers, and only a couple custom boards had actually been made by anyone other than Raspberry Pi themselves.
 Even though this leaves the community with less of a bulk of knowledge to work with, it has provided a sort of comradery with everyone helping each other learn the new Compute Module together.
-Members of the CM4 engineering team would even assist those on the forums to help them pick parts and design their own boards.
+Members of the CM4 engineering team would even assist those on the forums to help them pick parts and design their own boards \cite{cm4_datasheet_problem,rtc_kernelpatch}.
 
@@
 Thankfully due to most parts following international standards, many parts were chosen simply because they followed the same standard as the CM4 (such as USB 2.0, or \emph{1PORT 1000 BASE-T} Ethernet) and were in stock.
-What follows is a table of some components that had to be purchased apart from those that had been tested by the community, and the reason for their selection.
-A full Bill of Materials can be found in Appendix \ref{AppendixB}.
+What follows is a table of some components that were used in the custom PCB that may differ from those used by the CM4's official board, and the reason for their selection.
+A full Bill of Materials (BOM) can be found in Appendix \ref{AppendixB}.
 
@@
   Flat Flexible Cable Connector & \emph{FFC3B07-20-T} by \enquote{GCT} & A generic Flat Flexible Cable port that can be connected to anything.
-In this case, one is connected to the build-in screen's HDMI port, and another is connected to the screen's touch/power USB port.
+In this case, one is connected to the built-in screen's HDMI port, and another is connected to the screen's touch/power USB port.
 \\
   \hline
-  2-Pin Through Hole Header & \emph{PREC002SAAN-RC} by \enquote{Sullins Connector Solutions} & A generic 2-pin header than can be used to bypass the power circuit as discussed in Section \ref{sec:Prototyping}.
+  2-Pin Through Hole Header & \emph{PREC002SAAN-RC} by \enquote{Sullins Connector Solutions} & A generic 2-pin header than can be used to bypass the power circuit as will be discussed in Section \ref{sec:Prototyping}.
 \\
@@
   HDMI Power Regulator & \emph{RT9742ENGJ5} by \enquote{Richtek USA Inc.} & The CM4 recommends the \emph{RT9742SNGV} from the same manufacturer, but it is out of stock.
-This part has slight differences that call for tweaks in the final layout.
+This part has slight differences in functionality that require adjustments from the official I/O board design.
 \\
@@
   SD-Card Power Regulator & \emph{RT9742VGJ5} by \enquote{Richtek USA Inc.} & The CM4 recommends the \emph{RT9742GGJ5F} from the same manufacturer, but it is out of stock.
-This part has slight differences that call for tweaks in the final layout.
+This part has slight differences in functionality that require adjustments from the official I/O board design.
 \\
   \hline
-  Battery Voltage Regulator and Boost Converter & \emph{TPS61090RSAR} by \enquote{TPS61090RSAR} & Used to balance power usage from the USB-C port or the attached battery, favoring the external power when available.
+  Battery Voltage Regulator and Boost Converter & \emph{TPS61090RSAR} by \enquote{Texas Instruments} & Used to balance power usage from the USB-C port or the attached battery, favoring the external power when available.
 \\
@@
 \end{longtable}
-\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,inline]{Should the above table have pictures? If so, where? Since the table is already pretty dense}
 
@@
 
+Whenever working with electrical projects such as this, it is always a good idea to build up prototypes before committing to a final design that will be produced at the highest quality.
+Usually breadboards and some breakout boards for Surface Mounted Devices (SMD) are a great way to develop a circuit without creating a full PCB; but due to the nature of the complex microchips required for this project as well as the dependence on the CM4's tiny mezzanine connectors, building a prototype using a breadboard is impractical given the time and resources available.
 \begin{wrapfigure}{o}{0.37\textwidth}
@@
 \end{wrapfigure}
-Whenever working with electrical projects such as this, it is always a good idea to build up prototypes before committing to a final design that will be produced at the highest quality.
-Usually breadboards and some breakout boards for Surface Mounted Devices (SMD) are a great way to develop a circuit without creating a full PCB; but due to the nature of the complex microchips required for this project as well as the dependence on the CM4's tiny mezzanine connectors, building a prototype using a breadboard is impractical given the time and resources available.
 For example, the USB controller chip only has 0.5mm of space between each of it's 64 pins (Figure \ref{fig:USBController}), which would require a breakout board at least 7 times larger than the chip so that each each pin can be easily accessed by hand.
 Not to mention the difficulty in wiring up that many pins along with the other related components which would need their own breakout boards either purchased or custom-designed.
+
+Instead, the first board was built out in a piecewise fashion, allowing each section of the board a way to fail without interrupting the rest of the board.
+This allows each function of the board to be tested individually and in isolation so that changes could be made incrementally without needing to worry about their integration with the rest of the board.
+\todoquestion{This is that tricky sentence}For example, in order to build a system that would allow the board to be powered via USB or an internal battery, as well as charge the battery simultaneously, a complex circuit would have to be built that managed Lithium Polymer (LiPo) batteries which could cause fires if mishandled.
+Since power is obviously a critical component of the board, A secondary way of feeding power directly the board was built to bypass all that circuitry in case it fails.
+In doing so, the board can still be powered and it's other features can be tested while a fix for that portion of the circuitry is developed for the next iteration.
 
@@
 
-Instead, the first board was built out in a piecewise fashion, allowing each section of the board a way to fail without interrupting the rest of the board.
-This allows each function of the board to be tested individually and in isolation so that changes could be made incrementally without needing to worry about their integration with the rest of the board.
-\todoquestion{How do I convert this from first person to third person}For example, in order to build a system that would allow the board to be powered via USB or an internal battery, as well as charge the battery simultaneously, I would need to build a complex circuit consisting of parts neither I nor anyone I knew had used before.
-Since power is obviously a critical component of the board, A secondary way of feeding power directly the board was built to bypass all that circuitry in case it fails.
-In doing so, the board can still be powered and it's other features can be tested while a fix for that portion of the circuitry is developed for the next iteration.
-
 \section{Designing the Printed Circuit Board}\label{sec:DesigningThePCB}
 
-Because of the lack of prototyping options and the complexity of the circuit and it's components, the Printed Circuit Board was designed in a modular and piecewise fashion.
-This allowed any single part of the circuit to fail without interfering with the other components.
-Since the time and budget for this project was limited, the amount of revisioning was minimized by running the circuit by as many people as possible to catch any potential flaws before the board was manufactured.
+Because of the lack of prototyping options discussed in Section \ref{sec:Prototyping}, and ultimate decision to build the board in a way that could safely fail in any way, it became clear that multiple versions of the project were going to need to be designed and produced.
+Since the time and budget for this project was limited, the amount of revisioning was minimized by running the circuit by electrical engineering students and professors to catch any potential flaws before the board was manufactured.
 
-\begin{figure}[b]
+\subsection{Power}\label{subsec:DesigningPower}
+
+\begin{figure}[t]
   \centering
-  \includegraphics[width=0.33\textwidth]{Figures/kicad/close-ups/power-front}
+  \begin{subfigure}{.5\textwidth}
+    \centering
+    \includegraphics[width=.75\linewidth]{Figures/kicad/close-ups/power-front}
+  \end{subfigure}%
+  \begin{subfigure}{.5\textwidth}
+    \centering
+    \includegraphics[width=.75\linewidth]{Figures/pcb/crops/power}
+  \end{subfigure}
   \caption[PCB Power Circuit]{The front side of the PCB's power circuit}
   \label{fig:PowerCircuit}
+\end{figure}
+
+\begin{figure}[b!]
+  \centering
+  \includegraphics[width=.75\textwidth,page=2]{Figures/kicad/lathrum_thesis_schematic.pdf}
+  \caption[Power Schematic]{Schematic layout for the power circuit, including battery management. Expanded in Appendix \ref{app:pcb_schematic_battery}}
+  \label{fig:pcb_schematic_battery}
 \end{figure}
@@
 
+\subsection{USB}\label{subsec:DesigningUSB}
+
 \begin{figure}[t]
   \centering
-  \includegraphics[width=0.33\textwidth]{Figures/kicad/close-ups/usb-front}
+  \begin{subfigure}{.5\textwidth}
+    \centering
+    \includegraphics[width=.75\linewidth]{Figures/kicad/close-ups/usb-front}
+  \end{subfigure}%
+  \begin{subfigure}{.5\textwidth}
+    \centering
+    \includegraphics[width=.75\linewidth]{Figures/pcb/crops/usb}
+  \end{subfigure}
   \caption[PCB USB Circuit]{The front side of the PCB's USB circuit}
   \label{fig:USBCircuit}
+\end{figure}
+
+\begin{figure}[b!]
+  \centering
+  \includegraphics[width=.75\textwidth,page=4]{Figures/kicad/lathrum_thesis_schematic.pdf}
+  \caption[USB Schematic]{Schematic layout for the USB hub and ports. Expanded in Appendix \ref{app:pcb_schematic_usb}}
+  \label{fig:pcb_schematic_usb}
 \end{figure}
@@
 
-\begin{figure}[b]
+\subsection{HDMI}\label{subsec:DesigningHDMI}
+
+\begin{figure}[t]
   \centering
-  \begin{subfigure}{.5\textwidth}
+  \begin{subfigure}{.33\textwidth}
     \centering
-    \includegraphics[width=.6\linewidth]{Figures/kicad/close-ups/hdmi-front}
+    \includegraphics[width=1\linewidth]{Figures/kicad/close-ups/hdmi-front}
     \caption{The front side of the PCB}
@@
   \end{subfigure}%
-  \begin{subfigure}{.5\textwidth}
+  \begin{subfigure}{.33\textwidth}
     \centering
-    \includegraphics[width=.6\linewidth]{Figures/kicad/close-ups/hdmi-back}
+    \includegraphics[width=1\linewidth]{Figures/kicad/close-ups/hdmi-back}
     \caption{The back side of the PCB}
     \label{fig:HDMICircuitBack}
+  \end{subfigure}%
+  \begin{subfigure}{.33\textwidth}
+    \centering
+    \includegraphics[width=1\linewidth]{Figures/pcb/crops/hdmi}
+    \caption{The physical PCB}
+    \label{fig:HDMICircuitReal}
   \end{subfigure}
@@
   \label{fig:HDMICircuit}
+\end{figure}
+
+% While this is the image for the CM4 section, it makes sense to put it here
+\begin{figure}[b!]
+  \centering
+  \begin{subfigure}{.33\textwidth}
+    \centering
+    \includegraphics[width=.95\linewidth]{Figures/kicad/close-ups/cm4-front}
+    \caption{The front side of the PCB}
+    \label{fig:CM4CircuitFront}
+  \end{subfigure}%
+  \begin{subfigure}{.33\textwidth}
+    \centering
+    \includegraphics[width=.95\linewidth]{Figures/kicad/close-ups/cm4-back}
+    \caption{The back side of the PCB}
+    \label{fig:CM4CircuitBack}
+  \end{subfigure}%
+  \begin{subfigure}{.33\textwidth}
+    \centering
+    \includegraphics[width=.95\linewidth]{Figures/pcb/crops/cm4}
+    \caption{The physical PCB}
+    \label{fig:CM4CircuitReal}
+  \end{subfigure}
+  \caption{The circuitry surrounding the CM4's two mezzanine connectors}
+  \label{fig:CM4Circuit}
 \end{figure}
@@
 Even though their connectors are completely generic, meaning that they have no set pinout, it is possible to use them given that the receiving screen will take in the FFC in the same order as a HDMI cable.
-Using this knowledge, an internal FFC can be wired for the device's primary screen, while a second tradition HDMI port can be exposed to the user in case they want to use an external monitor.
+Using this knowledge, an internal FFC can be wired for the device's primary screen, while a second traditional HDMI port can be exposed to the user in case they want to use an external monitor.
 This, along with the FFC used for one of the four USB ports, allows a screen to be hooked up to the CM4 internally without needing to expose anything to the user.
+
+\subsection{CM4 and Miscellaneous}\label{subsec:DesigningCM4}
+
+\begin{figure}[t!]
+  \centering
+  \includegraphics[width=.75\textwidth,page=3]{Figures/kicad/lathrum_thesis_schematic.pdf}
+  \caption[USB Schematic]{Schematic layout for interfacing with the Compute Module 4. Expanded in Appendix \ref{app:pcb_schematic_cm4}}
+  \label{fig:pcb_schematic_cm4}
+\end{figure}
 
@@
 By removing the copper pour from this area on the PCB board and stitching the perimeter with vias, it prevents the board from acting like another antenna and interfering with the signals being sent to and from the CM4.
-Finally, each of the other components are connected to the CM4 board in the same manner as described previously in the chapter, such as the Ethernet adapter, the SD card reader, and remaining ports.
+Finally, each of the other components are connected to the CM4 board in the same manner as described throughout this chapter, such as the Ethernet adapter, the SD card reader, and remaining ports.
+\todoadvisor{Why are those components not described in the same way as power/USB/etc.?}\todoquestion{The Ethernet port is hooked directly into the CM4 connector with nothing interested except that they are also differential pairs, as described in the HDMI section.
+I suppose they could be explained fully, but I worry that it wouldn't add any useful information.}
 See Appendices \ref{AppendixA} and \ref{AppendixB} for further information.
-
-\begin{figure}[t]
-  \centering
-  \begin{subfigure}{.5\textwidth}
-    \centering
-    \includegraphics[width=.75\linewidth]{Figures/kicad/close-ups/cm4-front}
-    \caption{The front side of the PCB}
-    \label{fig:CM4CircuitFront}
-  \end{subfigure}%
-  \begin{subfigure}{.5\textwidth}
-    \centering
-    \includegraphics[width=.75\linewidth]{Figures/kicad/close-ups/cm4-back}
-    \caption{The back side of the PCB}
-    \label{fig:CM4CircuitBack}
-  \end{subfigure}
-  \caption{The circuitry surrounding the CM4's two mezzanine connectors}
-  \label{fig:CM4Circuit}
-\end{figure}
 
@@
 
-\begin{figure}[b]
+\begin{figure}[t]
   \centering
@@
 While ASU offers a basic printing service for use in classes and academic projects, the requirements of this board, specifically the tiny traces and pads required for the CM4's mezzanine connectors (pictured in Figure \ref{fig:rpi_cm4_mezzanine}), are outside of the printer's capabilities.
-Instead, by recommendation of a professor, the printing was done through JLCPC (\href{https://www.jlcpc.com/}{JLCPCB.com})\todoquestion{Would this be the right format for linking JLCPCB?
-It's not really a source or a reference that would be linked in the bibliography}, a company that provides quick turnaround and low costs for printing PCBs.
+Instead, by recommendation of a professor, the printing was done through an external company that provides quick turnaround and low costs for printing PCBs \cite{jlcpcb}.
 A set of five boards was ordered, allowing for multiple boards to be assembled in case of error, as well as a stencil.
@@
 This enables the easy application of solder paste onto all of the pads on the board as once, including the difficulty small pads of the CM4's mezzanine connectors.
-While the PCB was being printed, all of the parts found in the Bill of Materials (found in Appendix \ref{AppendixB}) were ordered from sites such as Digi-Key and Mouser.\todoquestion{Should I mention small issues such as the time when I ordered the PCB and before I could order the parts one of them went out of stock?
-I worry that that would break the flow of the paper, and I already mentioned stock issues in a previous section.
-I feel like it's a detail that makes an interesting story, but since this isn't a story paper it's not needed here.}
+While the PCB was being printed, all of the parts found in the Bill of Materials (found in Appendix \ref{AppendixB}) were ordered from sites such as Digi-Key and Mouser \cite{digikey,mouser}.
 In order to account for the production of multiple prototype boards and the possibility of broken parts, enough parts were ordered to produce five complete boards.
 
-Once the PCB and it's component parts arrived, it was time for assembly.
+\subsection{First Attempt}\label{subsec:Manufacturing1}
+
+Once the PCB and it's component parts arrived, the board could be assembled.
 The first board was assembled by hand using the PCB stencil and a handheld heat gun.
@@
 Once the parts were placed, a heat gun was used to melt the solder paste and fuse the components to the board.
-Following the SMDs, the through-hole components, such as the ports lining the side of the board, were then placed and soldered using a hand-held \todoquestion{soldering iron or solder iron?}soldering iron.
+Following the SMDs, the through-hole components, such as the ports lining the side of the board, were then placed and soldered using a hand-held soldering iron.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.4\textwidth]{Figures/pcb/handmade_mezzanine.JPG}
+  \caption[Hand-soldered mezzanine connector]{The leftovers of the broken-off mezzanine connector from the hand-soldered first board. Notice the exposed copper and warped pads.}
+  \label{fig:brokenmezzanine}
+\end{figure}
 
@@
 Firstly, due to improper use of the heat gun, most of the LEDs had popped, resulting in all hardware-level diagnostic indicators being inoperable.
-Secondly, the CM4 connectors were not properly connected to the board, and in an attempt to repair it, the connector components were irreparably damaged and the board itself was damaged slightly.\todo{Definitely want a picture here}
+Secondly, the CM4 connectors were not properly connected to the board, and in an attempt to repair it, the connector components were irreparably damaged and the board itself was damaged slightly, seen in Figure \ref{fig:brokenmezzanine}.
 Instead of attempting to assemble another board by hand, a local third-party company was contacted to assemble the board professionally.
 After delivering the board, stencil, and parts, the second board was assembled.
+
+\subsection{Second Attempt}\label{subsec:Manufacturing2}
 
@@
 Thankfully, as described in Section \ref{sec:Prototyping}, the secondary method of powering the board could be used to test the remaining features of the board.
-After making the necessary fixes, a second round of five PCB boards and a stencil was ordered from JLCPCB.
+After making the necessary fixes, a second round of five PCB boards and a stencil was ordered.
 Because there was still enough components to assemble three more boards, and no components needed to be changed, no more components were ordered at this stage.
 Once the boards were delivered, the revised board was again sent to be professionally assembled.
+
+\subsection{Third Attempt}\label{subsec:Manufacturing3}
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=0.8\textwidth]{Figures/pcb/crops/final}
+  \caption[Assembled PCB]{The assembled PCB}
+  \label{fig:AssembledPCB}
+\end{figure}
 

### 2022-03-31 paper/Chapters/Software.tex
Fixes from feedback, add images of physical board

@@
   \item The user can make a request to \emph{/applist} to get a list of applications that the host is able to stream.
-  \item The user can make a request to \emph{/launch} with a number of parameters to instruct the host to create a session with the given settings and launch the specified application.
+  \item The user makes a request to \emph{/launch} with a number of parameters to instruct the host to create a session with the given settings and launch the specified application.
         \begin{enumerate}
@@
   \item If the user unexpectedly or accidentally leaves a session without closing it, they can make a request to \emph{/resume} to instruct the host to reopen and resume the session the previous session.
-  \item Once the user is finished, that can make a request to \emph{/cancel} to instruct the host to close the session.
+  \item Once the user is finished, they make a request to \emph{/cancel} to instruct the host to close the session.
   \item If the user is completely finished with the host, they can return it to the state they found it in by making a request to \emph{/unpair} with their unique ID, unauthorizing themselves with the host.

### 2022-04-01 paper/Chapters/Software.tex
Start developing for ARM section

@@
 The project is structured as a core library written in \emph{C} with a number of community built clients built on top of it for various platforms.
-In essence, it turns the process described in the previous section (\ref{subsec:NVIDIAGameStream}) into a usable API.
-By reverse engineering the protocol originally implemented by the Nvidia Shield, Moonlight enables the development of a GameStream client for any other platform.
+In essence, it turns the process described in the previous section (\ref{subsec:NVIDIAGameStream}) into a usable programming interface.
+By reverse engineering the protocol originally implemented by the Nvidia Shield, Moonlight enables the development of a GameStream client for almost any operating system.
 The Moonlight community has already developed fantastic clients for many platforms such as Windows, Mac, Linux, Android, iOS, Web, and even other gaming consoles.
@@
 
-\todosection
+The CM4 runs on a 64-bit ARM Cortex-A72 processor \cite{rpi_cm4}, meaning that applications must be compiled for the \emph{ARM} architecture in order for the CPU to be able to run it.
+If a developer were to write a program on their \emph{x86} architecture Windows desktop computer, they could compile and run their program on any other Windows computer running on the same \emph{x86} architecture.
+They could also compile their application for other operating systems, but the target machine must also be running the same architecture.
+A computer running on another architecture such as \emph{ARM} is not able to run a program compiled for \emph{x86} machines, and a computer running on \emph{x86} is not able to run a program compiled for \emph{ARM} machines or any other architecture.
+
+In order to run the code for the project, given that the CM4 is the only \emph{ARM} computer available for this project, the code must either be compiled on the CM4 or built in a cross-compilation environment on the \emph{x86} machine.
+Using a cross-compilation environment would be overall quicker than that developing on the CM4 itself since it can use the full power of a desktop PC.
+Thankfully, setting up a cross-compilation environment for projects written in \emph{C} is straightforward.
+\emph{CMAKE}, a common compilation tool for \emph{C}, has built-in support for compiling for different target architectures with easy to use libraries.
+What isn't as straightforward however, is building a toolchain to compile the rest of the code for \emph{ARM}.
+
+In order to create the most straightforward experience possible for someone using the hardware for this project, a custom operating system image is created that has all the dependencies, libraries, and applications pre-installed.
+Manjaro provides good tools for building and managing custom images of the Manjaro operating system for \emph{ARM} architectures \cite{manjaro_arm_tools}.
+The only drawback is that it must be run an \emph{ARM} processor, and is much more difficult to run from another architecture.
+To make this process easier to develop, a toolchain was developed that could be run on the CM4 to automatically compile all the necessary code using \emph{cmake}, package all the necessary files using \emph{makepkg}, and build a fresh OS image that is ready to flash.
+This three step script allows the developer to pull the git repository of the project, run a set up script, and then run the build script to compile everything into an immediately deployable OS image.
+The following excerpt details the last step of the process, building the OS image that includes the custom application.
+
+\begin{lstlisting}[style=custombash,firstnumber=72,title=Excerpt from \emph{/code/build.sh}]
+pkgname=$(sed -n 's/^pkgname=//p' ./pkgbuild/PKGBUILD)
+pkgver=$(sed -n 's/^pkgver=//p' ./pkgbuild/PKGBUILD)
+pkgrel=$(sed -n 's/^pkgrel=//p' ./pkgbuild/PKGBUILD)
+# Copy local package to cache to be installed into new image
+sudo cp ./pkgbuild/$pkgname-$pkgver-$pkgrel-aarch64.pkg.tar.zst /var/cache/pacman/pkg/
+# Build image
+sudo buildarmimg -d rpi4 -e thesis -i $pkgname-$pkgver-$pkgrel-aarch64.pkg.tar.zst
+read -s -n 1 -p "Image built, press a key to continue (for sudo)"
+# Remove local package from cache
+sudo rm /var/cache/pacman/pkg/$pkgname-$pkgver-$pkgrel-aarch64.pkg.tar.zst -f
+# Move build image to build folder
+sudo mv /var/cache/manjaro-arm-tools/img/* ./build/
+\end{lstlisting}
+
+This script pulls the package's name, version, and release number from the \emph{PKGBUILD} file generated from the packaging step and installs it into the newly created image, which is then moved into the project's build directory.
+This image file is then ready to be flashed and installed like any other operating system.
 
@@
 
-\todosection
+In order to evaluate the performance of the project, a suite of tools must be developed to test both the newly developed application and the existing solutions.
+To do this, two sets of data will be collected: First, data directly from the application that logs statistics such as the number of frames rendered, dropped, or rendered, and data that runs separately from the application that logs the amount of time it takes for data that is from the client to be realized as a change on the screen.
+The first set of data is important to track the raw performance of the application, statistics like Frames Per Second (FPS)
+
+\begin{lstlisting}[style=customc,firstnumber=23,title=Excerpt from \emph{/src/video/stats.h}]
+typedef struct _VIDEO_STATS {
+  uint32_t receivedFrames;
+  uint32_t decodedFrames;
+  uint32_t renderedFrames;
+  uint32_t totalFrames;
+  uint32_t networkDroppedFrames;
+  uint32_t totalDecodeTime;
+  uint32_t totalRenderTime;
+  uint32_t lastRtt; // Rtt = Round Trip Time
+  uint32_t lastRttVariance;
+  float receivedFps;
+  float decodedFps;
+  float renderedFps;
+  uint32_t measurementStartTimestamp;
+} VIDEO_STATS;
+\end{lstlisting}

### 2022-04-01 paper/main.tex
(almost) finish draft of hardware section
Still need to investigate the USB issues and maybe add more detail to the end of the chapter, but other than that I have a couple of questions to resolve then I'm happyish with the chapter
Start the software section
Fixes from feedback, add images of physical board
Start developing for ARM section

@@
 \usepackage{longtable} % Tables that span multiple pages
+
+\usepackage{listings} % Required for the code examples
+% Code block settings
+\lstdefinelanguage{custombash}{
+	morekeywords={sudo,cp,rm,mv,read,buildarmimg},
+	morestring=[b]{"},
+	morestring=[b]{'},
+	morecomment=[l]{\#},
+	morekeywords=[2]{pkgname,pkgver,pkgrel,\$pkgname,\$pkgver,\$pkgrel},
+}
+\lstdefinestyle{custombash}{
+  language=custombash,
+  belowcaptionskip=1\baselineskip,
+  breaklines=true,
+  frame=single,
+	numbers=left,
+  % xleftmargin=\parindent,
+  showstringspaces=false,
+  basicstyle=\footnotesize\ttfamily,
+  commentstyle=\itshape\color{green!40!black},
+  keywordstyle=\bfseries\color{orange!70!black},
+  keywordstyle=[2]\bfseries\color{blue!90!black},
+  % identifierstyle=\color{blue},
+  stringstyle=\color{orange},
+}
+\lstdefinestyle{customc}{
+  language=C,
+	morekeywords={uint32_t},
+  belowcaptionskip=1\baselineskip,
+  breaklines=true,
+  frame=single,
+  % xleftmargin=\parindent,
+	numbers=left,
+  showstringspaces=false,
+  basicstyle=\footnotesize\ttfamily,
+  keywordstyle=\bfseries\color{blue!80!black},
+  commentstyle=\itshape\color{green!40!black},
+  % identifierstyle=\color{blue},
+  stringstyle=\color{orange},
+}
+\lstset{escapechar=@,style=customc}
 
@@
 	\addchaptertocentry{\abbrevname} % Add the Abbreviations to the table of contents
+	\textbf{BOM} & \textbf{B}ill \textbf{O}f \textbf{M}aterials\\
 	\textbf{CM4} & Raspberry Pi \textbf{C}ompute \textbf{M}odule \textbf{4}\\
@@
 	\textbf{FFC} & \textbf{F}lat \textbf{F}lexible \textbf{C}able\\
+	\textbf{FPS} & \textbf{F}rames \textbf{P}er \textbf{S}econd\\
 	\textbf{GPU} & \textbf{G}raphics \textbf{P}rocessing \textbf{U}nit\\
@@
 	\textbf{RDP} & \textbf{R}emote \textbf{D}esktop \textbf{P}rotocol\\
+	\textbf{RTC} & \textbf{R}eal \textbf{T}ime \textbf{C}lock\\
 	\textbf{SBC} & \textbf{S}ingle \textbf{B}oard \textbf{C}omputer\\

### 2022-04-01 paper/Chapters/Software.tex
Finish draft of software section

--- 2022-03-31 paper/Chapters/Software.tex
@@
 
-\todosection
+The second half of the project is composed of software that is used to connect the client machine to a host computer.
+Designing the software begins with identifying it's requirements in Section \ref{sec:SoftwareRequirements}.
+Then, potential avenues for development are discussed in Section \ref{sec:PotentialAvenues} so that the software does not need to be built from scratch.
+Following that, special considerations must be made while developing the software for the CM4 running on an \emph{ARM} processor in Section \ref{sec:DevelopingForARM}, and the development of the suite of tools used to record data is detailed in Section \ref{sec:DevelopingTestingAndMeasurementTools}.
 
@@
 The project is structured as a core library written in \emph{C} with a number of community built clients built on top of it for various platforms.
-In essence, it turns the process described in the previous section (\ref{subsec:NVIDIAGameStream}) into a usable API.
-By reverse engineering the protocol originally implemented by the Nvidia Shield, Moonlight enables the development of a GameStream client for any other platform.
+In essence, it turns the process described in the previous section (\ref{subsec:NVIDIAGameStream}) into a usable programming interface.
+By reverse engineering the protocol originally implemented by the Nvidia Shield, Moonlight enables the development of a GameStream client for almost any operating system.
 The Moonlight community has already developed fantastic clients for many platforms such as Windows, Mac, Linux, Android, iOS, Web, and even other gaming consoles.
-All that is needed for the purposes of this project is to make tweaks to best suit the CM4 that drives the hardware and develop a suite of tools to test it's functionality.
+All that is needed for the purposes of this project is to make a fork of the project that tweaks it to best suit the CM4 that drives the hardware and develop a suite of tools to test it's functionality.
 
@@
 
-\todosection
+The CM4 runs on a 64-bit ARM Cortex-A72 processor \cite{rpi_cm4}, meaning that applications must be compiled for the \emph{ARM} architecture in order for the CPU to be able to run it.
+If a developer were to write a program on their \emph{x86} architecture Windows desktop computer, they could compile their program for any other computer running on the same \emph{x86} architecture, but such a program would be incompatible with an \emph{ARM} processor.
+Similarly, if the program were compiled for \emph{ARM}, that program could be run on an \emph{ARM} computer but not one running \emph{x86}.
+
+In order to run the code for the project, given that the CM4 is the only \emph{ARM} computer available for this project, the code must either be compiled on the CM4 or built in a cross-compilation environment on an \emph{x86} machine.
+Using a cross-compilation environment would be overall quicker than that developing on the CM4 itself since it can use the full power of a desktop PC.
+Thankfully, setting up a cross-compilation environment for projects written in \emph{C} is straightforward.
+\emph{CMake}, a common compilation tool for \emph{C}, has built-in support for compiling for different target architectures with easy to use libraries.
+What isn't as straightforward however, is building a toolchain to compile the rest of the code for \emph{ARM}.
+
+In order to create an out-of-the-box experience for someone using the hardware for this project, a custom operating system image is created that has all the dependencies, libraries, and applications pre-installed and pre-configured.
+Manjaro provides good tools for building and managing custom images of the Manjaro operating system for \emph{ARM} architectures \cite{manjaro_arm_tools}, with the only drawback being that it is that it must be run on an \emph{ARM} processor, and is much more difficult to run on another architecture.
+To make development easier, a toolchain was built that could be run on the CM4 to build the code that is written on other computers.
+All the developer needs to do is to pull the git repository of the project, run a set up script to install dependencies, and then run the build script to compile everything into an immediately deployable OS image.
+This build script will automatically compile all the necessary code using \emph{cmake}, package all the necessary files using \emph{makepkg}, and build a fresh OS image that is ready to flash.
+The following excerpt details the three functions used to make this work.
+
+\begin{lstlisting}[style=custombash,firstnumber=50,title=Excerpt from \emph{/scripts/build.sh}]
+runMake()
+{
+  mkdir ./moonlight/build # Create build directory
+  cd ./moonlight/build # Go to build directory
+  cmake ../ # Run cmake
+  make # Build the project
+  sudo make install # Install the project
+  sudo ldconfig /usr/local/lib # Link libraries
+  cd "$parent_path" # Go back to original execution location
+}
+
+runPackage()
+{
+  cd ./pkgbuild # Go to pkgbuild directory
+  tar -cf moonlight.tar ../moonlight/ # Create tar archive of the project
+  makepkg -sLfc # Build the package
+  rm moonlight.tar # Remove the tar archive
+  cd "$parent_path" # Go back to original execution location
+}
+
+runBuild()
+{
+  pkgname=$(sed -n 's/^pkgname=//p' ./pkgbuild/PKGBUILD)
+  pkgver=$(sed -n 's/^pkgver=//p' ./pkgbuild/PKGBUILD)
+  pkgrel=$(sed -n 's/^pkgrel=//p' ./pkgbuild/PKGBUILD)
+  # Copy local package to cache to be installed into new image
+  sudo cp ./pkgbuild/$pkgname-$pkgver-$pkgrel-aarch64.pkg.tar.zst /var/cache/pacman/pkg/
+  # Build image
+  sudo buildarmimg -d rpi4 -e thesis -i $pkgname-$pkgver-$pkgrel-aarch64.pkg.tar.zst
+  read -s -n 1 -p "Image built, press a key to continue (for sudo)"
+  # Remove local package from cache
+  sudo rm /var/cache/pacman/pkg/$pkgname-$pkgver-$pkgrel-aarch64.pkg.tar.zst -f
+  # Move build image to build folder
+  sudo mv /var/cache/manjaro-arm-tools/img/* ./build/
+  cd "$parent_path" # Go back to original execution location
+}
+\end{lstlisting}
+
+This script starts by building the forked version of moonlight using \emph{cmake} and \emph{make} and links it's libraries using \emph{ldconfig}.
+Once the application is built, it then builds it into an installable package using \emph{makepkg}, before installing it into the newly created image.
+Finally, the image is moved to the project's build folder where it is ready to be flashed and installed like any other operating system.
 
@@
 
-\todosection
+In order to evaluate the performance of the project, a suite of tools must be developed to test both the newly developed application and the existing solutions.
+To do this, two sets of data will be collected: First, data is recorded directly from the forked version of moonlight that logs statistics such as the number of frames received, dropped, or rendered, and data that is recorded from a separate application that logs the amount of time it takes for the client machine to receive updates after it sends data to the host machine.
+The first set of data is important to track the raw performance of the application, with statistics like Frames Per Second (FPS) and decoding time which are difficult to record otherwise.
+These statistics provide insight into how performant the application is from a technical perspective.
+The following is a structure that holds the statistics recorded over a session of connecting to a host computer.
+The data recorded is discussed in Chapter \ref{Chapter6}.
+
+\begin{lstlisting}[style=customc,firstnumber=23,title=Excerpt from \emph{moonlight/src/video/stats.h}]
+typedef struct _VIDEO_STATS {
+  uint32_t receivedFrames;
+  uint32_t decodedFrames;
+  uint32_t renderedFrames;
+  uint32_t totalFrames;
+  uint32_t networkDroppedFrames;
+  uint32_t totalDecodeTime;
+  uint32_t totalRenderTime;
+  uint32_t lastRtt; // Rtt = Round Trip Time
+  uint32_t lastRttVariance;
+  float receivedFps;
+  float decodedFps;
+  float renderedFps;
+  uint32_t measurementStartTimestamp;
+} VIDEO_STATS;
+\end{lstlisting}
+
+The second set of data is recorded by measuring the total amount of time it takes for the host to receive data from the client, and for the client to receive updates back from the host.
+This is done using the following steps once the client is connected to the host:
+
+\begin{enumerate}
+  \item When a user hits a key on the client, record which key was pressed and at what timestamp.
+  \item When the host receives the instruction to press the key from the client, record which key and the timestamp.
+  \item Change the color of the host's screen to a different color.
+  \item When the client detects that the color of the screen has changed, record the current timestamp.
+\end{enumerate}
+
+This records two points of data, the time it takes for the host to implement the actions done by the user, and the time that the user sees the result of their actions.
+Because this method of data collection runs as a separate application on both the client and the server, it can be used to measure the performance of any of the remote desktop applications for comparison.
+This is again discussed in Chapter \ref{Chapter6}.
+
+Both the application running on the client and the server do the same thing, every time a key is pressed, record which key what pressed and at what time, and every time the pixel in the center of the screen changes color, record the new color and at what time it changed.
+This data is written to a Comma Separated Values (CSV) file for later processing.
+To change the color of the screen, a simple web page is used that changes color any time a key is pressed.
+It's source in it's entirety follows:
+
+\begin{lstlisting}[style=customhtml,title=\emph{/datalogger/web/index.html}]
+<!DOCTYPE html>
+<html lang="en">
+  <head>
+    <script>
+      let nextColor = "green";
+      document.addEventListener("keydown", () => {
+        document.body.setAttribute("style", "background-color:" + nextColor);
+        nextColor = nextColor === "green" ? "red" : "green";
+      });
+    </script>
+  </head>
+  <body></body>
+</html>
+\end{lstlisting}
+
+This is sufficient to change the color of the screen back and forth between easily distinguishable colors for both the client and the host to record.
+With both of these datasets recorded, the next step is to analyze the data and evaluate the performance of the application.

### 2022-04-01 paper/main.tex
Finish draft of software section

@@
 \lstdefinelanguage{custombash}{
-	morekeywords={sudo,cp,rm,mv,read,buildarmimg},
+	morekeywords={sudo,cp,cd,rm,mv,read,mkdir,cmake,make,makepkg,buildarmimg},
 	morestring=[b]{"},
@@
 	morekeywords=[2]{pkgname,pkgver,pkgrel,\$pkgname,\$pkgver,\$pkgrel},
+}
+\lstdefinelanguage{customhtml}{
+	morekeywords={DOCTYPE,html,lang,head,script,body,let},
+	morekeywords=[2]{nextColor,document,body},
+	morekeywords=[3]{addEventListener,setAttribute},
+	morestring=[b]{"},
+	morestring=[b]{'},
+	morecomment=[l]{\#},
 }
@@
   keywordstyle=[2]\bfseries\color{blue!90!black},
+  % identifierstyle=\color{blue},
+  stringstyle=\color{orange},
+}
+\lstdefinestyle{customhtml}{
+  language=customhtml,
+	morekeywords={uint32_t},
+  belowcaptionskip=1\baselineskip,
+  breaklines=true,
+  frame=single,
+  % xleftmargin=\parindent,
+	numbers=left,
+  showstringspaces=false,
+  basicstyle=\footnotesize\ttfamily,
+  keywordstyle=\bfseries\color{blue!80!black},
+  keywordstyle=[2]\bfseries\color{teal!80!white},
+  keywordstyle=[3]\bfseries\color{orange!70!black},
+  commentstyle=\itshape\color{green!40!black},
   % identifierstyle=\color{blue},
@@
 	\textbf{CRD} & \textbf{C}hrome \textbf{R}emote \textbf{D}esktop\\
+	\textbf{CSV} & \textbf{C}omma \textbf{S}eparated \textbf{V}alues\\
 	\textbf{FFC} & \textbf{F}lat \textbf{F}lexible \textbf{C}able\\

### 2022-04-03 paper/Chapters/Evaluation.tex
Start Evaluation section, add graphs

@@
 
-\todosection
+In order to ensure that the data collected for analysis is accurate and directly comparable, the testing environment must be kept exactly the same between trials and between different applications.
+This gives each application the exact same starting conditions, and the best chance to demonstrate it's unabated performance.
+For these tests, the host machine is a desktop computer running Windows 10 with a AMD Ryzen 9 5900X CPU and a Nvidia RTX 3070 GPU running the latest GeForce Experience version \emph{3.25.0.84}.
+This configuration does not change throughout testing.
+The client machine is the CM4 board attached to the PCB developed in Chapter \ref{Chapter4}, running the custom Manjaro linux distribution and forked version of Moonlight developed in Chapter \ref{Chapter5}.
+This configuration also does not change throughout testing.
+Each test is performed with both the host computer and the client machine running on the same network, with both devices connected to the same router using an ethernet cable.
+This ensures that the speed of the network is not a factor in the tests.
+The tests are performed multiple times under the following conditions:
+
+\begin{itemize}
+  \item Ideal Conditions: No other applications are running during testing.
+  \item CPU Stressed: The CPU is constantly at 100\% load during the entire test.
+  \item GPU Stressed: The GPU is constantly at 100\% load during the entire test.
+\end{itemize}
+
+\noindent
+For each test, the following steps are performed:
+
+\begin{enumerate}
+  \item Both the host and client computers are fully rebooted to ensure no other application is running.
+  \item The host computer starts up a single browser window with the web page described in Section \ref{sec:DevelopingTestingAndMeasurementTools}.
+  \item If the test calls for the CPU or the GPU to be stressed, the application \enquote{Blender} is started on the host machine, and a benchmark is performed for the CPU or the GPU.
+        This benchmark will run for longer than the duration of the test, and will keep the CPU or GPU at 100\% load for the entire duration of the benchmark.
+  \item Both devices start up their respective data collection tool that will record key-strokes and changes in the screen's color.
+  \item The client machine connects to the host computer using the application being tested.
+  \item The user presses the \enquote{0} key on the client's numpad to synchronize the two devices' data recording tools.
+  \item The user begins by typing on the client's keyboard slowly, allowing both computers to record their data.
+  \item The user then types on the client's keyboard quickly to ensure that quick successive inputs are handled correctly and no data is lost or duplicated.
+  \item Once typing is complete, both the client and the host computers close their data recording tools and the client ends the streaming session.
+\end{enumerate}
+
+\noindent
+This process is repeated for each of the following applications:
+
+\begin{itemize}
+  \item The forked version of Moonlight developed for this project.
+  \item Chrome Remote Desktop (CRD, Section \ref{subsec:ChromeRemoteDesktop}).
+  \item Microsoft's Remote Desktop Connection (RDP, Section \ref{subsec:RemoteDesktopProtocol}).
+  \item Virtual Network Computing (VNC, Section \ref{subsec:VirtualNetworkComputing}).
+\end{itemize}
+
+\noindent
+Once all the data is collected, the data from the host and client machines for a single trial are organized into a single CSV file for processing.
+In order to ensure that the data is comparable, all extraneous factors must be removed from the data.
+Namely, the network latency and difference in computer clocks must be removed.
+This is partially accounted for in step \emph{6} of the above process, and the remaining inaccuracy can now be removed by subtracting the time difference between the when the client presses the button and the time the host executes that command.
+This ensures that any inaccuracies in each computer's internal clock will not impact the data.
+
+After the data is cleaned up, the data can be processed by matching each key-stroke and change in the screen's color detected by the client and host machines and calculating how much time elapsed between the events.
+This reveals the amount of delay the user experiences between when they press a key and when the host executes the command, and between when they press a key and can visually see the response.
 
@@
 \todosection
+
+\begin{figure}[h]
+  \centering
+  \begin{tikzpicture}
+    \begin{axis}[
+      boxplot/draw direction=y,
+      ylabel={Time (ms)},
+      scale only axis,
+      height=8cm,
+      width=\textwidth,
+      ymin=0,ymax=200,
+      cycle list={{black},{blue},{red}},
+      legend style={
+          legend pos=outer north east,
+          font=\small
+        },
+      legend cell align=left,
+      legend image code/.code={%
+          \draw[#1] (0cm,-0.1cm) rectangle (0.6cm,0.1cm);
+        },
+      boxplot={
+          % Three boxplots for each column
+          draw position={1/4 + floor(\plotnumofactualtype/3) + 1/4*mod(\plotnumofactualtype,3)},
+          % Each plot takes up a quarter of the column
+          box extend=0.2,
+        },
+      % 1 unit in x controls the width:
+      x=2cm,
+      % ... and it means that we should describe intervals:
+      xtick={0,1,2,...,10},
+      x tick label as interval,
+      xticklabels={%
+      {Forked Moonlight\\{\tiny ideal/cpu/gpu}},%
+      {CRD\\{\tiny ideal/cpu/gpu}},%
+      {RDP\\{\tiny ideal/cpu/gpu}},%
+      {VNC\\{\tiny ideal/cpu/gpu}},%
+      },
+      x tick label style={
+          text width=2.5cm,
+          align=center
+        },
+      ]
+
+      % Forked Moonlight
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/ideal/moonlight_key_delay.csv};
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/stresscpu/moonlight_key_delay.csv};
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/stressgpu/moonlight_key_delay.csv};
+
+      % CRD
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/ideal/crd_key_delay.csv};
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/stresscpu/crd_key_delay.csv};
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/stressgpu/crd_key_delay.csv};
+
+      % RDP
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/ideal/rdp_key_delay.csv};
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/stresscpu/rdp_key_delay.csv};
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/stressgpu/rdp_key_delay.csv};
+
+      % VNC
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/ideal/vnc_key_delay.csv};
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/stresscpu/vnc_key_delay.csv};
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/stressgpu/vnc_key_delay.csv};
+
+      \legend{Ideal,CPU stressed,GPU stressed}
+    \end{axis}
+  \end{tikzpicture}
+  \caption[Input Delay Data]{Time for host to press key from client.}
+  \label{fig:InputDelay}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \begin{tikzpicture}
+    \begin{axis}[
+      boxplot/draw direction=y,
+      ylabel={Time (ms)},
+      scale only axis,
+      height=8cm,
+      width=\textwidth,
+      ymin=0,ymax=200,
+      cycle list={{black},{blue},{red}},
+      legend style={
+          legend pos=outer north east,
+          font=\small
+        },
+      legend cell align=left,
+      legend image code/.code={%
+          \draw[#1] (0cm,-0.1cm) rectangle (0.6cm,0.1cm);
+        },
+      boxplot={
+          % Three boxplots for each column
+          draw position={1/4 + floor(\plotnumofactualtype/3) + 1/4*mod(\plotnumofactualtype,3)},
+          % Each plot takes up a quarter of the column
+          box extend=0.2,
+        },
+      % 1 unit in x controls the width:
+      x=2cm,
+      % ... and it means that we should describe intervals:
+      xtick={0,1,2,...,10},
+      x tick label as interval,
+      xticklabels={%
+      {Forked Moonlight\\{\tiny ideal/cpu/gpu}},%
+      {CRD\\{\tiny ideal/cpu/gpu}},%
+      {RDP\\{\tiny ideal/cpu/gpu}},%
+      {VNC\\{\tiny ideal/cpu/gpu}},%
+      },
+      x tick label style={
+          text width=2.5cm,
+          align=center
+        },
+      ]
+
+      % Forked Moonlight
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/ideal/moonlight_color_delay.csv};
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/stresscpu/moonlight_color_delay.csv};
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/stressgpu/moonlight_color_delay.csv};
+
+      % CRD
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/ideal/crd_color_delay.csv};
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/stresscpu/crd_color_delay.csv};
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/stressgpu/crd_color_delay.csv};
+
+      % RDP
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/ideal/rdp_color_delay.csv};
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/stresscpu/rdp_color_delay.csv};
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/stressgpu/rdp_color_delay.csv};
+
+      % VNC
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/ideal/vnc_color_delay.csv};
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/stresscpu/vnc_color_delay.csv};
+      \addplot
+      table[col sep=comma,x=Time difference] {Data/stressgpu/vnc_color_delay.csv};
+
+      \legend{Ideal,CPU stressed,GPU stressed}
+    \end{axis}
+  \end{tikzpicture}
+  \caption[Color Delay Data]{Time for client to receive update to screen.}
+  \label{fig:ColorDelay}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+  \begin{tikzpicture}
+    \begin{axis}[
+      ylabel={Color changes not rendered (\%)},
+      scale only axis,
+      height=8cm,
+      width=\textwidth,
+      enlarge x limits=0.25,
+      ybar=2*\pgflinewidth,
+      bar width=14pt,
+      ymin=0,ymax=100,
+      cycle list={{black},{blue},{red}},
+      legend style={
+          legend pos=outer north east,
+          font=\small
+        },
+      legend cell align=left,
+      legend image code/.code={%
+          \draw[#1] (0cm,-0.1cm) rectangle (0.6cm,0.1cm);
+        },
+      % 1 unit in x controls the width:
+      x=2cm,
+      % ... and it means that we should describe intervals:
+      xtick={0,1,2,...,10},
+      % x tick label as interval,
+      xticklabels={%
+      {Forked Moonlight\\{\tiny ideal/cpu/gpu}},%
+      {CRD\\{\tiny ideal/cpu/gpu}},%
+      {RDP\\{\tiny ideal/cpu/gpu}},%
+      {VNC\\{\tiny ideal/cpu/gpu}},%
+      },
+      x tick label style={
+          text width=2.5cm,
+          align=center
+        },
+      ]
+      \addplot[style={black,fill=black,mark=none}]
+      coordinates {(0, 0) (1,0) (2,0) (3,15.32)};
+
+      \addplot[style={blue,fill=blue,mark=none}]
+      coordinates {(0,0) (1,32.28) (2,0) (3,19.51)};
+
+      \addplot[style={red,fill=red,mark=none}]
+      coordinates {(0,0) (1,2.34) (2,0) (3,15.5)};
+
+      \legend{Ideal,CPU stressed,GPU stressed}
+    \end{axis}
+  \end{tikzpicture}
+  \caption[Color Delay Loss Data]{Percentage of color changes not rendered.}
+  \label{fig:ColorDelayLoss}
+\end{figure}
 

### 2022-04-04 paper/Chapters/Evaluation.tex
Continue evaluation chapter

@@
 \noindent
+This is to examine whether each application can perform under intensive conditions.
 For each test, the following steps are performed:
@@
 
-\todosection
+The first factor in determining the performance of the application is to take a look at how responsive it is to use.
+Responsivity is defined as how quick the application responds to inputs and how quickly the user can see the results of their inputs.
+This is important for general user experience and applications where continuous input is required.
+This is also what is generally refered to when determining whether an application has a positive user experience, since a program that can take seconds to respond to the user begets lower productivity and a more frustrating experience.
+To analyse this, two points of data are recorded: First, the amount of time between when the client records a keystroke and when it is actually inputted on the host machine, and seccondly, the amount of time between the host machine updating their screen and when the client actually displays that update.
+The first datum is generally referred to as the latency of the application, since even though each trial is done using the same network setup different programs may have different optimizations built in that reduce the time needed for the host to implement keystrokes made by the client.
+Figure \ref{fig:InputDelay} shows the amount of time between the client and the host registering the keystroke under the different test conditions.
 
@@
 
+Each application does a good job sending data from the client to the host in a timely manner, with the majority of keystrokes being realized within a tenth of a seccond of being pressed on the client machine.
+The forked Moonlight implementation keeps a latency under 20ms regardless of the load of the host machine, with more than 75\% of all keystrokes being realized within 10ms.
+CRD proves to be a CPU limited application, with latency while idle or under GPU load also staying below 20ms, but climbing up above 50ms when under CPU load.
+This is important to keep in mind when thinking about what sort of situations the host machine will be in while being accessed.
+If the user is going to be rendering video remotely, a GPU intensive process, the latency of the stream wont be impacted, but if the user is going to be running a CPU intensive process, the latency can be expected to be negatively impacted.
+RDP has a similar story, with extremely low latency while idle or under GPU load, and upwards of 40ms while under CPU load.
+VNC has the highest variance, with no discernable affinity to any kind of load.
+This lines up with it's intended use described in Section \ref{subsec:VirtualNetworkComputing}, being the simplest and most straightforward method of remote desktop control without much focus on performance or security.
+That's not so say VNC is unsuable or even difficult to use; latency of 160ms is still usable in all but the most precise situations.
+Though this is only one metric to consider, the other is the amount of time it take for the client to display the results of their input.
+This is recorded as the amount of time it takes between the host machine changing the color of it's screen and the client rendering the change.
+Figure \ref{fig:ColorDelay} shows the amount of time for each application under each condition.
+
 \begin{figure}[h]
@@
       width=\textwidth,
-      ymin=0,ymax=200,
+      ymin=0,ymax=250,
       cycle list={{black},{blue},{red}},
@@
   \end{tikzpicture}
-  \caption[Color Delay Data]{Time for client to receive update to screen.}
+  \caption[Color Delay Data]{Time for client to render update to the screen.}
   \label{fig:ColorDelay}
 \end{figure}
+
+This data demonstrates a similar but more extreme pattern to the input delay figure, with most application performing well under ideal conditions but struggling a little under high stress.
+The forked Moonlight implementation is able to render updates on the client device in under 70ms in nearly all cases, with an even spread regardless of the load of the host machine.
+CRD is able to render all updates in under 160ms while the host is idle, but struggles to keep up while under load.
+Again, CRD struggles the most while under CPU load, with the quickest update happening in just under 200ms, and the median update taking over a full second.
+CRD performs a little better under GPU load, but still experiences much high delays with the median update taking over half of a second.
+Both of these values push the usability of the system as any kind of intensive program will bring the responsivity of the application to a crawl.
+RDP again performs extremely well, with all render updates taking less than 50ms while under ideal conditions or GPU load, and all updates taking less that 70ms while under full CPU load.
+VNC is again the slowest and most inconsistent application, with all conditions resulting in a frequent render delays of more than a quarter of a second.
+This makes VNC considerably less suitable for the task at hand.
+
+The delay, however, is only half the story in this case.
+Another factor to consider is the percentage of updates that are not rendered by the client. 
+This, often referred to as dropped frames, records how much data is lost when the host sends it's screen to the client device.
+A higher percentage results in a more choppy and difficult to use experience, with a 50\% loss meaning that 50\% of the updates sent by the host are not rendered by the client.
+Figure \ref{fig:ColorDelayLoss} shows the percentage of color changes not rendered by the client for each of the load conditions.
 
@@
 
-\section{Performance}
-
-\todosection
+These data points adds context to the previous figures, showing where the applications struggle to keep up with sending data back to the client device.
+The forked Moonlight implementation does not experience any frame drops, as does CRD under ideal conditions.
+But once the host machine is put under heavy load, CRD begins to drop frames, with a few being dropped under GPU stress, and 32\% of frames being dropped while under CPU stress.
+This can make CRD unpleasent and frustrating to use while the host is running a CPU intensive program as not only does the client take much more time to update the screen to reflect that status of the host computer, but many of the updates are lost completely and not shown to the user.
+RDP is also able to display all the data sent by the host without dropping any frames under any condition.
+VNC is again proved to be the least performant, with frames being dropped regardless of the load being put on the host computer.
+This is still in line with it's intended use of basic remote control, though it is dissapointing to see about 15\% loss even under ideal conditions.
 
@@
 
-\todosection
+Apart from the performance of each application, the quality of the user experience must also be taken into account when considering each application.
+When analyzing the forked version of Moonlight, specific quantitative data described in Section \ref{sec:DevelopingTestingAndMeasurementTools} can be collected directly from the application while it is running.
+\todo{put in data}
+
+Chrome Remote Desktop also provides some quantitative data.\todo{V8 compression}
+
+RDP does not provide quantitative to analyse it's quality, but speaking from a user experience perspective\dots
+
+VNC provides little in the form of quality statistics, but given it's performance and responsiveness discussed in the previous section, it is clear that the quality of the stream suffers heavily whenever the host is under load.
 
@@
 
+By combining the information from the previous two sections, a wholistic picture of the performance of each application can be drawn.
+
 \todosection

### 2022-04-04 paper/Chapters/Evaluation.tex
Finish draft of evaluation chapter

@@
 
-\section{Testing Methodology}
+In order to evaluate the performance of the forked version of Moonlight developed for this paper, it must be tested against the other existing solutions.
+The testing methodology employed to ensure the accuracy of the testing is described in Section \ref{sec:TestingMethodology}.
+Then, the different results are evaluated and analyzed in Sections \ref{sec:ResponsivityAndLatency} and \ref{sec:Quality}.
+Finally, the results are summarized and verdicts on each of the solutions' applicability to the research questions will be presented in Section \ref{sec:EvaluationSummary}.
+
+\section{Testing Methodology}\label{sec:TestingMethodology}
 
@@
 
-\section{Responsivity and Latency}
+\section{Responsivity and Latency}\label{sec:ResponsivityAndLatency}
 
@@
 This is important for general user experience and applications where continuous input is required.
-This is also what is generally refered to when determining whether an application has a positive user experience, since a program that can take seconds to respond to the user begets lower productivity and a more frustrating experience.
-To analyse this, two points of data are recorded: First, the amount of time between when the client records a keystroke and when it is actually inputted on the host machine, and seccondly, the amount of time between the host machine updating their screen and when the client actually displays that update.
+This is also what is generally referred to when determining whether an application has a positive user experience, since a program that can take seconds to respond to the user begets lower productivity and a more frustrating experience.
+To analyze this, two points of data are recorded: First, the amount of time between when the client records a keystroke and when it is actually inputted on the host machine, and secondly, the amount of time between the host machine updating their screen and when the client actually displays that update.
 The first datum is generally referred to as the latency of the application, since even though each trial is done using the same network setup different programs may have different optimizations built in that reduce the time needed for the host to implement keystrokes made by the client.
@@
 
-Each application does a good job sending data from the client to the host in a timely manner, with the majority of keystrokes being realized within a tenth of a seccond of being pressed on the client machine.
+Each application does a good job sending data from the client to the host in a timely manner, with the majority of keystrokes being realized within a tenth of a second of being pressed on the client machine.
 The forked Moonlight implementation keeps a latency under 20ms regardless of the load of the host machine, with more than 75\% of all keystrokes being realized within 10ms.
@@
 This lines up with it's intended use described in Section \ref{subsec:VirtualNetworkComputing}, being the simplest and most straightforward method of remote desktop control without much focus on performance or security.
-That's not so say VNC is unsuable or even difficult to use; latency of 160ms is still usable in all but the most precise situations.
+That's not so say VNC is unusable or even difficult to use; latency of 160ms is still usable in all but the most precise situations.
 Though this is only one metric to consider, the other is the amount of time it take for the client to display the results of their input.
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
   \centering
@@
 But once the host machine is put under heavy load, CRD begins to drop frames, with a few being dropped under GPU stress, and 32\% of frames being dropped while under CPU stress.
-This can make CRD unpleasent and frustrating to use while the host is running a CPU intensive program as not only does the client take much more time to update the screen to reflect that status of the host computer, but many of the updates are lost completely and not shown to the user.
+This can make CRD unpleasant and frustrating to use while the host is running a CPU intensive program as not only does the client take much more time to update the screen to reflect that status of the host computer, but many of the updates are lost completely and not shown to the user.
 RDP is also able to display all the data sent by the host without dropping any frames under any condition.
 VNC is again proved to be the least performant, with frames being dropped regardless of the load being put on the host computer.
-This is still in line with it's intended use of basic remote control, though it is dissapointing to see about 15\% loss even under ideal conditions.
+This is still in line with it's intended use of basic remote control, though it is disappointing to see about 15\% loss even under ideal conditions.
 
-\section{Quality}
+\section{Quality}\label{sec:Quality}
 
@@
 When analyzing the forked version of Moonlight, specific quantitative data described in Section \ref{sec:DevelopingTestingAndMeasurementTools} can be collected directly from the application while it is running.
-\todo{put in data}
+Once the streaming session concludes, the statistics are collated into human-readable numbers and presented in the console.
+What follows is an example output from one of the test streaming sessions:
 
-Chrome Remote Desktop also provides some quantitative data.\todo{V8 compression}
+\begin{lstlisting}[style=plaintext,title=Statistics generated after a streaming session]
+Global Video Stats
+----------------------------------------------------------
+Incoming frame rate from network: 62.18 FPS
+Decoding frame rate: 62.18 FPS
+Rendering frame rate: 62.18 FPS
+Frames dropped by your network connection: 0.00%
+Average network latency: 1 ms (variance: 0 ms)
+Average decoding time: 4.27 ms
+Average rendering time (including monitor V-sync latency): 0.06 ms
 
-RDP does not provide quantitative to analyse it's quality, but speaking from a user experience perspective\dots
+Total dropped frames: 0
+\end{lstlisting}
+
+This data shows that Moonlight is able to render all frames sent by the host without dropping any frames due to the network or decoding delays.
+It is also able to keep the stream running at over 60 FPS throughout the entire session, which far exceeds the requirements of a pleasing to use system.
+Taking into account the average frame decoding time of $4.27ms$ and the average rendering time of $0.06ms$, the program still has plenty of time in-between frames in case the streaming conditions are less favorable:
+$(4.27ms+0.06ms) = 4.33ms < 16.\overline{66}ms = (\frac{1000ms}{60fps})$.
+It is also worth noting that while testing the streaming quality, there were very few visual artifacts while monitoring the screen.
+There were very few times where the screen appeared to be run through a compression filter, and it for the most part kept the visual fidelity of the host machine.
+
+Chrome Remote Desktop also provides some quantitative data during streaming, which gives an insight into it's performance usage and speed.
+Figure \ref{fig:crdStats} shows screenshots of the stream during the most demanding test session.
+
+\begin{figure}[h]
+  % \centering
+  % \begin{subfigure}{.25\textwidth}
+  %   \centering
+  %   \includegraphics[width=1\linewidth]{Figures/crd/bandwidth}
+  %   \caption{Bandwidth usage in bits per second}
+  %   \label{fig:crdBandwidth}
+  % \end{subfigure}%
+  \begin{subfigure}{.33\textwidth}
+    \centering
+    \includegraphics[width=1\linewidth]{Figures/crd/delay}
+    \captionsetup{width=.85\linewidth}
+    \caption{delay of image rendering in milliseconds}
+    \label{fig:crdDelay}
+  \end{subfigure}%
+  \begin{subfigure}{.33\textwidth}
+    \centering
+    \includegraphics[width=1\linewidth]{Figures/crd/fps}
+    \captionsetup{width=.85\linewidth}
+    \caption{Frames per second encoded by host and decoded by client}
+    \label{fig:crdFPS}
+  \end{subfigure}%
+  \begin{subfigure}{.33\textwidth}
+    \centering
+    \includegraphics[width=1\linewidth]{Figures/crd/quality}
+    \captionsetup{width=.85\linewidth}
+    \caption{VP8 compression quality of current frames. Lower means CRD is struggling to keep up}
+    \label{fig:crdQuality}
+  \end{subfigure}
+  \captionsetup{width=.8\linewidth}
+  \caption{Statistics from CRD during the most demanding test session. Data recorded over 60 seconds.}
+  \label{fig:crdStats}
+\end{figure}
+
+Unfortunately, as there is no official documentation for this data, analysis on these figures is limited but insights on the stream quality can still be found.
+As seen in Figure \ref{fig:crdDelay}, the delay of image rendering climbs as the CPU is put under stress, with frames sometimes taking entire seconds to be rendered on the client machine.
+It is unclear on how much of the delay is due to the host machine versus the client machine, as the delay obviously rises as the host's CPU is put under stress, but it is clear that as the host machine is under heavy load the frames can take entire seconds to be rendered on the client machine.
+Figure \ref{fig:crdQuality} shows the quality setting of the VP8 compression algorithm being adjusted over time to keep up with demand.
+This is a decent indication of the quality of the video stream over time, as the quality drops in an attempt to keep the stream updating the client with as much frequency as possible instead of sending full frames.
+With the knowledge that CRD leverages the VP8 compression algorithm \cite{miniorange_chromoting}, it is possible to infer than as the quality goes down, the frames being sent contain less information about the current frame and instead more temporal data to construct the new frame from the previous frame.
+This is also seen with visual smudging on the client's screen as the stream slows down.
+
+Microsoft's Remote Desktop Protocol does not provide quantitative data to analyze it's quality, but speaking from a user experience perspective, and considering the data provided in the previous section, RDP performed very well under each of the testing conditions.
+By dropping zero frames and visually keeping up with the host machine, it is clear that RDP works extremely well over a local network.
+Similar to Moonlight, there was very little loss of visual fidelity in the client's screen compared to the host's screen.
+However, noticing that the bandwidth usage can climb upwards of 10 Mb/s, such streaming may cause issues while attempting to access the host machine while outside of the local network \cite{rdp_bandwidth}.
+This will be further discussed in Section \ref{sec:ConclusionLimitations}.
 
 VNC provides little in the form of quality statistics, but given it's performance and responsiveness discussed in the previous section, it is clear that the quality of the stream suffers heavily whenever the host is under load.
+As discussed in Section \ref{subsec:VirtualNetworkComputing}, VNC was never built to be a highly performant protocol.
+The testing done for this paper really stretches the limits of the protocol and pushes the use case of VNC to the extreme, but it is still worth looking into a common and widely used tool.
 
-\section{Summary}
+\section{Summary}\label{sec:EvaluationSummary}
 
 By combining the information from the previous two sections, a wholistic picture of the performance of each application can be drawn.
+Firstly, Virtual Network Computing works as a bare-bones method of connecting to a host machine from a client device in the simplest way possible, but struggles at keeping up with even mildly demanding tasks.
+As seen in Figure \ref{fig:ColorDelayLoss}, VNC failed to provide a smooth experience for the user even in ideal conditions by simply not rendering updates to the screen in time for the next frame to be shown.
+This, combined with a slow response time and delayed inputs, causes a poor experience for the user when attempting to use demanding applications or anything that requires speed.
 
-\todosection
+Secondly, Chrome Remote Desktop performs well under idea conditions and even succeeds at at maintaining steady performance under some GPU load, but it struggles to keep up whenever the host's CPU is under stress.
+Due to CRD's reliance on web technologies, it too suffers from the same limitations as seen in other web streaming applications such as visual smearing and dependence on the CPU.
+In fact, these issues are something technology companies like Google have been trying so solve, with the release of their new VP9 codec used for YouTube starting in 2014, but such technology has not yet been introduced to CRD \cite{google_io_vp9}.
+Unfortunately, due to the dependence on the CPU, and difficulty in handling conditions where the host computer is running strenuous tasks, Chrome Remote Desktop is not a suitable software solution for the purposes of this paper.
+
+Thirdly, Microsoft's Remote Desktop Protocol proves to be a very performant protocol that can handle nearly every situation without a problem.
+It is slightly CPU limited, as seen in Figure \ref{fig:InputDelay}, but even then it is able to keep input latency under a very reasonable 40ms, and it still manages to render the host's screen without dropping any frames.
+It even proves to be the most stable and consistent solution in Figure \ref{fig:ColorDelay}, producing the smallest variance in the render delay of the client's screen.
+If it weren't for the limitations brought up in Section \ref{subsec:RemoteDesktopProtocol}, RDP would be a great solution for streaming demanding applications to a remote client.
+However, because of these drawbacks, RDP is not suitable for every application.
+
+Finally, Moonlight proves to be a high-performant solution that isn't limited by the host's CPU or GPU being under stress.
+Each test resulted in highly similar results, and while it does on average have slightly higher response times compared to RDP (Figures \ref{fig:InputDelay} and \ref{fig:ColorDelay}), it only falls behind by 5-10ms -- \todoquestion{is this appropriate?}quicker than the blink of an eye.
+By not dropping any frames, and numerically proving it's ability to keep the stream above 60 frames per second, Moonlight proves to be a great solution for streaming demanding applications to a remote client without compromising its usability.

### 2022-04-05 paper/Chapters/Conclusion.tex
Finish draft of evaluation chapter
Draft Intro and Conclusion chapters

@@
 
-\section{Summary}
+\section{Summary}\label{sec:ConclusionSummary}
 
-\todosection
+This thesis has proved that it is possible to produce a hardware and software solution that is capable of leveraging the power of a desktop computer remotely without sacrificing the user's experience.
+It improves upon existing solutions by continuing to offer a performant and high-quality stream of the host's screen and audio to the user even as the host is running highly demanding applications.
+Each of the questions posed in Section \ref{sec:ResearchQuestions} can be answered individually.
 
-\section{Limitations}
+\begin{enumerate}
+  \setcounter{enumi}{0} % Start the enumeration at X+1
+  \item \textbf{Hardware Feasibility:} \emph{What hardware is needed in order to power a mobile device capable of acting as a client?}
+\end{enumerate}
 
-\todosection
+\noindent
+This thesis project is able to run on embedded Single Board Computers and hardware available for purchase today.
+While this project built a custom PCB for the Raspberry Pi Compute Module 4 (Discussed in section \ref{sec:ChoosingParts}), the software could also run on consumer-grade Raspberry Pi models as well.
 
-\section{Future Research}
+\begin{enumerate}
+  \setcounter{enumi}{1} % Start the enumeration at X+1
+  \item \textbf{Hardware Cost:} \emph{Is it feasible to construct such as device at a lower or comparable cost to existing solutions?}
+\end{enumerate}
 
-\todosection
+\noindent
+Yes, the total cost for a single hardware unit, including the board and all its components, totalled in under \$150, and could even be produced for less (Discussed in Section \ref{subsec:HardwareCost}).
+This cost is considerably cheaper than buying a new laptop device, and often cheaper than buying a used laptop with comparable power.
+
+\begin{enumerate}
+  \setcounter{enumi}{2} % Start the enumeration at X+1
+  \item \textbf{Communication Protocol:} \emph{Does a protocol exist that is efficient enough to stream demanding applications to a client?}
+\end{enumerate}
+
+\noindent
+Yes, Nvidia's GameStream protocol proved itself to be capable of streaming highly demanding applications to a client device without any loss of quality or performance (Discussed in Section \ref{sec:EvaluationSummary}).
+Even though it's original purpose and namesake is to stream video games, the high demands of such a use case have made it applicable to much more than just that.
+While Microsoft's Remote Desktop Protocol proved to be powerful enough to stream demanding applications to a client device, it's limitations and closed-source nature prevent it from being the viable solution for certain use cases (Discussed in \ref{subsec:RemoteDesktopProtocol}).
+
+\begin{enumerate}
+  \setcounter{enumi}{3} % Start the enumeration at X+1
+  \item \textbf{Software Application:} \emph{Can a software solution be built to utilize this protocol in a manner that performant enough?}
+\end{enumerate}
+
+\noindent
+Yes, the solution produced in this thesis is built upon the open source project Moonlight, which utilizes the GameStream protocol, to provide a software application that is capable of streaming demanding applications to a client device while remaining performant enough to provide a positive user experience (Discussed in Section \ref{sec:EvaluationSummary}).
+
+\section{Limitations}\label{sec:ConclusionLimitations}
+
+While this thesis was constructed with the utmost attentiveness with regards to the time and resources available, there are still limitations in what could be accomplished.
+The largest limitation is the lack of testing in alternative environments.
+All of the testing performed in Chapter \ref{Chapter6} was performed in an ideal network environment, where there was minimal network latency between the host and client machines due to them both being connected to the same network by ethernet.
+While this is a good indicator of the best possible performance, it is not very indicative of the performance that could be expected in a real-world environment.
+More testing should be done where the host and client machines are located on different networks akin to the use case described in Section \ref{sec:ApplicationToModernDay}.
+
+Secondly, due to the ongoing chip shortage and the lack of experience in developing PCB boards, it is very likely that there are better ways to develop the PCB board for this project.
+While the board was able to meet the requirements for this project, there we're many instances (some described in table \ref{tab:SelectedComponents}) where alternative parts had to be chosen and designed around because the first-choice was out of stock.
+It is very possible that the board could have been constructed quicker and at a cheaper price if more parts were available for purchase.
+
+\section{Future Research}\label{sec:ConclusionFutureResearch}
+
+Should this project be continued in the future, there are multiple ways that additional time and resources could further improve this project.
+Similar to the previous section, if more time were available, more testing could be done in various environments to identify any weaknesses in the software that may have gone unnoticed.
+After experimenting with streaming over different networks, it would be beneficial to look at polishing the hardware of the project into a self-contained product.
+As it stands, it is currently a PCB with peripherals attached that make it cumbersome to carry around.
+The groundwork for this has already been laid out in Section \ref{subsec:DesigningHDMI}, with FFC connectors established for one HDMI and one USB port for connection to an embedded screen, and with the footprint of the final PCB and battery fitting within the footprint of the chosen screen.
+All that is left is to build an enclosure to house the final product.

### 2022-04-05 paper/Chapters/Introduction.tex
Draft Intro and Conclusion chapters

@@
 
-\todosection
+As computers continue to grow to become more and more integrated into our daily lives, purchasing a computer has always been a balance.
+Whether its deciding between something cheap or something long-lasting, a machine that is pre-built or a machine that is custom tailored and custom built, or, a tradeoff that has become much more common in recent years, something powerful or something portable.
+As with anything, compromises can always be made and a middle ground may be found, but the only way to get the best of both worlds currently is to pay a price premium.
+If someone were to look for a computer that is powerful, long-lasting, and upgradable, they would be directed towards a desktop computer.
+However, if they needed something with the same power but portable, their only option would be to purchase a laptop costing much more than its desktop counterpart and often with limitations in upgradeability down the line.
+Without purchasing both an expensive desktop and an expensive laptop, there currently is no way to get the benefits of both power and portability.
 
@@
 
-\todosection
+The primary goal of this thesis is to create a portable hardware and software solution that can be used to control a desktop computer remotely while leveraging all of it's power.
+There are existing software solutions that give a user remote control of their desktop computer from a different location, but many of them fail to provide a high-quality and performant solution to working with highly demanding programs that put the desktop computer under intense load.
+Once attempting to run an intensive application such as rendering video, training Artificial Intelligence, or developing a video game, existing solutions fail to keep up.
+
+In other words, this thesis seeks to provide a user who already owns a powerful desktop computer with a mobile solution that gives them the ability to leverage it's power remotely.
 
@@
 
-\todosection
+This thesis is split into seven chapters.
+Chapter \ref{Chapter2} introduces background knowledge surrounding existing computers, the shift to a more mobile computing world, and the modern day applications for this technology.
+Chapter \ref{Chapter3} discusses the current the state of the art in terms of existing solutions for remote computer access, and presents a list of questions that this thesis seeks to answer.
+Chapters \ref{Chapter4} and \ref{Chapter5} detail the requirements, design, and production of the hardware and software portions of the project respectively.
+This design is tested against the requirements and the existing solutions in Chapter \ref{Chapter6}.
+Finally, Chapter \ref{Chapter7} concludes the thesis with a summary of the project's innovations, limitations, and future research.

### 2022-04-05 paper/Chapters/Background.tex
Prepare Complete (not final) draft
This draft will be sent of to each of the committee members to read through before the thesis defense.
The PDF has been rendered in "final" mode, so no comments will show through.

small.pdf is a compressed version of the main pdf, with no differences except each of the images have been compressed to 300dpi.

@@
-\chapter{Background} % Main chapter title
+\chapter{Background}
 
-\label{Chapter2} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
+\label{Chapter2}
 
@@
 As time passed and technology progressed, a single computer could be produced that served multiple functions as well as be adaptable programmable to handle new tasks that were not considered when the machine was originally built.
-This gave rise to what was known as \enquote{Autonomic Computing}, a system of characteristics that are built into a computer to help it self-manage it's resources and adapt to it's administrator's requirements without the need to be redesigned for it's new purpose \citep{AutonomicComputing}.
+This gave rise to what was known as \enquote{Autonomic Computing}, a system of characteristics that are built into a computer to help it self-manage it's resources and adapt to it's administrator's requirements without the need to be redesigned for it's new purpose \cite{AutonomicComputing}.
 Autonomic Computing was designed to combat the exponential complexity crisis that came from the widespread availability of computers in different disciplines, since now anyone who could afford a computer could program it for any purpose.
 As new use cases were found for computers in day-to-day life and new technologies were being developed to interface with the world around us, the need for componentization became ever more apparent.
-Researchers at IBM knew that the best way to address the looming problem of runaway complexity in computers was to develop a way for the computers to automatically interface with any new components that are installed, and to configure itself to only use what is necessary for the job at hand \citep[p.~43]{AutonomicComputing}.
+Researchers at IBM knew that the best way to address the looming problem of runaway complexity in computers was to develop a way for the computers to automatically interface with any new components that are installed, and to configure itself to only use what is necessary for the job at hand \cite{AutonomicComputing}.
 This allowed hardware and software developers to focus on building their products to work with a common standard rather than having to manually integrate their products with every computer.
@@
 Once the idea of modularity began to take hold in the computer industry, attention was turned once again towards specializing individual computers.
-Now that a computer's physical footprint can be minified, and peripheral components can be added and removed without a complete reengineering of the device, a single computer can be optimized to handle a single task without the overhead cost of building a monolithic computer \citep{Burbeck2007ComplexityAT}.
+Now that a computer's physical footprint can be minified, and peripheral components can be added and removed without a complete reengineering of the device, a single computer can be optimized to handle a single task without the overhead cost of building a monolithic computer \cite{Burbeck2007ComplexityAT}.
 Now a computer can be specialized to serve a particular purpose or set of purposes while keeping development time comparatively short.
@@
 Pieces of the computer could be bought separately, and single parts could be upgraded or replaced by anyone will a little hardware knowledge.
-Even though they aren't very portable, desktop computers allow consumers to access greater processor power for a lower cost compared to portable devices \citep{Meyer20145RS}.
+Even though they aren't very portable, desktop computers allow consumers to access greater processor power for a lower cost compared to portable devices \cite{Meyer20145RS}.
 Due to this, a desktop PC is often seen at one's home or place of work hardwired to the wall where it remains unmoving unless it needs upgrading, cleaning, or repairs.

### 2022-04-05 paper/Chapters/Hardware.tex
Prepare Complete (not final) draft
This draft will be sent of to each of the committee members to read through before the thesis defense.
The PDF has been rendered in "final" mode, so no comments will show through.

small.pdf is a compressed version of the main pdf, with no differences except each of the images have been compressed to 300dpi.

@@
-\chapter{Developing the Hardware} % Main chapter title
+\chapter{Developing the Hardware}
 
-\label{Chapter4} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
+\label{Chapter4}
 
@@
   \end{subfigure}
-  \caption{The PCB's HDMI circuit}
+  \caption[PCB HDMI Circuit]{The PCB's HDMI circuit}
   \label{fig:HDMICircuit}
@@
   \end{subfigure}
-  \caption{The circuitry surrounding the CM4's two mezzanine connectors}
+  \caption[PCB CM4 Circuit]{The circuitry surrounding the CM4's two mezzanine connectors}
   \label{fig:CM4Circuit}
@@
   \includegraphics[width=.75\textwidth,page=3]{Figures/kicad/lathrum_thesis_schematic.pdf}
-  \caption[USB Schematic]{Schematic layout for interfacing with the Compute Module 4. Expanded in Appendix \ref{app:pcb_schematic_cm4}}
+  \caption[CM4 Schematic]{Schematic layout for interfacing with the Compute Module 4. Expanded in Appendix \ref{app:pcb_schematic_cm4}}
   \label{fig:pcb_schematic_cm4}
@@
 \end{figure}
-
-\todo[inline]{Insert pictures of PCB. Thinking at least one showing the flaws of the first board, another of the completed board}
 
@@
 In particular, USB devices plugged into any of the ports on the board were able to receive power from the board but were unable to transfer data in either direction.
-\todo[inline]{I am currently going through and diagnosing this issue}
+Thankfully, this did not prevent testing of the board from being completed, but it did stop the board from being considered a final complete product.
+This board could still be used for testing by connecting peripherals such as the keyboard and mouse through bluetooth and powering the screen through USB.
+Figure \ref{fig:AssembledPCB} shows the final assembled board, with the final PCB layouts in Figures \ref{app:pcb_layout_front} and \ref{app:pcb_layout_back}.

### 2022-04-05 paper/main.tex
Start Evaluation section, add graphs
Finish draft of evaluation chapter
Prepare Complete (not final) draft
This draft will be sent of to each of the committee members to read through before the thesis defense.
The PDF has been rendered in "final" mode, so no comments will show through.

small.pdf is a compressed version of the main pdf, with no differences except each of the images have been compressed to 300dpi.

@@
 % draft, % Uncomment to enable draft mode (no pictures, no links, overfull hboxes indicated)
-% final, % Uncomment to enable final mode (hide todos)
+final, % Uncomment to enable final mode (hide todos)
 %nolistspacing, % If the document is onehalfspacing or doublespacing, uncomment this to set spacing in lists to single
@@
 }
+\lstdefinestyle{plaintext}{
+  belowcaptionskip=1\baselineskip,
+  breaklines=true,
+  frame=single,
+	numbers=none,
+  showstringspaces=false,
+  basicstyle=\footnotesize\ttfamily,
+}
 \lstset{escapechar=@,style=customc}
+
+\usepackage{pgfplots} % Required for the plots
+\usepgfplotslibrary{statistics}
+\pgfplotsset{compat=1.8}
 
@@
 	\addchaptertocentry{\acknowledgementname} % Add the acknowledgements to the table of contents
-	The acknowledgments and the people to thank go here, don't forget to include your project advisor\ldots
+	I would like to express my graditude to my thesis director, Dr.
+Robert Heinrichs, and to each of my committee members, Professor Ruben Acu\~{n}a and Dr.
+Shawn Jordan, for their help in the development of this thesis.
+	Without their guidance, I would not have been able to step so far outside of my comfort zone and learn about so many new topics to make this thesis happen.
+	I would also like to thank my friends and family who were always there to bounce ideas off of and inevitably tell me when I was trying to fit too many words into a single section.
 \end{acknowledgements}
@@
 
-\dedicatory{For/Dedicated to/To my\ldots}
+\dedicatory{For my parents, and their endless support, encouragement, and guidance.}
 

### 2022-04-05 paper/Chapters/Conclusion.tex

@@
-\chapter{Conclusion} % Main chapter title
+\chapter{Conclusion}
 
-\label{Chapter7} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
+\label{Chapter7}
 
@@
 Should this project be continued in the future, there are multiple ways that additional time and resources could further improve this project.
+The first improvement to make is to fix the issue with USB data described in section \ref{subsec:Manufacturing3}.
+This is the only flaw in the hardware preventing it from being considered a completed product.
 Similar to the previous section, if more time were available, more testing could be done in various environments to identify any weaknesses in the software that may have gone unnoticed.

### 2022-04-05 paper/Chapters/StateOfTheArt.tex
Tweaks and updates to the hardware sections
Now a nice big sprint can begin. Just need the time
(almost) finish draft of hardware section
Still need to investigate the USB issues and maybe add more detail to the end of the chapter, but other than that I have a couple of questions to resolve then I'm happyish with the chapter

@@
-\chapter{State of the Art} % Main chapter title
+\chapter{State of the Art}
 
-\label{Chapter3} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
+\label{Chapter3}
 
@@
 
-Google Chromoting\todo{Should this chapter be called Chromoting?
-"Chrome Remote Desktop" is the much more recognizable name though}, implemented publicly as Chrome Remote Desktop, is an open source protocol that enables remote desktop control through any web browser running on the Chromium engine \cite{chromotingBuildInstructions}.
+Google Chromoting, implemented publicly as Chrome Remote Desktop, is an open source protocol that enables remote desktop control through any web browser running on the Chromium engine \cite{chromotingBuildInstructions}.
 It prioritizes low bandwidth usage and ease of access, allowing easy server install on every major desktop operating system, and enables any web browser or mobile device to act as a client.
@@
 
-\section{Summary}\label{sec:StateOfTheArtSummary}
-
-\todosection
-
 \section{Research Questions}\label{sec:ResearchQuestions}

### 2022-04-13 paper/Chapters/Evaluation.tex
Prepare Complete (not final) draft
This draft will be sent of to each of the committee members to read through before the thesis defense.
The PDF has been rendered in "final" mode, so no comments will show through.

small.pdf is a compressed version of the main pdf, with no differences except each of the images have been compressed to 300dpi.
Feedback from advisor and thesis defense

@@
-\chapter{Evaluation} % Main chapter title
+\chapter{Evaluation}
 
-\label{Chapter6} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
+\label{Chapter6}
+
+\todoadvisor{How many samples were taken to create the figures? How many times was the test run?}
 
@@
   \item The client machine connects to the host computer using the application being tested.
-  \item The user presses the \enquote{0} key on the client's numpad to synchronize the two devices' data recording tools.
-  \item The user begins by typing on the client's keyboard slowly, allowing both computers to record their data.
-  \item The user then types on the client's keyboard quickly to ensure that quick successive inputs are handled correctly and no data is lost or duplicated.
+  \item The user presses the \enquote{0} key on the client's numpad to synchronize \todoadvisor{Syncronize what?}the two devices' data recording tools.\todoadvisor{How far off are they?}
+  \item The user begins by typing on the client's keyboard \todoadvisor{what does slowly mean?}slowly, allowing both computers to record their data.
+  \item The user then types on the client's keyboard quickly to ensure that \todoadvisor{Why is there a seperation here?}quick successive inputs are handled correctly and no data is lost or duplicated.
   \item Once typing is complete, both the client and the host computers close their data recording tools and the client ends the streaming session.
@@
 \end{figure}
+\todoadvisor{Maybe put full figure in the appendix?}
 
@@
 Apart from the performance of each application, the quality of the user experience must also be taken into account when considering each application.
+Each application will be assessed individually, and then compared in Section \ref{sec:EvaluationSummary}
 When analyzing the forked version of Moonlight, specific quantitative data described in Section \ref{sec:DevelopingTestingAndMeasurementTools} can be collected directly from the application while it is running.
@@
   \captionsetup{width=.8\linewidth}
-  \caption{Statistics from CRD during the most demanding test session. Data recorded over 60 seconds.}
+  \caption[CRD Data]{Statistics from CRD during the most demanding test session. Data recorded over 60 seconds.}
   \label{fig:crdStats}
@@
 This is a decent indication of the quality of the video stream over time, as the quality drops in an attempt to keep the stream updating the client with as much frequency as possible instead of sending full frames.
-With the knowledge that CRD leverages the VP8 compression algorithm \cite{miniorange_chromoting}, it is possible to infer than as the quality goes down, the frames being sent contain less information about the current frame and instead more temporal data to construct the new frame from the previous frame.
+With the knowledge that CRD leverages the VP8 compression algorithm \cite{miniorange_chromoting}, it is possible to infer that as the quality goes down, the frames being sent contain less information about the current frame and instead more temporal data to construct the new frame from the previous frame.
 This is also seen with visual smudging on the client's screen as the stream slows down.
@@
 As seen in Figure \ref{fig:ColorDelayLoss}, VNC failed to provide a smooth experience for the user even in ideal conditions by simply not rendering updates to the screen in time for the next frame to be shown.
-This, combined with a slow response time and delayed inputs, causes a poor experience for the user when attempting to use demanding applications or anything that requires speed.
+This, combined with a slow response time and delayed inputs, causes a poor experience for the user when attempting to use demanding applications or anything that requires speed, making it unsuitable for the purposes of this paper.
 
-Secondly, Chrome Remote Desktop performs well under idea conditions and even succeeds at at maintaining steady performance under some GPU load, but it struggles to keep up whenever the host's CPU is under stress.
+Secondly, Chrome Remote Desktop performs well under ideal conditions and even succeeds at at maintaining steady performance under some GPU load, but it struggles to keep up whenever the host's CPU is under stress.
 Due to CRD's reliance on web technologies, it too suffers from the same limitations as seen in other web streaming applications such as visual smearing and dependence on the CPU.
@@
 Finally, Moonlight proves to be a high-performant solution that isn't limited by the host's CPU or GPU being under stress.
-Each test resulted in highly similar results, and while it does on average have slightly higher response times compared to RDP (Figures \ref{fig:InputDelay} and \ref{fig:ColorDelay}), it only falls behind by 5-10ms -- \todoquestion{is this appropriate?}quicker than the blink of an eye.
+Each test resulted in highly similar results, and while it does on average have slightly higher response times compared to RDP (Figures \ref{fig:InputDelay} and \ref{fig:ColorDelay}), it only falls behind by 5-10ms -- quicker than the blink of an eye.
 By not dropping any frames, and numerically proving it's ability to keep the stream above 60 frames per second, Moonlight proves to be a great solution for streaming demanding applications to a remote client without compromising its usability.

### 2022-04-13 paper/Chapters/Hardware.tex
Feedback from advisor and thesis defense

@@
 This section will explain the different constraints taken into account while developing the hardware portion of this project.
+\todoadvisor{Target user or audience of the product should be described more accurately.
+Since no studies or research has been conducted regarding which group of people are interested in / will use this product, no specific requirements can be derived for the system.
+At this point the only source of the requirements is the author of the thesis and hence it should be worded this way.
+The thesis should initially state that the requirements for the user experience are based on the authors requirements/preferences, to make clear that no other users were involved or a study conducted with other persons.}
 First, the hardware must be performant enough to provide a positive user experience (\ref{subsec:HardwarePerformance}), and secondly the product must be cheap enough to justify it over similar alternatives (\ref{subsec:HardwareCost}).
@@
 \todoadvisor{Why are those components not described in the same way as power/USB/etc.?}\todoquestion{The Ethernet port is hooked directly into the CM4 connector with nothing interested except that they are also differential pairs, as described in the HDMI section.
-I suppose they could be explained fully, but I worry that it wouldn't add any useful information.}
+I suppose they could be explained fully, but I worry that it wouldn't add any useful information.}\todoadvisor{Explain why it's skimed over}
 See Appendices \ref{AppendixA} and \ref{AppendixB} for further information.
@@
 The only issue apart from some ports being too close together with real-world cables was one component having the wrong footprint on the board.
-Unfortunately, the component \emph{U2 (MCP73871-2CCI\_ML)}, is described as using the \emph{QFN-20-1EP\_4x4mm} footprint in it's datasheet\todoquestion{Does this need to be cited?
-Can you cite a datasheet like this?}, but actually uses a revised version of the footprint, shown in Figure \ref{fig:FootprintQFN204x4}.
+Unfortunately, the component \emph{U2 (MCP73871-2CCI\_ML)}, is described as using the \emph{QFN-20-1EP\_4x4mm} footprint in its datasheet\todoquestion{Does this need to be cited?
+Can you cite a datasheet like this?}\todoadvisor{Yes}, but actually uses a revised version of the footprint, shown in Figure \ref{fig:FootprintQFN204x4}.
 This discrepancy meant that that component could not be soldered, and it's associated circuit would not function properly.

### 2022-04-13 paper/Chapters/Software.tex
Feedback from advisor and thesis defense

@@
-\chapter{Developing the Software} % Main chapter title
+\chapter{Developing the Software}
 
-\label{Chapter5} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
+\label{Chapter5}
 
@@
 Following that, special considerations must be made while developing the software for the CM4 running on an \emph{ARM} processor in Section \ref{sec:DevelopingForARM}, and the development of the suite of tools used to record data is detailed in Section \ref{sec:DevelopingTestingAndMeasurementTools}.
+\todo{Link git repo}
 
@@
 A GameStream session is started by first connecting to a host computer which has enabled the GameStream service through Nvidia's GeForce Experience application.
-Once the service is enabled, the computer will listen on port \emph{47989}\todoquestion{Should this be italicized? quoted? plaintext?} for the following web requests:
+Once the service is enabled, the computer will listen on port \emph{47989} for the following web requests:
 \begin{itemize}
@@
 
+\todoadvisor{Mention this earlier, maybe to the state of the art}
 Moonlight is an unofficial third-party open-source client for the Nvidia GameStream protocol \cite{moonlight}.
@@
 In order to create an out-of-the-box experience for someone using the hardware for this project, a custom operating system image is created that has all the dependencies, libraries, and applications pre-installed and pre-configured.
-Manjaro provides good tools for building and managing custom images of the Manjaro operating system for \emph{ARM} architectures \cite{manjaro_arm_tools}, with the only drawback being that it is that it must be run on an \emph{ARM} processor, and is much more difficult to run on another architecture.
+Manjaro provides good tools for building and managing custom images of the Manjaro operating system for \emph{ARM} architectures \cite{manjaro_arm_tools}, with the only drawback being that it must be run on an \emph{ARM} processor, and is much more difficult to run on another architecture.
 To make development easier, a toolchain was built that could be run on the CM4 to build the code that is written on other computers.
-All the developer needs to do is to pull the git repository of the project, run a set up script to install dependencies, and then run the build script to compile everything into an immediately deployable OS image.
+All the developer needs to do is to pull the git repository\todoadvisor{Cite the git repository} of this project onto the \emph{ARM} architecture, run a set up script to install dependencies, and then run the build script to compile everything into an immediately deployable OS image.
 This build script will automatically compile all the necessary code using \emph{cmake}, package all the necessary files using \emph{makepkg}, and build a fresh OS image that is ready to flash.
@@
   cd ./moonlight/build # Go to build directory
-  cmake ../ # Run cmake
+  cmake ../ # Run cmake on the project directory
   make # Build the project
@@
 
+\todoadvisor{Explicitly mention that this is work I did}
 In order to evaluate the performance of the project, a suite of tools must be developed to test both the newly developed application and the existing solutions.
@@
 This is done using the following steps once the client is connected to the host:
+\todoadvisor{Mention that the times need to be syncronized}
 
@@
 This is sufficient to change the color of the screen back and forth between easily distinguishable colors for both the client and the host to record.
+\todoadvisor{Mention that this is recorded by the dataloggerz}
 With both of these datasets recorded, the next step is to analyze the data and evaluate the performance of the application.
+
+\todoadvisor{Summary}

### 2022-04-16 paper/Chapters/Hardware.tex

--- 2022-04-05 paper/Chapters/Hardware.tex
@@
 
+The entirety of the hardware schematics and designs developed for this project are available on GitHub at: \url{https://github.com/Dylancyclone/thesis/kicad}.
+
 \section{Requirements}\label{sec:HardwareRequirements}
@@
 First, the hardware must be performant enough to provide a positive user experience (\ref{subsec:HardwarePerformance}), and secondly the product must be cheap enough to justify it over similar alternatives (\ref{subsec:HardwareCost}).
+
+Due to the lack of preliminary studies or research into the topic of target audience for this project, the requirements laid out in this section are based on the author's own requirements and preferences.
+The requirements are constructed with the target user being someone who already has a working desktop computer and wishes to leverage it's power remotely without sacrificing the user experience.
 
@@
 This allows each function of the board to be tested individually and in isolation so that changes could be made incrementally without needing to worry about their integration with the rest of the board.
-\todoquestion{This is that tricky sentence}For example, in order to build a system that would allow the board to be powered via USB or an internal battery, as well as charge the battery simultaneously, a complex circuit would have to be built that managed Lithium Polymer (LiPo) batteries which could cause fires if mishandled.
+For example, in order to build a system that would allow the board to be powered via USB or an internal battery, as well as charge the battery simultaneously, a complex circuit would have to be built that managed Lithium Polymer (LiPo) batteries which could cause fires if mishandled.
 Since power is obviously a critical component of the board, A secondary way of feeding power directly the board was built to bypass all that circuitry in case it fails.
@@
 Finally, each of the other components are connected to the CM4 board in the same manner as described throughout this chapter, such as the Ethernet adapter, the SD card reader, and remaining ports.
-\todoadvisor{Why are those components not described in the same way as power/USB/etc.?}\todoquestion{The Ethernet port is hooked directly into the CM4 connector with nothing interested except that they are also differential pairs, as described in the HDMI section.
-I suppose they could be explained fully, but I worry that it wouldn't add any useful information.}
+These ports are connected by simply connecting the appropriate pin on the CM4's mezzanine connector to the corresponding pin on the component.
+Apart from the concerns brought up by keeping the differential pairs in line and the power carefully managed as with the previous components described in detail, there are no special considerations that need to be made while designing the layout for these components.
 See Appendices \ref{AppendixA} and \ref{AppendixB} for further information.
@@
 The only issue apart from some ports being too close together with real-world cables was one component having the wrong footprint on the board.
-Unfortunately, the component \emph{U2 (MCP73871-2CCI\_ML)}, is described as using the \emph{QFN-20-1EP\_4x4mm} footprint in it's datasheet\todoquestion{Does this need to be cited?
-Can you cite a datasheet like this?}, but actually uses a revised version of the footprint, shown in Figure \ref{fig:FootprintQFN204x4}.
+Unfortunately, the component \emph{U2 (MCP73871-2CCI\_ML)}, is described as using the \emph{QFN-20-1EP\_4x4mm} footprint in its datasheet \cite{MCP73871-2CCI/ML-ND_datasheet}, but actually uses a revised version of the footprint, shown in Figure \ref{fig:FootprintQFN204x4}.
 This discrepancy meant that that component could not be soldered, and it's associated circuit would not function properly.

### 2022-04-16 paper/Chapters/StateOfTheArt.tex
Feedback from advisor and thesis defense

@@
 In 2013, a group of students at Case Western Reserve University developed Moonlight, an open source implementation of the GameStream protocol \cite{moonlight}.
-This open source implementation allows the development of GameStream compatible clients that can run on other devices, from other computers to embedded ARM devices.
+The project is structured as a core library written in \emph{C} with a number of community built clients built on top of it for various platforms.
+In essence, it turns the closed-source black-box protocol developed by Nvidia into a usable programming interface.
+By reverse engineering the protocol originally implemented by the Nvidia Shield, Moonlight enables the development of a GameStream client for almost any operating system.
+The Moonlight community has already developed fantastic clients for many platforms such as Windows, Mac, Linux, Android, iOS, Web, and even other gaming consoles.
+All that is needed for the purposes of this project is to make a fork of the project that tweaks it to best suit the CM4 that drives the hardware and develop a suite of tools to test it's functionality.
 This stands as a good starting place to develop the solution described in this paper.
@@
   \item \textbf{Communication Protocol:} \emph{Does a protocol exist that is efficient enough to stream demanding applications to a client?}
-  \item \textbf{Software Application:} \emph{Can a software solution be built to utilize this protocol in a manner that performant enough?}
+  \item \textbf{Software Application:} \emph{Can a software solution be built to utilize this protocol in a manner that is performant enough?}
 \end{enumerate}

### 2022-04-16 paper/Chapters/Conclusion.tex
Feedback from advisor and thesis defense

@@
 Similar to the previous section, if more time were available, more testing could be done in various environments to identify any weaknesses in the software that may have gone unnoticed.
-After experimenting with streaming over different networks, it would be beneficial to look at polishing the hardware of the project into a self-contained product.
+
+Another aspect of the project worth testing in more detail is the performance of the hardware produced in Chapter \ref{Chapter4} compared to the hardware performance of the development board that is sold with the CM4 (Section \ref{sec:ChoosingParts}).
+The ideal outcome is that the two boards provide the same performance, meaning that there is no unneeded overhead introduced by the board produced in this thesis that should be fixed.
+In the event that there is slowdown somewhere within the custom board, the specific problem should be identified and redesigned to provide optimal performance.
+
+After experimenting with streaming over different networks and testing the project under different conditions, it would be beneficial to look at polishing the hardware of the project into a self-contained product.
 As it stands, it is currently a PCB with peripherals attached that make it cumbersome to carry around.

### 2022-04-16 paper/Chapters/Evaluation.tex

@@
 \label{Chapter6}
-
-\todoadvisor{How many samples were taken to create the figures? How many times was the test run?}
 
@@
   \item The client machine connects to the host computer using the application being tested.
-  \item The user presses the \enquote{0} key on the client's numpad to synchronize \todoadvisor{Syncronize what?}the two devices' data recording tools.\todoadvisor{How far off are they?}
-  \item The user begins by typing on the client's keyboard \todoadvisor{what does slowly mean?}slowly, allowing both computers to record their data.
-  \item The user then types on the client's keyboard quickly to ensure that \todoadvisor{Why is there a seperation here?}quick successive inputs are handled correctly and no data is lost or duplicated.
+  \item The tester presses the \enquote{0} key on the client's numpad to synchronize the two devices' clocks and data recording tools (Detailed below).
+  \item The tester begins by typing on the client's keyboard at a steady pace as close to 50 words per minute as possible, allowing both computers to record their data.
+  \item The tester then types on the client's keyboard at a quicker pace of approximately 100 words per minute to ensure that quick successive inputs are handled correctly and no data is lost or duplicated.
+  Due to the wide variance in what the user can do using the software, it is nearly impossible to test real world scenarios.
+  Instead, high frequency inputs coupled with updating the entire screen mimic high-intensity applications.
   \item Once typing is complete, both the client and the host computers close their data recording tools and the client ends the streaming session.
@@
 Namely, the network latency and difference in computer clocks must be removed.
+By keeping the host and client computers attached to the same network through ethernet, the network latency is minimized.
+But the task of keeping two computer's internal clocks synchronized is not trivial.
+Even after instructing the operating system to synchronize its clock with an internet time server, the two computers may still report different times.
+In some cases, the difference between the host and client computer's system time was noticed to be upwards of a full second after resynchronizing both clocks \cite{time.is}.
 This is partially accounted for in step \emph{6} of the above process, and the remaining inaccuracy can now be removed by subtracting the time difference between the when the client presses the button and the time the host executes that command.
@@
 The first datum is generally referred to as the latency of the application, since even though each trial is done using the same network setup different programs may have different optimizations built in that reduce the time needed for the host to implement keystrokes made by the client.
+Each of the following figures shows data over one trial for each category.
+Each trial lasted approximately 60 seconds following the steps detailed in Section \ref{sec:TestingMethodology}, totaling to approximately 200 keystrokes per trial.
+Limitations in this data is discussed in Section \ref{sec:ConclusionLimitations}.
 Figure \ref{fig:InputDelay} shows the amount of time between the client and the host registering the keystroke under the different test conditions.
@@
 \end{figure}
-\todoadvisor{Maybe put full figure in the appendix?}
 

### 2022-04-16 paper/Chapters/Software.tex

@@
 Designing the software begins with identifying it's requirements in Section \ref{sec:SoftwareRequirements}.
-Then, potential avenues for development are discussed in Section \ref{sec:PotentialAvenues} so that the software does not need to be built from scratch.
+Then, the Nvidia GameStream Protocol is analyzed in depth in Section \ref{sec:UtilizingNVIDIAGameStream} so that the software does not need to be built from scratch.
 Following that, special considerations must be made while developing the software for the CM4 running on an \emph{ARM} processor in Section \ref{sec:DevelopingForARM}, and the development of the suite of tools used to record data is detailed in Section \ref{sec:DevelopingTestingAndMeasurementTools}.
-\todo{Link git repo}
+
+The entirety of the code developed for this project is available on GitHub at: \url{https://github.com/Dylancyclone/thesis/code}.
 
@@
 
-\section{Potential Avenues}\label{sec:PotentialAvenues}
+\section{Utilizing NVIDIA GameStream}\label{sec:UtilizingNVIDIAGameStream}
 
 As with any software project, there are many potential ways to solve the problem.
-In many cases, the easiest place to start is to begin by ruling out potential avenues that are not feasible.
-For example, as evident by the number of existing solutions that attempt to solve this problem (Chapter \ref{Chapter3}), taking into special account Nvidia's recent foray into remote gaming with the Shield (\ref{subsec:NvidiaShieldAndGameStream}), it is not reasonable to attempt to create a brand new software solution from scratch with the time and resources available.
+As evident by the number of existing solutions that attempt to solve this problem (Chapter \ref{Chapter3}), taking into special account Nvidia's recent foray into remote gaming with the Shield (\ref{subsec:NvidiaShieldAndGameStream}), it is not reasonable to attempt to create a brand new software solution from scratch with the time and resources available for this thesis.
 It has taken Nvidia, a company worth hundreds of billions of dollars, years to get to the point they are at now \cite{brown_2013}.
@@
 Instead, given that Nvidia's GameStream technology is the current leader in the realm of high-performance remote computing, the best way to build a software product that meets the requirements laid out in Section \ref{sec:ResearchQuestions} is to build upon what already exists.
-
-\subsection{NVIDIA GameStream}\label{subsec:NVIDIAGameStream}
-
-\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,inline]{Should this subsection and the following subsection be top level sections?
-The organization of the headers is a little strange, but the content flows pretty well}
 
@@
 
-\subsection{Moonlight}\label{subsec:Moonlight}
-
-\todoadvisor{Mention this earlier, maybe to the state of the art}
-Moonlight is an unofficial third-party open-source client for the Nvidia GameStream protocol \cite{moonlight}.
-The project is structured as a core library written in \emph{C} with a number of community built clients built on top of it for various platforms.
-In essence, it turns the process described in the previous section (\ref{subsec:NVIDIAGameStream}) into a usable programming interface.
-By reverse engineering the protocol originally implemented by the Nvidia Shield, Moonlight enables the development of a GameStream client for almost any operating system.
-The Moonlight community has already developed fantastic clients for many platforms such as Windows, Mac, Linux, Android, iOS, Web, and even other gaming consoles.
-All that is needed for the purposes of this project is to make a fork of the project that tweaks it to best suit the CM4 that drives the hardware and develop a suite of tools to test it's functionality.
+By taking advantage of the Moonlight library briefly described in Section \ref{subsec:NvidiaShieldAndGameStream}, the previously described process is translated  into a usable programming interface.
+To accomplish this, a fork of the Moonlight project is created so the GameStream protocol can be leveraged in this thesis.
+With the fork, much of the code required to create the software solution is already provided, all that is left is to tweak the code to ease development for the \emph{ARM} architecture and to build a suite of testing tools that hook into Moonlight to extract statistics.
 
@@
 To make development easier, a toolchain was built that could be run on the CM4 to build the code that is written on other computers.
-All the developer needs to do is to pull the git repository\todoadvisor{Cite the git repository} of this project onto the \emph{ARM} architecture, run a set up script to install dependencies, and then run the build script to compile everything into an immediately deployable OS image.
+All the developer needs to do is to pull the git repository of this project onto a machine running the \emph{ARM} architecture \cite{ProjectGitHub}, run a set up script to install dependencies, and then run the build script to compile everything into an immediately deployable OS image.
 This build script will automatically compile all the necessary code using \emph{cmake}, package all the necessary files using \emph{makepkg}, and build a fresh OS image that is ready to flash.
@@
 
-\todoadvisor{Explicitly mention that this is work I did}
 In order to evaluate the performance of the project, a suite of tools must be developed to test both the newly developed application and the existing solutions.
@@
 These statistics provide insight into how performant the application is from a technical perspective.
-The following is a structure that holds the statistics recorded over a session of connecting to a host computer.
+The following is a structure that was developed for this thesis to hold the statistics recorded over a session of connecting to a host computer.
+Corresponding code was written throughout the fork of the Moonlight application to populate this structure while a streaming session is running.
 The data recorded is discussed in Chapter \ref{Chapter6}.
@@
 This is done using the following steps once the client is connected to the host:
-\todoadvisor{Mention that the times need to be syncronized}
 
@@
 Because this method of data collection runs as a separate application on both the client and the server, it can be used to measure the performance of any of the remote desktop applications for comparison.
-This is again discussed in Chapter \ref{Chapter6}.
+An important obstacle to using this data is the difficulty in synchronizing the host and client machines' clocks.
+Even if each machine accurately recorded the amount of time between each event and the order in which they occur, the data is useless if the two machines do not agree at which time the test began.
+This is again discussed in Chapter \ref{Chapter6}, with the issue of clock synchronization being discussed in Section \ref{sec:TestingMethodology}.
 
@@
 
-This is sufficient to change the color of the screen back and forth between easily distinguishable colors for both the client and the host to record.
-\todoadvisor{Mention that this is recorded by the dataloggerz}
+This is sufficient to change the color of the screen back and forth between easily distinguishable colors for the data loggers of both the client and the host to record.
 With both of these datasets recorded, the next step is to analyze the data and evaluate the performance of the application.

### 2022-04-16 code/datalogger/linux/logger.cpp
Rework linux datalogger and add autotyping tool
This allows the testing to be automated and exact rather than based on human inperfection

@@ -17,11 +17,14 @@ Adapted from theju/linux-keylogger
 #include <chrono>
 #include <thread>
 #include <math.h>
+#include <map>
 #include <X11/Xlib.h>
 #include <X11/Xutil.h>
 
 #define MAX_EVENTS 100
 #define MAXKEYS 104
+
+#define MS_BETWEEN_KEYSTROKE 100
 
 using namespace std;
 
@@ -37,238 +40,158 @@ class KeyLogger
 private:
   int fd, efd, cfg, ret;
   char *temp;
-  ENTRY ep, *search_result;
-  struct input_event ev;
+  ENTRY ep;
+  struct input_event ev, inject_ev;
   struct epoll_event ee, *events;
-  struct code2keymap
+
+  string test_data = "this is a test string that is exactly 400 characters long that is being sent from client machine to the host computer in order to test the latency and response time. the computer is typing each character with a constant delay between each keystroke to ensure a consistent flow of data, this delay can be configured from test to test to evaluate the streaming application's abilities to transfer data.";
+
+
+  // Given a map from keys to values, creates a new map from values to keys
+  template <typename K, typename V>
+  static map<V, K> reverse_map(const map<K, V> &m)
   {
-    const char *code;
-    const char *keyname;
+    map<V, K> r;
+    for (const auto &kv : m)
+      r[kv.second] = kv.first;
+    return r;
+  }
+
+  map<int, string> keymap{
+      {0, "[RESERVED]"},
+      {1, "[ESC]"},
+      {2, "1"},
+      {3, "2"},
+      {4, "3"},
+      {5, "4"},
+      {6, "5"},
+      {7, "6"},
+      {8, "7"},
+      {9, "8"},
+      {10, "9"},
+      {11, "0"},
+      {12, "-"},
+      {13, "="},
+      {14, "[BACKSPACE]"},
+      {15, "[TAB]"},
+      {16, "q"},
+      {17, "w"},
+      {18, "e"},
+      {19, "r"},
+      {20, "t"},
+      {21, "y"},
+      {22, "u"},
+      {23, "i"},
+      {24, "o"},
+      {25, "p"},
+      {26, "["},
+      {27, "]"},
+      {28, "[RETURN]"},
+      {29, "[LEFT_CTRL]"},
+      {30, "a"},
+      {31, "s"},
+      {32, "d"},
+      {33, "f"},
[…]

### 2022-04-16 code/datalogger/windows/logger.cpp

@@ -18,7 +18,7 @@ Adapted from GiacomoLaw/Keylogger
 const std::map<int, std::string> keyname{
 		{VK_BACK, "[BACKSPACE]"},
 		{VK_RETURN, "[RETURN]"},
-		{VK_SPACE, "[SPACE]"},
+		{VK_SPACE, " "}, // Space
 		{VK_TAB, "[TAB]"},
 		{VK_SHIFT, "[SHIFT]"},
 		{VK_LSHIFT, "[LEFT_SHIFT]"},

### 2022-04-20 code/datalogger/linux/logger.cpp
Rewrite Data logger again for more consistent output
Also fix that annoying bug that caused the windows logger to randomly slow the entire computer to a crawl if left on too long

@@ -18,255 +18,90 @@ Adapted from theju/linux-keylogger
 #include <thread>
 #include <math.h>
 #include <map>
+#include <xdo.h>
 #include <X11/Xlib.h>
 #include <X11/Xutil.h>
+#include <X11/XKBlib.h>
 
-#define MAX_EVENTS 100
-#define MAXKEYS 104
-
-#define MS_BETWEEN_KEYSTROKE 100
+#define MS_BETWEEN_KEYSTROKE 500
 
 using namespace std;
 
 FILE *pFile;
 FILE *pFile2;
+Display *display;
 int keysInputCounter = 0;
 int colorsInputCounter = 0;
+string test_data = "this is a test string that is exactly 400 characters long that is being sent from client machine to the host computer in order to test the latency and response time. the computer is typing each character with a constant delay between each keystroke to ensure a consistent flow of data. this delay can be configured from test to test to evaluate the streaming application's abilities to transfer data.";
 
 void CTRL_C_Handler(int);
 
-class KeyLogger
+void type_string()
 {
-private:
-  int fd, efd, cfg, ret;
-  char *temp;
-  ENTRY ep;
-  struct input_event ev, inject_ev;
-  struct epoll_event ee, *events;
+  usleep(MS_BETWEEN_KEYSTROKE * 1000);
+  xdo_t *xdo = xdo_new(NULL);
+  xdo_enter_text_window(xdo, CURRENTWINDOW, test_data.c_str(), MS_BETWEEN_KEYSTROKE * 1000);
+  xdo_free(xdo);
+}
 
-  string test_data = "this is a test string that is exactly 400 characters long that is being sent from client machine to the host computer in order to test the latency and response time. the computer is typing each character with a constant delay between each keystroke to ensure a consistent flow of data, this delay can be configured from test to test to evaluate the streaming application's abilities to transfer data.";
+void log()
+{
+  Window root = DefaultRootWindow(display);
 
+  Window current_window;
+  int rev;
+  XGetInputFocus(display, &current_window, &rev);
 
-  // Given a map from keys to values, creates a new map from values to keys
[…]

### 2022-04-20 code/datalogger/windows/logger.cpp

@@ -18,7 +18,7 @@ Adapted from GiacomoLaw/Keylogger
 const std::map<int, std::string> keyname{
 		{VK_BACK, "[BACKSPACE]"},
 		{VK_RETURN, "[RETURN]"},
-		{VK_SPACE, " "}, // Space
+		{VK_SPACE, "space"},
 		{VK_TAB, "[TAB]"},
 		{VK_SHIFT, "[SHIFT]"},
 		{VK_LSHIFT, "[LEFT_SHIFT]"},
@@ -38,11 +38,11 @@ const std::map<int, std::string> keyname{
 		{VK_DOWN, "[DOWN]"},
 		{VK_PRIOR, "[PG_UP]"},
 		{VK_NEXT, "[PG_DOWN]"},
-		{VK_OEM_PERIOD, "."},
-		{VK_DECIMAL, "."},
-		{VK_OEM_PLUS, "+"},
+		{VK_OEM_PERIOD, "period"},
+		{VK_DECIMAL, "period"},
+		{VK_OEM_PLUS, "equal"},
 		{VK_OEM_MINUS, "-"},
-		{VK_ADD, "+"},
+		{VK_ADD, "equal"},
 		{VK_SUBTRACT, "-"},
 		{VK_CAPITAL, "[CAPSLOCK]"},
 };
@@ -103,11 +103,16 @@ int Save(int key_stroke)
 	std::stringstream output;
 	HKL layout = NULL;
 
-	if (key_stroke == VK_RETURN)
+	if (key_stroke == VK_OEM_PLUS)
 	{
-		// If Numpad 0 is hit, reset the counter for easy data comprehension
+		// If plus is hit, reset the counter for easy data comprehension
 		keysInputCounter = 0;
 		colorsInputCounter = 0;
+	}
+	if (key_stroke == VK_OEM_MINUS)
+	{
+		// If minus is hit, exit
+		exit(0);
 	}
 
 	if (keyname.find(key_stroke) != keyname.end())
@@ -155,10 +160,10 @@ void colorThread()
 	double prev_r, prev_g, prev_b = 0;
 	if (_hGDI)
 	{
+		HDC _hdc = GetDC(NULL);
 		while (true)
 		{
 			GETPIXEL pGetPixel = (GETPIXEL)GetProcAddress(_hGDI, "GetPixel");
-			HDC _hdc = GetDC(NULL);
 			if (_hdc)
 			{
 				// get center of screen
@@ -176,7 +181,8 @@ void colorThread()
 				{
 					// get time
[…]

### 2022-04-21 paper/Chapters/Evaluation.tex
Record new data and update evaluation chapter

@@
   \item GPU Stressed: The GPU is constantly at 100\% load during the entire test.
+  \item Both Stressed: Both the CPU and the GPU are constantly at 100\% load during the entire test.
 \end{itemize}
@@
   \item The client machine connects to the host computer using the application being tested.
-  \item The tester presses the \enquote{0} key on the client's numpad to synchronize the two devices' clocks and data recording tools (Detailed below).
-  \item The tester begins by typing on the client's keyboard at a steady pace as close to 50 words per minute as possible, allowing both computers to record their data.
-  \item The tester then types on the client's keyboard at a quicker pace of approximately 100 words per minute to ensure that quick successive inputs are handled correctly and no data is lost or duplicated.
+  \item The tester presses the \enquote{=} to begin the automated testing process.
         Due to the wide variance in what the user can do using the software, it is nearly impossible to test real world scenarios.
-  Instead, high frequency inputs coupled with updating the entire screen mimic high-intensity applications.
-  \item Once typing is complete, both the client and the host computers close their data recording tools and the client ends the streaming session.
+        Instead, the automated test simulates an active working environment by sending 400 keystrokes with a 500ms delay between inputs coupled with updating the entire screen for each keystroke to mimic high-intensity applications.
+  \item Once the trial is complete, both the client and the host computers close their data recording tools and the client ends the streaming session.
 \end{enumerate}
@@
 Each of the following figures shows data over one trial for each category.
-Each trial lasted approximately 60 seconds following the steps detailed in Section \ref{sec:TestingMethodology}, totaling to approximately 200 keystrokes per trial.
+Each trial consists of the client machine programmatically sending 400 keystrokes with 500ms between each input.
 Limitations in this data is discussed in Section \ref{sec:ConclusionLimitations}.
@@
         ymin=0,ymax=200,
-      cycle list={{black},{blue},{red}},
+        cycle list={{green!70!black},{red},{blue},{black}},
         legend style={
@@
             % Three boxplots for each column
-          draw position={1/4 + floor(\plotnumofactualtype/3) + 1/4*mod(\plotnumofactualtype,3)},
-          % Each plot takes up a quarter of the column
-          box extend=0.2,
+            draw position={1/5 + floor(\plotnumofactualtype/4) + 1/5*fpumod(\plotnumofactualtype,4)},
+            % Each plot takes up a 15% of the column
+            box extend=0.15,
           },
@@
         xticklabels={%
-      {Forked Moonlight\\{\tiny ideal/cpu/gpu}},%
-      {CRD\\{\tiny ideal/cpu/gpu}},%
-      {RDP\\{\tiny ideal/cpu/gpu}},%
-      {VNC\\{\tiny ideal/cpu/gpu}},%
+            {Forked Moonlight},%
+            {CRD},%
+            {RDP},%
+            {VNC},%
           },
@@
       \addplot
-      table[col sep=comma,x=Time difference] {Data/ideal/moonlight_key_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/ideal/moonlight_key_delay.csv};
       \addplot
-      table[col sep=comma,x=Time difference] {Data/stresscpu/moonlight_key_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/stresscpu/moonlight_key_delay.csv};
       \addplot
-      table[col sep=comma,x=Time difference] {Data/stressgpu/moonlight_key_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/stressgpu/moonlight_key_delay.csv};
+      \addplot
+      table[x expr=\coordindex+1, y index=0] {Data/stressboth/moonlight_key_delay.csv};
 
@@
       \addplot
-      table[col sep=comma,x=Time difference] {Data/ideal/crd_key_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/ideal/crd_key_delay.csv};
       \addplot
-      table[col sep=comma,x=Time difference] {Data/stresscpu/crd_key_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/stresscpu/crd_key_delay.csv};
       \addplot
-      table[col sep=comma,x=Time difference] {Data/stressgpu/crd_key_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/stressgpu/crd_key_delay.csv};
+      \addplot
+      table[x expr=\coordindex+1, y index=0] {Data/stressboth/crd_key_delay.csv};
 
@@
       \addplot
-      table[col sep=comma,x=Time difference] {Data/ideal/rdp_key_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/ideal/rdp_key_delay.csv};
       \addplot
-      table[col sep=comma,x=Time difference] {Data/stresscpu/rdp_key_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/stresscpu/rdp_key_delay.csv};
       \addplot
-      table[col sep=comma,x=Time difference] {Data/stressgpu/rdp_key_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/stressgpu/rdp_key_delay.csv};
+      \addplot
+      table[x expr=\coordindex+1, y index=0] {Data/stressboth/rdp_key_delay.csv};
 
@@
       \addplot
-      table[col sep=comma,x=Time difference] {Data/ideal/vnc_key_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/ideal/vnc_key_delay.csv};
       \addplot
-      table[col sep=comma,x=Time difference] {Data/stresscpu/vnc_key_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/stresscpu/vnc_key_delay.csv};
       \addplot
-      table[col sep=comma,x=Time difference] {Data/stressgpu/vnc_key_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/stressgpu/vnc_key_delay.csv};
+      \addplot
+      table[x expr=\coordindex+1, y index=0] {Data/stressboth/vnc_key_delay.csv};
 
-      \legend{Ideal,CPU stressed,GPU stressed}
+      \legend{Ideal,CPU stressed,GPU stressed, Both stressed}
     \end{axis}
   \end{tikzpicture}
-  \caption[Input Delay Data]{Time for host to press key from client.}
+  \caption[Input Delay Data]{Time for host to press key from client. n=400.}
   \label{fig:InputDelay}
@@
 If the user is going to be rendering video remotely, a GPU intensive process, the latency of the stream wont be impacted, but if the user is going to be running a CPU intensive process, the latency can be expected to be negatively impacted.
-RDP has a similar story, with extremely low latency while idle or under GPU load, and upwards of 40ms while under CPU load.
+RDP has a similar results, with extremely low latency while idle or under GPU load, and upwards of 40ms while under CPU load.
 VNC has the highest variance, with no discernable affinity to any kind of load.
 This lines up with it's intended use described in Section \ref{subsec:VirtualNetworkComputing}, being the simplest and most straightforward method of remote desktop control without much focus on performance or security.
-That's not so say VNC is unusable or even difficult to use; latency of 160ms is still usable in all but the most precise situations.
+That's not so say VNC is unusable or even difficult to use; latency of 140ms is still usable in all but the most precise situations.
 Though this is only one metric to consider, the other is the amount of time it take for the client to display the results of their input.
 This is recorded as the amount of time it takes between the host machine changing the color of it's screen and the client rendering the change.
+As noted before, the color of the host's screen changes with each input sent by the client, which occurs every 500ms for 400 keystrokes.
 Figure \ref{fig:ColorDelay} shows the amount of time for each application under each condition.
@@
         ymin=0,ymax=250,
-      cycle list={{black},{blue},{red}},
+        cycle list={{green!70!black},{red},{blue},{black}},
         legend style={
@@
       \addplot
-      table[col sep=comma,x=Time difference] {Data/ideal/moonlight_color_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/ideal/moonlight_color_delay.csv};
       \addplot
-      table[col sep=comma,x=Time difference] {Data/stresscpu/moonlight_color_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/stresscpu/moonlight_color_delay.csv};
       \addplot
-      table[col sep=comma,x=Time difference] {Data/stressgpu/moonlight_color_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/stressgpu/moonlight_color_delay.csv};
+      \addplot
+      table[x expr=\coordindex+1, y index=0] {Data/stressboth/moonlight_color_delay.csv};
 
@@
       \addplot
-      table[col sep=comma,x=Time difference] {Data/ideal/crd_color_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/ideal/crd_color_delay.csv};
       \addplot
-      table[col sep=comma,x=Time difference] {Data/stresscpu/crd_color_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/stresscpu/crd_color_delay.csv};
       \addplot
-      table[col sep=comma,x=Time difference] {Data/stressgpu/crd_color_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/stressgpu/crd_color_delay.csv};
+      \addplot
+      table[x expr=\coordindex+1, y index=0] {Data/stressboth/crd_color_delay.csv};
 
@@
       \addplot
-      table[col sep=comma,x=Time difference] {Data/ideal/rdp_color_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/ideal/rdp_color_delay.csv};
       \addplot
-      table[col sep=comma,x=Time difference] {Data/stresscpu/rdp_color_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/stresscpu/rdp_color_delay.csv};
       \addplot
-      table[col sep=comma,x=Time difference] {Data/stressgpu/rdp_color_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/stressgpu/rdp_color_delay.csv};
+      \addplot
+      table[x expr=\coordindex+1, y index=0] {Data/stressboth/rdp_color_delay.csv};
 
@@
       \addplot
-      table[col sep=comma,x=Time difference] {Data/ideal/vnc_color_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/ideal/vnc_color_delay.csv};
       \addplot
-      table[col sep=comma,x=Time difference] {Data/stresscpu/vnc_color_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/stresscpu/vnc_color_delay.csv};
       \addplot
-      table[col sep=comma,x=Time difference] {Data/stressgpu/vnc_color_delay.csv};
+      table[x expr=\coordindex+1, y index=0] {Data/stressgpu/vnc_color_delay.csv};
+      \addplot
+      table[x expr=\coordindex+1, y index=0] {Data/stressboth/vnc_color_delay.csv};
 
-      \legend{Ideal,CPU stressed,GPU stressed}
+      \legend{Ideal,CPU stressed,GPU stressed,Both stressed}
     \end{axis}
   \end{tikzpicture}
-  \caption[Color Delay Data]{Time for client to render update to the screen.}
+  \caption[Color Delay Data]{Time for client to render update to the screen. n=400.}
   \label{fig:ColorDelay}
@@
 CRD is able to render all updates in under 160ms while the host is idle, but struggles to keep up while under load.
-Again, CRD struggles the most while under CPU load, with the quickest update happening in just under 200ms, and the median update taking over a full second.
+Again, CRD struggles the most while under CPU load, with the quickest update happening in just under 180ms, and the median update taking over a full second.
 CRD performs a little better under GPU load, but still experiences much high delays with the median update taking over half of a second.
-Both of these values push the usability of the system as any kind of intensive program will bring the responsivity of the application to a crawl.
+Once both the CPU and the GPU are under load, the worst of both scenarios is realized as the median update taking nearly three quarters of a second.
+All three of these values push the usability of the system as any kind of intensive program will bring the responsivity of the application to a crawl.
 RDP again performs extremely well, with all render updates taking less than 50ms while under ideal conditions or GPU load, and all updates taking less that 70ms while under full CPU load.
@@
         ymin=0,ymax=100,
-      cycle list={{black},{blue},{red}},
+        cycle list={{green!70!black},{red},{blue},{black}},
         legend style={
@@
       ]
-      \addplot[style={black,fill=black,mark=none}]
+
+      \addplot[style={green!70!black,fill=green!70!black,mark=none}]
       coordinates {(0, 0) (1,0) (2,0) (3,15.32)};
 
-      \addplot[style={blue,fill=blue,mark=none}]
+      \addplot[style={red,fill=red,mark=none}]
       coordinates {(0,0) (1,32.28) (2,0) (3,19.51)};
 
-      \addplot[style={red,fill=red,mark=none}]
+      \addplot[style={blue,fill=blue,mark=none}]
       coordinates {(0,0) (1,2.34) (2,0) (3,15.5)};
 
-      \legend{Ideal,CPU stressed,GPU stressed}
+      \addplot[style={black,fill=black,mark=none}]
+      coordinates {(0, 0) (1,34.65) (2,0) (3,39.83)};
+
+      \legend{Ideal,CPU stressed,GPU stressed, Both stressed}
     \end{axis}
   \end{tikzpicture}
-  \caption[Color Delay Loss Data]{Percentage of color changes not rendered.}
+  \caption[Color Delay Loss Data]{Percentage of color changes not rendered. n=400.}
   \label{fig:ColorDelayLoss}

### 2022-04-24 paper/Chapters/Evaluation.tex
Write real world evaluation section

@@
 
+\section{Real World Testing}\label{sec:RealWorldTesting}
+
+While the testing conducted in the previous two sections provide an overview of how each application performs under controlled conditions, testing in the real world provides better insight into how a user may experience the final product.
+Unfortunately, there are many limitations in how testing can be conducted in the real world.
+First and foremost, due to the closed nature of of the existing solutions, collecting data while using them in real-world scenarios is impossible while leveraging the tools from the previous sections.
+Because of the reliance on using exact and repeatable testing conditions, a dynamic test of each application is much more difficult to conduct without direct access to the performance of the program, all data that could be collected otherwise would amount to quantitative data subject to interpretation.
+However, because of the logging tools that were developed in Section \ref{sec:DevelopingTestingAndMeasurementTools}, it is possible to collect data from the forked version of moonlight that shows the performance of the application over an entire streaming session.
+While it is impossible to predict and exhaustively test what the user may try to do with the final product, it is possible to test common use cases and evaluate whether the streaming experience is satisfactory.
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}{1\textwidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/realworld/kicad}
+  \end{subfigure}
+  \begin{subfigure}{1\textwidth}
+    \centering
+    \includegraphics[width=.5\linewidth]{Figures/realworld/kicadstats}
+  \end{subfigure}
+  \caption[Streaming Electronics Design software]{Working on the PCB for this project using KiCad.}
+  \label{fig:RealWorldKicad}
+\end{figure}
+
+The first use case that was tested was streaming KiCad to the client machine to inspect the PCB layout of the custom board (Figure \ref{fig:RealWorldKicad}).
+This was to test the ability to work with highly accurate CAD programs to ensure that the small client device could be used for technical work.
+The first thing that was noticed was that since the chosen screen for the client device was small, everything had to be zoomed in much further than normal to be comfortable.
+This is where the secondary HDMI output could be used to plug in a larger, more conventional monitor, but in the event the user did not have access to such a hardware, the small screen needed to be sufficient.
+Once zoomed in, KiCad was readable and usable to without any further compromise.
+The client's inputs were responsive and accurate enough for precise movements to be made without error and without the worry of needing to wait for the screen to catch up to render the changes.
+Once the five minute testing session was complete, the following statistics were recorded:
+
+\begin{lstlisting}[style=plaintext,title=Statistics recorded while streaming KiCad (Figure \ref{fig:RealWorldKicad})]
+Global Video Stats
+----------------------------------------------------------
+Incoming frame rate from network: 61.23 FPS
+Decoding frame rate: 61.23 FPS
+Rendering frame rate: 61.23 FPS
+Frames dropped by your network connection: 0.00%
+Average network latency: 1 ms (variance: 0 ms)
+Average decoding time: 4.24 ms
+Average rendering time (including monitor V-sync latency): 0.06 ms
+
+Total dropped frames: 0
+\end{lstlisting}
+
+These results are very similar to the results recorded in the previous sections, where the frame rate is able to be kept above 60 FPS throughout the test, and zero frames were dropped.
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}{1\textwidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/realworld/ai}
+  \end{subfigure}
+  \begin{subfigure}{1\textwidth}
+    \centering
+    \includegraphics[width=.5\linewidth]{Figures/realworld/aistats}
+  \end{subfigure}
+  \caption[Streaming the training of an AI]{Programing while training an AI model.}
+  \label{fig:RealWorldAI}
+\end{figure}
+
+The next use case that was tested was programing while training an AI model (Figure \ref{fig:RealWorldAI}).
+While programming on it's own is not a very intensive process, and processes such as compiling can be done while the user isn't actively interacting with the computer, training an AI model can be a very time consuming process that may need to continue in the background while the user continues to do other work.
+As this scenario is extremely similar to the CPU stress test from Section \ref{sec:ResponsivityAndLatency}, the same results are expected.
+
+\begin{lstlisting}[style=plaintext,title=Statistics recorded while training an AI (Figure \ref{fig:RealWorldAI})]
+Global Video Stats
+----------------------------------------------------------
+Incoming frame rate from network: 62.05 FPS
+Decoding frame rate: 62.05 FPS
+Rendering frame rate: 62.05 FPS
+Frames dropped by your network connection: 0.00%
+Average network latency: 1 ms (variance: 0 ms)
+Average decoding time: 4.26 ms
+Average rendering time (including monitor V-sync latency): 0.05 ms
+
+Total dropped frames: 0
+\end{lstlisting}
+
+As with before, the stream remains performant and usable throughout the test.
+There was no point during the test that the stream started to struggle as the AI training process kept the CPU at 100\% usage.
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}{1\textwidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/realworld/blender}
+  \end{subfigure}
+  \begin{subfigure}{1\textwidth}
+    \centering
+    \includegraphics[width=.5\linewidth]{Figures/realworld/blenderstats}
+  \end{subfigure}
+  \caption[Streaming a 3D modeling, animation, and rendering program]{Working on a blender animation while rendering the scene.}
+  \label{fig:RealWorldBlender}
+\end{figure}
+
+The next use case that was tested was streaming the 3D modeling, animation, and rendering program Blender (Figure \ref{fig:RealWorldBlender}).
+To test this, a demo scene provided for free by the Blender Foundation was loaded and edited in one window while being rendered from the camera's point of view in real time in another window \cite{BlenderDemoScene}.
+This put inconsistent strain on the computer as it constantly had to rerender the scene as changes were made.
+Whenever the scene was rendered to a reasonable resolution, Blender would automatically stop rendering to keep resource usage low, resulting in the CPU and GPU usage to bounce back and forth between high and low usage.
+This proved to be the most strenuous test of the application:
+
+\begin{lstlisting}[style=plaintext,title=Statistics recorded while streaming Blender (Figure \ref{fig:RealWorldBlender})]
+  Global Video Stats
+----------------------------------------------------------
+Incoming frame rate from network: 57.30 FPS
+Decoding frame rate: 57.30 FPS
+Rendering frame rate: 57.30 FPS
+Frames dropped by your network connection: 0.06%
+Average network latency: 1 ms (variance: 0 ms)
+Average decoding time: 4.70 ms
+Average rendering time (including monitor V-sync latency): 0.05 ms
+
+Total dropped frames: 16
+\end{lstlisting}
+
+This was the only test in which not all frames rendered successfully, though the majority 99.4\% of frames rendered as expected.
+In fact, due to the stream running at an average 57 FPS, the 0.06\% loss of frames was imperceivable during the test and only became known once the statistics were produced.
+During the test, the stream remained performant and stable.
+Though applications such as this are more difficult to use on a smaller screen, it was by no means unusable; and because of the minimal delay from the stream, there was no point during the test where it felt like the output of the tester's inputs were not visible as quick as they normally would.
+
+\begin{figure}[t]
+  \centering
+  \begin{subfigure}{1\textwidth}
+    \centering
+    \includegraphics[width=.9\linewidth]{Figures/realworld/gamedev}
+  \end{subfigure}
+  \begin{subfigure}{1\textwidth}
+    \centering
+    \includegraphics[width=.5\linewidth]{Figures/realworld/gamedevstats}
+  \end{subfigure}
+  \caption[Streaming the development of a video game level]{Developing a video game level while inspecting it in game.}
+  \label{fig:RealWorldGameDev}
+\end{figure}
+
+The last use case that was tested was streaming the development of a video game level while inspecting in the game (Figure \ref{fig:RealWorldGameDev}).
+This was intended to be an general all-round stress test that would put the host under multiple kinds of stress at the same time.
+Since the host computer had to run not only the video game's development tools but also the game itself, the stream had to keep up with the precise nature of geometric level design as well as interacting with the game.
+During the test, the client machine always felt as though it was not compromising the experience of working on the host machine save for the small display; inputs in the development tools stayed precise and accurate, and interacting with the game was as smooth as playing on the host machine.
+Once the test concluded, the following statistics were recorded:
+
+\begin{lstlisting}[style=plaintext,title=Statistics recorded while streaming the development of a video game level (Figure \ref{fig:RealWorldGameDev})]
+Global Video Stats
+----------------------------------------------------------
+Incoming frame rate from network: 61.93 FPS
+Decoding frame rate: 61.93 FPS
+Rendering frame rate: 61.93 FPS
+Frames dropped by your network connection: 0.00%
+Average network latency: 1 ms (variance: 0 ms)
+Average decoding time: 4.27 ms
+Average rendering time (including monitor V-sync latency): 0.06 ms
+
+Total dropped frames: 0
+\end{lstlisting}
+
+As with the most of the previous tests, the stream kept a constant frame rate above 60 FPS and never dropped any frames.
+
 \section{Summary}\label{sec:EvaluationSummary}
 
-By combining the information from the previous two sections, a wholistic picture of the performance of each application can be drawn.
+By combining the information from the previous three sections, a wholistic picture of the performance of each application can be drawn.
 Firstly, Virtual Network Computing works as a bare-bones method of connecting to a host machine from a client device in the simplest way possible, but struggles at keeping up with even mildly demanding tasks.

### 2022-04-24 paper/Chapters/Background.tex
Implement Director feedback
Everything except for summaries for each section and a last look-over

@@
 This allows the user to utilize the power of the host machine from a cheap computer that usually doesn't have enough power to be used as a standalone computer.
-Because of it's low power, the thin client cannot be used as a standalone computer, but it's low cost and portability make it a great solution for a corporate user.
+Because of this, the thin client cannot be used as a standalone computer, but it's low cost and portability make it a great solution for a corporate user.
 While these clients work well for simple interfacing tasks such as data entry or system management, they are not designed to be used for complex tasks that require quick response times or graphical fidelity.

### 2022-04-24 paper/Chapters/Conclusion.tex
Implement Director feedback
Everything except for summaries for each section and a last look-over

@@
 This thesis has proved that it is possible to produce a hardware and software solution that is capable of leveraging the power of a desktop computer remotely without sacrificing the user's experience.
-It improves upon existing solutions by continuing to offer a performant and high-quality stream of the host's screen and audio to the user even as the host is running highly demanding applications.
+It improves upon existing solutions by continuing to offer a performant and high-quality stream of the host's screen to the user even as the host is running highly demanding applications.
 Each of the questions posed in Section \ref{sec:ResearchQuestions} can be answered individually.
@@
 This thesis project is able to run on embedded Single Board Computers and hardware available for purchase today.
-While this project built a custom PCB for the Raspberry Pi Compute Module 4 (Discussed in section \ref{sec:ChoosingParts}), the software could also run on consumer-grade Raspberry Pi models as well.
+While this project built a custom PCB for the Raspberry Pi Compute Module 4 (Discussed in section \ref{sec:ChoosingParts}), the software could also run on consumer-grade main-line Raspberry Pi models as well.
 

### 2022-04-24 paper/Chapters/Evaluation.tex
Implement Director feedback
Everything except for summaries for each section and a last look-over

@@
 The testing methodology employed to ensure the accuracy of the testing is described in Section \ref{sec:TestingMethodology}.
-Then, the different results are evaluated and analyzed in Sections \ref{sec:ResponsivityAndLatency} and \ref{sec:Quality}.
+Then, the responsivity and latency of each solution is evaluated in Section \ref{sec:ResponsivityLatency}, and the quality of each solution's stream is evaluated in Section \ref{sec:Quality}.
+Then, the forked version of Moonlight is evaluated in real world scenarios in Section \ref{sec:RealWorldTesting}.
 Finally, the results are summarized and verdicts on each of the solutions' applicability to the research questions will be presented in Section \ref{sec:EvaluationSummary}.
@@
   \item If the test calls for the CPU or the GPU to be stressed, the application \enquote{Blender} is started on the host machine, and a benchmark is performed for the CPU or the GPU.
-        This benchmark will run for longer than the duration of the test, and will keep the CPU or GPU at 100\% load for the entire duration of the benchmark.
-  \item Both devices start up their respective data collection tool that will record key-strokes and changes in the screen's color.
+        This benchmark will run for longer than the duration of the test, and will keep the CPU and/or GPU at 100\% load for the entire duration of the benchmark.
+  \item Both devices start up their respective data collection tool that will record keystrokes and changes in the screen's color.
   \item The client machine connects to the host computer using the application being tested.
-  \item The tester presses the \enquote{=} to begin the automated testing process.
-        Due to the wide variance in what the user can do using the software, it is nearly impossible to test real world scenarios.
-        Instead, the automated test simulates an active working environment by sending 400 keystrokes with a 500ms delay between inputs coupled with updating the entire screen for each keystroke to mimic high-intensity applications.
+  \item The tester presses the \enquote{=} key on the client machine to begin the automated testing process.
+        Due to the wide variance in what the user can do using the software, it is nearly impossible to test every possible real world scenario.
+        Instead, the automated test simulates an active working environment by sending a repeatable series of keystrokes that cause the entire screen to change colors to mimic high-intensity applications.
+        For consistent testing, 400 keystrokes are automatically sent with a 500ms delay between inputs.
   \item Once the trial is complete, both the client and the host computers close their data recording tools and the client ends the streaming session.
@@
 This is partially accounted for in step \emph{6} of the above process, and the remaining inaccuracy can now be removed by subtracting the time difference between the when the client presses the button and the time the host executes that command.
+It is important to not subtract the entire round trip time of the input, as that is what is being testing in the following section, but rather just the time between when the client sends the input and when the host receives it.
 This ensures that any inaccuracies in each computer's internal clock will not impact the data.
 
-After the data is cleaned up, the data can be processed by matching each key-stroke and change in the screen's color detected by the client and host machines and calculating how much time elapsed between the events.
+After the data is cleaned up, the data can be processed by matching each keystroke and change in the screen's color detected by the client and host machines and calculating how much time elapsed between the events.
 This reveals the amount of delay the user experiences between when they press a key and when the host executes the command, and between when they press a key and can visually see the response.
@@
 To analyze this, two points of data are recorded: First, the amount of time between when the client records a keystroke and when it is actually inputted on the host machine, and secondly, the amount of time between the host machine updating their screen and when the client actually displays that update.
-The first datum is generally referred to as the latency of the application, since even though each trial is done using the same network setup different programs may have different optimizations built in that reduce the time needed for the host to implement keystrokes made by the client.
+The first datum is generally referred to as the latency of the application, since even though each trial is done using the same network setup different solutions may have different optimizations built in that reduce the time needed for the host to implement keystrokes made by the client.
+This means that as the network delay is removed from the data, all that is left is the time that it takes the software on the host's machine to actually execute the input.
 Each of the following figures shows data over one trial for each category.
@@
 Limitations in this data is discussed in Section \ref{sec:ConclusionLimitations}.
-Figure \ref{fig:InputDelay} shows the amount of time between the client and the host registering the keystroke under the different test conditions.
-
 \begin{figure}[h]
@@
 \end{figure}
+Figure \ref{fig:InputDelay} shows the amount of time between the client and the host registering the keystroke under the different test conditions.
 
@@
 That's not so say VNC is unusable or even difficult to use; latency of 140ms is still usable in all but the most precise situations.
+
 Though this is only one metric to consider, the other is the amount of time it take for the client to display the results of their input.
@@
   \end{tikzpicture}
-  \caption[Color Delay Data]{Time for client to render update to the screen. n=400.}
+  \caption[Color Delay Data]{Time for client to render update to the screen. Values over 250ms push the usability of the system. n=400.}
   \label{fig:ColorDelay}
@@
 
-\begin{figure}[t]
+\begin{figure}[h]
   \centering
@@
 Apart from the performance of each application, the quality of the user experience must also be taken into account when considering each application.
-Each application will be assessed individually, and then compared in Section \ref{sec:EvaluationSummary}
+Each application will be assessed individually, and then compared in Section \ref{sec:EvaluationSummary}.
+Due to the closed nature of some of the protocols, only limited amounts of data could be extracted and compared.
+This prevents direct comparisons, but can still contribute to the understanding of the overall performance of the solution.
+
+\subsection{Moonlight}\label{subsec:QualityMoonlight}
+
 When analyzing the forked version of Moonlight, specific quantitative data described in Section \ref{sec:DevelopingTestingAndMeasurementTools} can be collected directly from the application while it is running.
@@
 Average rendering time (including monitor V-sync latency): 0.06 ms
-
 Total dropped frames: 0
@@
 There were very few times where the screen appeared to be run through a compression filter, and it for the most part kept the visual fidelity of the host machine.
+
+\subsection{Chrome Remote Desktop}\label{subsec:QualityCRD}
 
@@
 
+\subsection{Remote Desktop Protocol}\label{subsec:QualityRDP}
+
 Microsoft's Remote Desktop Protocol does not provide quantitative data to analyze it's quality, but speaking from a user experience perspective, and considering the data provided in the previous section, RDP performed very well under each of the testing conditions.
@@
 This will be further discussed in Section \ref{sec:ConclusionLimitations}.
+
+\subsection{Virtual Network Computing}\label{subsec:QualityVNC}
 
@@
 
-While the testing conducted in the previous two sections provide an overview of how each application performs under controlled conditions, testing in the real world provides better insight into how a user may experience the final product.
-Unfortunately, there are many limitations in how testing can be conducted in the real world.
-First and foremost, due to the closed nature of of the existing solutions, collecting data while using them in real-world scenarios is impossible while leveraging the tools from the previous sections.
-Because of the reliance on using exact and repeatable testing conditions, a dynamic test of each application is much more difficult to conduct without direct access to the performance of the program, all data that could be collected otherwise would amount to quantitative data subject to interpretation.
-However, because of the logging tools that were developed in Section \ref{sec:DevelopingTestingAndMeasurementTools}, it is possible to collect data from the forked version of moonlight that shows the performance of the application over an entire streaming session.
-While it is impossible to predict and exhaustively test what the user may try to do with the final product, it is possible to test common use cases and evaluate whether the streaming experience is satisfactory.
-
-\begin{figure}[t]
+\begin{figure}[!b]
   \centering
@@
   \end{subfigure}
-  \caption[Streaming Electronics Design software]{Working on the PCB for this project using KiCad.}
+  \caption[Streaming Electronics Design software]{Working on the PCB for this thesis using KiCad.}
   \label{fig:RealWorldKicad}
 \end{figure}
+
+While the testing conducted in the previous two sections provide an overview of how each application performs under controlled conditions, testing in the real world provides better insight into how a user may experience the final product.
+Unfortunately, there are many limitations in how testing can be conducted in the real world.
+First and foremost, due to the closed nature of of the existing solutions, collecting data while using them in real-world scenarios is impossible while leveraging the tools from the previous sections.
+Because of the reliance on using exact and repeatable testing conditions, a dynamic test of each application is much more difficult to conduct without direct access to the performance of the program, all data that could be collected otherwise would amount to quantitative data subject to interpretation.
+However, because of the logging tools that were developed in Section \ref{sec:DevelopingTestingAndMeasurementTools}, it is possible to collect data from the forked version of moonlight that shows the performance of the application over an entire streaming session.
+While it is impossible to predict and exhaustively test what the user may try to do with the final product, it is possible to test common use cases and evaluate whether the streaming experience is satisfactory.
 
@@
 Average rendering time (including monitor V-sync latency): 0.05 ms
-
 Total dropped frames: 0
@@
 Average rendering time (including monitor V-sync latency): 0.05 ms
-
 Total dropped frames: 16

### 2022-04-24 paper/Chapters/Hardware.tex
Implement Director feedback
Everything except for summaries for each section and a last look-over

@@
 
-This section will explain the different constraints taken into account while developing the hardware portion of this project.
+This section will explain the different constraints taken into account while developing the hardware portion of this thesis.
 First, the hardware must be performant enough to provide a positive user experience (\ref{subsec:HardwarePerformance}), and secondly the product must be cheap enough to justify it over similar alternatives (\ref{subsec:HardwareCost}).
 
-Due to the lack of preliminary studies or research into the topic of target audience for this project, the requirements laid out in this section are based on the author's own requirements and preferences.
+Since no specific studies regarding possible users of the system were conducted, the requirements laid out in this section are based on the author's own requirements and preferences.
 The requirements are constructed with the target user being someone who already has a working desktop computer and wishes to leverage it's power remotely without sacrificing the user experience.
@@
 Instead, a better suited solution would be to utilize a \emph{Raspberry Pi Compute Module 4 (CM4)} which has \enquote{The power of Raspberry Pi 4 in a compact form factor for deeply embedded applications} \cite{rpi_cm4}.
-Boasting an incredibly small size and streamlined I/O, there only way to communicate with the Compute Module was through two tiny mezzanine connectors as shown in Figure \ref{fig:rpi_cm4_mezzanine}.
-With no other pin outs for anything such as power, HDMI, or USB, everything had to be built on a custom Printed Circuit Board (PCB).
+As this perfectly lines up with the requirements for the project, this thesis will be built using the CM4.
+However, due to it's small size and deeply embedded nature, the only way to communicate with the Compute Module is through two tiny mezzanine connectors as shown in Figure \ref{fig:rpi_cm4_mezzanine}.
+With no other pin outs for anything such as power, HDMI, or USB, the rest of the hardware must be built on a custom Printed Circuit Board (PCB) that tailors the CM4 for this thesis.
 

### 2022-04-24 paper/Chapters/Introduction.tex
Prepare Complete (not final) draft
This draft will be sent of to each of the committee members to read through before the thesis defense.
The PDF has been rendered in "final" mode, so no comments will show through.

small.pdf is a compressed version of the main pdf, with no differences except each of the images have been compressed to 300dpi.
Implement Director feedback
Everything except for summaries for each section and a last look-over

@@
-\chapter{Introduction} % Main chapter title
+\chapter{Introduction}
 
-\label{Chapter1} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}
+\label{Chapter1}
 
@@
 The primary goal of this thesis is to create a portable hardware and software solution that can be used to control a desktop computer remotely while leveraging all of it's power.
-There are existing software solutions that give a user remote control of their desktop computer from a different location, but many of them fail to provide a high-quality and performant solution to working with highly demanding programs that put the desktop computer under intense load.
-Once attempting to run an intensive application such as rendering video, training Artificial Intelligence, or developing a video game, existing solutions fail to keep up.
+There are existing software solutions that give a user remote control of their desktop computer from a different location, which will be introduced in Chapter \ref{Chapter3}, but such solutions may not provide a performant and high-quality  experience for the user.
+These solutions will be tested while running highly demanding applications such as rendering animation, training Artificial Intelligence models, or developing video games to see if they are capable of working in such situations.
 

### 2022-04-24 paper/Chapters/Software.tex
Implement Director feedback
Everything except for summaries for each section and a last look-over

@@
 To accomplish this, a fork of the Moonlight project is created so the GameStream protocol can be leveraged in this thesis.
-With the fork, much of the code required to create the software solution is already provided, all that is left is to tweak the code to ease development for the \emph{ARM} architecture and to build a suite of testing tools that hook into Moonlight to extract statistics.
+With the fork, much of the code required to create the software solution is already provided, the next step is to build on top of the existing code to ease development for the \emph{ARM} architecture and to write a suite of testing tools that hook into Moonlight to extract statistics.
 
@@
 Corresponding code was written throughout the fork of the Moonlight application to populate this structure while a streaming session is running.
-The data recorded is discussed in Chapter \ref{Chapter6}.
+The data recorded is evaluated in Chapter \ref{Chapter6}.
 

### 2022-04-24 paper/main.tex
Feedback from advisor and thesis defense
Record new data and update evaluation chapter
Implement Director feedback
Everything except for summaries for each section and a last look-over

@@
 % draft, % Uncomment to enable draft mode (no pictures, no links, overfull hboxes indicated)
-final, % Uncomment to enable final mode (hide todos)
+% final, % Uncomment to enable final mode (hide todos)
 %nolistspacing, % If the document is onehalfspacing or doublespacing, uncomment this to set spacing in lists to single
@@
 \pgfplotsset{compat=1.8}
+% Fix mod function bug
+% borrowed from <https://tex.stackexchange.com/a/145967/95441>
+\pgfmathdeclarefunction{fpumod}{2}{%
+		\pgfmathfloatdivide{#1}{#2}%
+		\pgfmathfloatint{\pgfmathresult}%
+		\pgfmathfloatmultiply{\pgfmathresult}{#2}%
+		\pgfmathfloatsubtract{#1}{\pgfmathresult}%
+		% replaced `0' by `5' to make it work for this problem
+		\pgfmathfloatifapproxequalrel{\pgfmathresult}{#2}{\def\pgfmathresult{5}}{}%
+}
 
@@
 			\begin{flushright} \large
-				\emph{Supervisor:} \\
+				\emph{Director:} \\
 				\href{https://isearch.asu.edu/profile/3180015}{\supname} % Supervisor name - remove the \href bracket to remove the link  
+			\end{flushright}
+		\end{minipage}\\[.5cm]
+
+		\begin{minipage}[t]{0.8\textwidth}
+			\begin{flushright} \large
+				\emph{Second Reader:}\\
+				\href{https://isearch.asu.edu/profile/1232839/}{Ruben Acu\~{n}a}
+			\end{flushright}
+		\end{minipage}\\[.5cm]
+		\begin{minipage}[t]{0.8\textwidth}
+			\begin{flushright} \large
+				\emph{Third Reader:} \\
+				\href{https://isearch.asu.edu/profile/1625795}{Dr. Shawn Jordan}
 			\end{flushright}
@@
 
-	\noindent Signed:\\
+	\noindent Signed: Dylan Lathrum\\
 	\rule[0.5em]{25em}{0.5pt} % This prints a line for the signature
 
-	\noindent Date:\\
+	\noindent Date: 4/15/2022\\
 	\rule[0.5em]{25em}{0.5pt} % This prints a line to write the date
@@
 
+This thesis will explore the development of a custom PCB that utilizes a Raspberry Pi Computer Module 4, as well as the development of a fork of the Open Source project Moonlight to stream a host machine's screen to a remote client.
+This implementation will be compared against other existing remote desktop solutions to analyze it's performance and quality.
+
 \end{abstract}
@@
 	I would like to express my graditude to my thesis director, Dr.
-Robert Heinrichs, and to each of my committee members, Professor Ruben Acu\~{n}a and Dr.
+Robert Heinrichs, and to each of my committee members, Ruben Acu\~{n}a and Dr.
 Shawn Jordan, for their help in the development of this thesis.
@@
 	\addchaptertocentry{\abbrevname} % Add the Abbreviations to the table of contents
+	\negmedspace % Gets rid of weird extra space
 	\textbf{BOM} & \textbf{B}ill \textbf{O}f \textbf{M}aterials\\
@@
 
-% \include{Chapters/Example}
 \include{Chapters/Introduction}
@@
 \include{Appendices/AppendixB}
-%\include{Appendices/AppendixC}
+\include{Appendices/AppendixC}
 

### 2022-04-25 paper/Chapters/Evaluation.tex
Add summaries

@@
 The testing methodology employed to ensure the accuracy of the testing is described in Section \ref{sec:TestingMethodology}.
-Then, the responsivity and latency of each solution is evaluated in Section \ref{sec:ResponsivityLatency}, and the quality of each solution's stream is evaluated in Section \ref{sec:Quality}.
+Then, the responsivity and latency of each solution is evaluated in Section \ref{sec:ResponsivityAndLatency}, and the quality of each solution's stream is evaluated in Section \ref{sec:Quality}.
 Then, the forked version of Moonlight is evaluated in real world scenarios in Section \ref{sec:RealWorldTesting}.

### 2022-04-25 paper/Chapters/Hardware.tex
Add summaries

@@
 
+As discussed in Chapter \ref{Chapter2}, there currently is no commercial solution for a cheap, portable device that can leverage the power of a remote computer; hence, this thesis will investigate the development of such a solution.
 There are a lot of factors that go into designing and building an electrical system that fits the needs for this project.
@@
 
-\begin{figure}[b]
+The second board was assembled mostly successfully, and allowed the first stage of testing to begin.
+The only issue apart from some ports being too close together with real-world cables was one component having the wrong footprint on the board.
+Unfortunately, the component \emph{U2 (MCP73871-2CCI\_ML)}, is described as using the \emph{QFN-20-1EP\_4x4mm} footprint in its datasheet \cite{MCP73871-2CCI/ML-ND_datasheet}, but actually uses a revised version of the footprint, shown in Figure \ref{fig:FootprintQFN204x4}.
+This discrepancy meant that that component could not be soldered, and it's associated circuit would not function properly.
+In this case, the component was a part of the battery charging circuit, causing the entire primary power supply to be unusable.
+Thankfully, as described in Section \ref{sec:Prototyping}, the secondary method of powering the board could be used to test the remaining features of the board.
+After making the necessary fixes, a second round of five PCB boards and a stencil was ordered.
+Because there was still enough components to assemble three more boards, and no components needed to be changed, no more components were ordered at this stage.
+Once the boards were delivered, the revised board was again sent to be professionally assembled.
+
+\begin{figure}[t]
   \centering
@@
 
-The second board was assembled mostly successfully, and allowed the first stage of testing to begin.
-The only issue apart from some ports being too close together with real-world cables was one component having the wrong footprint on the board.
-Unfortunately, the component \emph{U2 (MCP73871-2CCI\_ML)}, is described as using the \emph{QFN-20-1EP\_4x4mm} footprint in its datasheet \cite{MCP73871-2CCI/ML-ND_datasheet}, but actually uses a revised version of the footprint, shown in Figure \ref{fig:FootprintQFN204x4}.
-This discrepancy meant that that component could not be soldered, and it's associated circuit would not function properly.
-In this case, the component was a part of the battery charging circuit, causing the entire primary power supply to be unusable.
-Thankfully, as described in Section \ref{sec:Prototyping}, the secondary method of powering the board could be used to test the remaining features of the board.
-After making the necessary fixes, a second round of five PCB boards and a stencil was ordered.
-Because there was still enough components to assemble three more boards, and no components needed to be changed, no more components were ordered at this stage.
-Once the boards were delivered, the revised board was again sent to be professionally assembled.
-
 \subsection{Third Attempt}\label{subsec:Manufacturing3}
 
-\begin{figure}[t]
+\begin{figure}[!b]
   \centering
@@
 Figure \ref{fig:AssembledPCB} shows the final assembled board, with the final PCB layouts in Figures \ref{app:pcb_layout_front} and \ref{app:pcb_layout_back}.
+
+\section{Summary}\label{sec:HardwareSummary}
+
+In order to produce a hardware device capable of fitting the needs of the project, a careful balance between power and cost had to be found.
+The Raspberry Pi Compute Module 4 offered enough power to power the project while maintaining a low cost and slim profile that enabled it to fit in a small mobile device.
+One the CM4 was chosen, a custom PCB had to be designed and manufactured to tailor the Compute module to the needs of the project.
+This included choosing compatible electronic components, designing a board that provided the necessary functionality, and developing multiple iterations of the board to fix problems as they arose.

### 2022-04-26 paper/Chapters/Evaluation.tex
Changes from feedback
Almost done

@@
   \end{tikzpicture}
-  \caption[Color Delay Data]{Time for client to render update to the screen. Values over 250ms push the usability of the system. n=400.}
+  \caption[Color Delay Data]{Time for client to render update to the screen. Values over 250ms indicate a poorly performing system. n=400.}
   \label{fig:ColorDelay}
@@
 
-\begin{figure}[!b]
+While the testing conducted in the previous two sections provide an overview of how each application performs under controlled conditions, testing in the real world provides better insight into how a user may experience the final product.
+Unfortunately, there are many limitations in how testing can be conducted in the real world.
+First and foremost, due to the closed nature of of the existing solutions, collecting data while using them in real-world scenarios is impossible while leveraging the tools from the previous sections.
+Because of the reliance on using exact and repeatable testing conditions, a dynamic test of each application is much more difficult to conduct without direct access to the performance of the program, all data that could be collected otherwise would amount to quantitative data subject to interpretation.
+However, because of the logging tools that were developed in Section \ref{sec:DevelopingTestingAndMeasurementTools}, it is possible to collect data from the forked version of Moonlight that shows the performance of the application over an entire streaming session.
+While it is impossible to predict and exhaustively test what the user may try to do with the final product, it is possible to test common use cases and evaluate whether the streaming experience is satisfactory.
+
+Each of the following subsections aims to represent a different use case that may be encountered in the real world.
+First, the project is tested with a CAD application in Section \ref{subsec:RealWorldKiCad}, representing precise graphical design using software that may struggle to run on portable hardware.
+Then, to represent a programming environment while utilizing the power of the host, the project is tested while training an AI model in Section \ref{subsec:RealWorldAI}.
+Then, the project is tested while using Blender to edit and render a 3D scene simultaneously to represent artistic workloads and GPU accelerated rendering in Section \ref{subsec:RealWorldBlender}.
+Finally, to represent an all around stressful situation for the host, the project is tested while developing a level for a video game while inspecting it in-game in Section \ref{subsec:RealWorldVideoGame}.
+
+\subsection{CAD}\label{subsec:RealWorldKiCad}
+
+\begin{figure}[t]
   \centering
@@
 
-While the testing conducted in the previous two sections provide an overview of how each application performs under controlled conditions, testing in the real world provides better insight into how a user may experience the final product.
-Unfortunately, there are many limitations in how testing can be conducted in the real world.
-First and foremost, due to the closed nature of of the existing solutions, collecting data while using them in real-world scenarios is impossible while leveraging the tools from the previous sections.
-Because of the reliance on using exact and repeatable testing conditions, a dynamic test of each application is much more difficult to conduct without direct access to the performance of the program, all data that could be collected otherwise would amount to quantitative data subject to interpretation.
-However, because of the logging tools that were developed in Section \ref{sec:DevelopingTestingAndMeasurementTools}, it is possible to collect data from the forked version of moonlight that shows the performance of the application over an entire streaming session.
-While it is impossible to predict and exhaustively test what the user may try to do with the final product, it is possible to test common use cases and evaluate whether the streaming experience is satisfactory.
-
 The first use case that was tested was streaming KiCad to the client machine to inspect the PCB layout of the custom board (Figure \ref{fig:RealWorldKicad}).
@@
 
+\vspace{3mm} % Get the header of the block onto the next page. This can be removed if formatting changes
 \begin{lstlisting}[style=plaintext,title=Statistics recorded while streaming KiCad (Figure \ref{fig:RealWorldKicad})]
@@
 \end{figure}
+
+\subsection{AI Training}\label{subsec:RealWorldAI}
 
@@
 \end{figure}
+
+\subsection{3D Rendering}\label{subsec:RealWorldBlender}
 
@@
 \end{figure}
+
+\subsection{Video Game Development}\label{subsec:RealWorldVideoGame}
 
@@
 Each test resulted in highly similar results, and while it does on average have slightly higher response times compared to RDP (Figures \ref{fig:InputDelay} and \ref{fig:ColorDelay}), it only falls behind by 5-10ms -- quicker than the blink of an eye.
-By not dropping any frames, and numerically proving it's ability to keep the stream above 60 frames per second, Moonlight proves to be a great solution for streaming demanding applications to a remote client without compromising its usability.
+Moonlight also provided an uncompromising experience throughout each of the real world use cases.
+By dropping zero frames across seven of the eight tests conducted for this thesis, and only 0.06\% of frames on the final test, and numerically proving it's ability to keep the stream above 60 frames per second in nearly every situation, Moonlight proves to be a great solution for streaming demanding applications to a remote client without compromising its usability.

### 2022-04-26 paper/main.tex
Clarify latency section and prepare for submission
This is the last thing that needed to be tweaked before the thesis is finalized

@@
 11pt, % The default document font size, options: 10pt, 11pt, 12pt
-%oneside, % Two side (alternating margins) for binding by default, uncomment to switch to one side
+oneside, % Two side (alternating margins) for binding by default, uncomment to switch to one side
 english, % ngerman for German
@@
 
+\usepackage{chronosys} % Required for timelines
+
 %----------------------------------------------------------------------------------------
@@
 
-	\noindent Date: 4/15/2022\\
+	\noindent Date: \today\\
 	\rule[0.5em]{25em}{0.5pt} % This prints a line to write the date

### 2022-04-26 paper/Chapters/Background.tex

@@
 
-It doesn't make sense to build or purchase a powerful PC for use at home and another power laptop to handle working on the go, especially with rising prices, so many people are moving to exclusively powerful laptops or looking for other solutions.
+\section{Summary}\label{sec:BackgroundSummary}
+
+Even after decades of technological innovation, every consumer computer is still a compromise between power, portability, and price.
+The desktop computer has been the best option for high power machines with a lower overall cost, but lacks the portability that has become a necessity in modern times.
+Laptops and gaming laptops have risen in popularity for their portability and flexibility, but their power comes at a cost premium that makes it inaccessible for many consumers.
+While thin clients are a viable solution for the corporate world the power is provided by a dedicated server and end-user costs are kept at a minimum, their lack in portability highlights a gap in the currently available technology.
+There currently isn't a solution for a cheap an portable device that can leverage the power of an existing computer.
+Specifically, it doesn't make sense to build or purchase a powerful PC for use at home and another power laptop to handle working on the go.

### 2022-04-26 paper/Chapters/Evaluation.tex

@@
 But the task of keeping two computer's internal clocks synchronized is not trivial.
+
+\begin{figure}[h]
+  \startchronology[startyear=0,stopyear=10, height=0pt, dates=false, arrow=false, box=true]
+  \chronograduation[event][dateselevation=0pt]{11}
+
+  \chronoevent[date=false]{2}{Key is pressed on client}
+  \chronoevent[date=false,markdepth=35pt]{6}{Key is received by host}
+  \chronoevent[date=false]{8}{Key is pressed on host}
+
+  \chronoperiode[textstyle=\raggedleft\colorbox{cyan!50}, color=cyan, startdate=false, bottomdepth=5pt, topheight=10pt, textdepth=-15pt,stopdate=false]{2}{6}{Network Latency}
+  \chronoperiode[textstyle=\colorbox{green!50}, color=green, startdate=false, bottomdepth=5pt, topheight=10pt, textdepth=-15pt,stopdate=false]{6}{8}{Application Latency}
+  \chronoperiode[textstyle=\colorbox{orange!50}, color=orange, startdate=false, bottomdepth=0pt, topheight=5pt, textdepth=-40pt,stopdate=false]{2}{8}{Total Latency}
+
+  \stopchronology
+  \caption[Timeline of a keystroke event]{Timeline of a keystroke event}
+  \label{fig:KeystrokeTimeline}
+\end{figure}
+
 Even after instructing the operating system to synchronize its clock with an internet time server, the two computers may still report different times.
 In some cases, the difference between the host and client computer's system time was noticed to be upwards of a full second after resynchronizing both clocks \cite{time.is}.
-This is partially accounted for in step \emph{6} of the above process, and the remaining inaccuracy can now be removed by subtracting the time difference between the when the client presses the button and the time the host executes that command.
-It is important to not subtract the entire round trip time of the input, as that is what is being testing in the following section, but rather just the time between when the client sends the input and when the host receives it.
-This ensures that any inaccuracies in each computer's internal clock will not impact the data.
+This is partially accounted for in step \emph{6} of the above process, but the remaining inaccuracy must be subtracted away.
+Figure \ref{fig:KeystrokeTimeline} shows the timeline of a single keystroke event; after the client presses a key, it must first be sent and received by the host computer, and then the host must execute that input.
+The time that this process takes is the total latency of the input.
+For the purposes of evaluating the different applications, this delay must be split into the time it takes for the host to receive the input, called the network latency, and the time it takes for the host to actually execute the input, called the application latency.
+Because the data loggers record the exact time that the test begins on both the client and host machines, the two times can be subtracted from each other to determine the network latency at the beginning of the test.
+After recording the data, the total latency of each input recorded by the data loggers is subtracted by this value so that the time remaining is the application latency.
+This ensures a fair and consistent comparison between each of the applications so that any inaccuracies in each computer's internal clock or the network speed has no impact on the data, and the only delay is that which is introduced by the application itself.
 
@@
 
-\vspace{3mm} % Get the header of the block onto the next page. This can be removed if formatting changes
+\vspace{15mm} % Get the header of the block onto the next page. This can be removed if formatting changes
 \begin{lstlisting}[style=plaintext,title=Statistics recorded while streaming KiCad (Figure \ref{fig:RealWorldKicad})]

### 2022-04-26 paper/Chapters/Hardware.tex

@@
 Instead, a better suited solution would be to utilize a \emph{Raspberry Pi Compute Module 4 (CM4)} which has \enquote{The power of Raspberry Pi 4 in a compact form factor for deeply embedded applications} \cite{rpi_cm4}.
-As this perfectly lines up with the requirements for the project, this thesis will be built using the CM4.
-However, due to it's small size and deeply embedded nature, the only way to communicate with the Compute Module is through two tiny mezzanine connectors as shown in Figure \ref{fig:rpi_cm4_mezzanine}.
+Preliminary investigations showed that the CM4 is a promising candidate for the system and hence was chosen to be the starting point for the development of this project.
+It's most important characteristic to keep in mind is due to it's small size and deeply embedded nature, the only way to communicate with the Compute Module is through two tiny mezzanine connectors as shown in Figure \ref{fig:rpi_cm4_mezzanine}.
 With no other pin outs for anything such as power, HDMI, or USB, the rest of the hardware must be built on a custom Printed Circuit Board (PCB) that tailors the CM4 for this thesis.
@@
   \captionsetup{width=.8\linewidth}
-  \caption[SMD Breakout Board]{A breakout board similar to what would be needer for the USB controller used on the custom PCB}
+  \caption[SMD Breakout Board]{A breakout board similar to what would be needed for the USB controller used on the custom PCB}
   \label{fig:smt_breakout}
@@
 The Raspberry Pi Compute Module 4 offered enough power to power the project while maintaining a low cost and slim profile that enabled it to fit in a small mobile device.
-One the CM4 was chosen, a custom PCB had to be designed and manufactured to tailor the Compute module to the needs of the project.
+Once the CM4 was chosen, a custom PCB had to be designed and manufactured to tailor the Compute Module to the needs of the project.
 This included choosing compatible electronic components, designing a board that provided the necessary functionality, and developing multiple iterations of the board to fix problems as they arose.
+After three iterations, a board was finally produced that could be used to develop the software portion of the thesis.

### 2022-04-26 paper/Chapters/Software.tex

@@
 
-\todoadvisor{Summary}
+\section{Summary}\label{sec:SoftwareSummary}
+
+In order to connect the client device to the host computer, a software solution had to be developed to compliment the hardware produced in the previous chapter.
+Instead of reinventing the wheel and attempting to build a solution from scratch, focus was turned towards Nvidia's GameStream protocol, which has proven itself to be a reliable and capable solution that fits the needs of this thesis.
+The open source project Moonlight has taken the closed source protocol and provides a usable API to build a custom client.
+A fork of Moonlight was made for this thesis, and work was done to ease the development of the project on the \emph{ARM} architecture, as well as to build a suite of testing tools to evaluate the application against the existing solutions.
